### 2024-10-12 paper/paper.qmd
Trial on glm on variables (answer,population,sample_size)

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
# Set a CRAN mirror
# Set a specific CRAN mirror (this example uses the US CRAN mirror)
options(repos = c(CRAN = "https://cran.rstudio.com/"))

# Example of installing a package after setting the mirror
install.packages("tidyverse")

library(tidyverse)
# Install janitor package if not installed
install.packages("janitor")

# Load the janitor package
library(janitor)

install.packages("caret")
install.packages("ROCR")
install.packages("rstanarm")
library(rstanarm)
install.packages("modelsummary")
library(modelsummary)

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)

# Load the dataset
polls <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))

```

```{r}

polls |>
  ggplot(aes(x = population, fill = answer)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  labs(
    x = "Population",
    y = "Number of respondents",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
# Finding Median of Sample Size
median_sample_size <- median(polls$sample_size, na.rm = TRUE)
print(median_sample_size)

# The U.S. uses an Electoral College system, where each state is assigned a certain number of electoral votes based on its population (each state gets a number of electors equal to its total number of Senators and Representatives in Congress).
# Most states use a winner-takes-all system, where the candidate who wins the popular vote in the state gets all its electoral votes. (Exceptions: Nebraska and Maine, which allocate electoral votes proportionally.)
# A total of 538 electoral votes are available. To win the presidency, a candidate must secure at least 270 electoral votes.

# Differentiating sample sizes in presidential election polls is important because larger samples generally provide more reliable results with smaller margins of error, while smaller samples may result in less precise and potentially skewed outcomes. Additionally, larger sample sizes better capture the diversity of key demographic groups, leading to more accurate reflections of voter behavior.
```

```{r}

library(ggplot2)
library(dplyr)

library(ggplot2)
library(dplyr)
library(tidyr)

# Define population levels (adjust based on your actual data)
population_levels <- c("Adults", "Voters", "Likely Voters", "Registered Voters")  # Update these based on your dataset

# Ensure all factor levels are present and handle missing combinations
polls_clean <- polls %>%
  mutate(
    population = factor(population, levels = population_levels),
    sample_category = ifelse(sample_size < 1006, "Less than 1006", "1006 or more"),
    answer = as.factor(answer)  # Convert answer to factor if it represents candidates
  ) %>%
  complete(population, answer, sample_category, fill = list(n = 0))  # Fill missing combinations

# Plot with facet_wrap
polls_clean %>%
  ggplot(aes(x = population, fill = answer)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  labs(
    x = "Population",
    y = "Number of respondents",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  facet_wrap(~sample_category, scales = "free_y")

```

```{r}
set.seed(304)

polls_reduced <- 
  polls_clean |> 
  slice_sample(n = 1000)

political_preferences <-
  stan_glm(
    answer ~  + sample_category + population,
    data = polls_reduced,
    family = binomial(link = "logit"),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = 
      normal(location = 0, scale = 2.5, autoscale = TRUE),
    seed = 853
  )

saveRDS(
  political_preferences,
  file = "political_preferences.rds"
)

```

```{r}
political_preferences <-
  readRDS(file = "political_preferences.rds")

```

```{r}
modelsummary(
  list(
    "Support Biden" = political_preferences
  ),
  statistic = "mad"
  )

```

```{r}
modelplot(political_preferences, conf_level = 0.9) +
  labs(x = "90 per cent credibility interval")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References

### 2024-10-12 scripts/00-simulate_data.R

@@ -11,42 +11,48 @@
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
-
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
-
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
 
 # Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
+# Load necessary libraries
+library(tidyverse)
+
+# Set seed for reproducibility
+set.seed(853)
+
+# Number of observations for the simulation
+num_obs <- 1000
+
+# Simulating data based on variables in the provided president polls CSV
+president_poll_simulation <- tibble(
+  pollster = sample(c("InsiderAdvantage", "YouGov", "ActiVote"), size = num_obs, replace = TRUE),
+  state = sample(c("Pennsylvania", "Arizona", "Texas", "Florida", "Ohio"), size = num_obs, replace = TRUE),
+  sample_size = sample(400:1500, size = num_obs, replace = TRUE),
+  methodology = sample(c("Text", "App Panel", "Online Panel"), size = num_obs, replace = TRUE),
+  population = sample(c("lv", "rv"), size = num_obs, replace = TRUE), # Likely Voters (lv), Registered Voters (rv)
+  candidate_name = sample(c("Kamala Harris", "Donald Trump", "Jill Stein"), size = num_obs, replace = TRUE),
+  party = case_when(
+    candidate_name == "Kamala Harris" ~ "DEM",
+    candidate_name == "Donald Trump" ~ "REP",
+    candidate_name == "Jill Stein" ~ "GRE"
   ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
+  pct = rnorm(num_obs, mean = 50, sd = 5), # Simulating percentage of support
[…]

### 2024-10-12 scripts/03-clean_data.R

@@ -1,44 +1,37 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw polling data, focusing on Kamala Harris and Donald Trump
+# Author: [Your Name]
+# Date: [Today's Date]
+# Contact: [Your Email]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Requires 'president_polls.csv' file in 'data/01-raw_data/'
+# Any other information needed? This script filters polling data for Harris and Trump only.
 
 #### Workspace setup ####
-library(tidyverse)
+# Load necessary library
+library(dplyr)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Read the dataset
+polls <- read.csv("data/01-raw_data/president_polls.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+# Filter for only Kamala Harris and Donald Trump
+polls_clean <- polls %>%
+  filter(answer %in% c("Harris", "Trump")) %>%
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
[…]

### 2024-10-12 paper/paper.qmd
changed author names

@@
 author: 
-  - First author
-  - Another author
+  - Tim Chen
+  - Steven Li
+  - Tommy Fu
 thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
@@
 #| message: false
-# Set a CRAN mirror
-# Set a specific CRAN mirror (this example uses the US CRAN mirror)
-options(repos = c(CRAN = "https://cran.rstudio.com/"))
-
-# Example of installing a package after setting the mirror
-install.packages("tidyverse")
 
 library(tidyverse)
-# Install janitor package if not installed
-install.packages("janitor")
-
-# Load the janitor package
-library(janitor)
-
-install.packages("caret")
-install.packages("ROCR")
-install.packages("rstanarm")
-library(rstanarm)
-install.packages("modelsummary")
-library(modelsummary)
-
+library(palmerpenguins)
 ```
@@
 ```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
-# Load the dataset
-polls <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = "none") +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
 
-```{r}
+Talk more about it.
 
-polls |>
-  ggplot(aes(x = population, fill = answer)) +
-  stat_count(position = "dodge") +
-  theme_minimal() +
-  labs(
-    x = "Population",
-    y = "Number of respondents",
-    fill = "Voted for"
-  ) +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
 ```{r}
-# Finding Median of Sample Size
-median_sample_size <- median(polls$sample_size, na.rm = TRUE)
-print(median_sample_size)
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| warning: false
+#| message: false
 
-# The U.S.
-uses an Electoral College system, where each state is assigned a certain number of electoral votes based on its population (each state gets a number of electors equal to its total number of Senators and Representatives in Congress).
-# Most states use a winner-takes-all system, where the candidate who wins the popular vote in the state gets all its electoral votes.
-(Exceptions: Nebraska and Maine, which allocate electoral votes proportionally.)
-# A total of 538 electoral votes are available. To win the presidency, a candidate must secure at least 270 electoral votes.
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
-# Differentiating sample sizes in presidential election polls is important because larger samples generally provide more reliable results with smaller margins of error, while smaller samples may result in less precise and potentially skewed outcomes.
-Additionally, larger sample sizes better capture the diversity of key demographic groups, leading to more accurate reflections of voter behavior.
-```
-
-```{r}
-
-library(ggplot2)
-library(dplyr)
-
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-
-# Define population levels (adjust based on your actual data)
-population_levels <- c("Adults", "Voters", "Likely Voters", "Registered Voters")  # Update these based on your dataset
-
-# Ensure all factor levels are present and handle missing combinations
-polls_clean <- polls %>%
-  mutate(
-    population = factor(population, levels = population_levels),
-    sample_category = ifelse(sample_size < 1006, "Less than 1006", "1006 or more"),
-    answer = as.factor(answer)  # Convert answer to factor if it represents candidates
-  ) %>%
-  complete(population, answer, sample_category, fill = list(n = 0))  # Fill missing combinations
-
-# Plot with facet_wrap
-polls_clean %>%
-  ggplot(aes(x = population, fill = answer)) +
-  stat_count(position = "dodge") +
+analysis_data |> 
+  ggplot(aes(x = width, y = length)) +
+  geom_point(alpha = 0.8) +
   theme_minimal() +
-  labs(
-    x = "Population",
-    y = "Number of respondents",
-    fill = "Voted for"
-  ) +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  facet_wrap(~sample_category, scales = "free_y")
-
-```
-
-```{r}
-set.seed(304)
-
-polls_reduced <- 
-  polls_clean |> 
-  slice_sample(n = 1000)
-
-political_preferences <-
-  stan_glm(
-    answer ~  + sample_category + population,
-    data = polls_reduced,
-    family = binomial(link = "logit"),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = 
-      normal(location = 0, scale = 2.5, autoscale = TRUE),
-    seed = 853
-  )
-
-saveRDS(
-  political_preferences,
-  file = "political_preferences.rds"
-)
-
-```
-
-```{r}
-political_preferences <-
-  readRDS(file = "political_preferences.rds")
-
-```
-
-```{r}
-modelsummary(
-  list(
-    "Support Biden" = political_preferences
-  ),
-  statistic = "mad"
-  )
-
-```
-
-```{r}
-modelplot(political_preferences, conf_level = 0.9) +
-  labs(x = "90 per cent credibility interval")
+  labs(x = "Wing width (mm)",
+       y = "Wing length (mm)")
 ```
@@
 
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
 ```
@@
 
+plot(first_model, "trace")
+
+plot(first_model, "rhat")
 ```

### 2024-10-18 scripts/03-clean_data.R
code update from class

@@ -6,32 +6,35 @@
 # License: MIT
 # Pre-requisites: Requires 'president_polls.csv' file in 'data/01-raw_data/'
 # Any other information needed? This script filters polling data for Harris and Trump only.
-
+install.packages("janitor")
 #### Workspace setup ####
 # Load necessary library
 library(dplyr)
+library(tidyverse)
+library(janitor)
 
 # Read the dataset
-polls <- read.csv("data/01-raw_data/president_polls.csv")
+data <- read_csv("data/01-raw_data/president_polls.csv") |>
+  clean_names()
 
-# Filter for only Kamala Harris and Donald Trump
-polls_clean <- polls %>%
-  filter(answer %in% c("Harris", "Trump")) %>%
+# Filter data to Harris estimates based on high-quality polls after she declared
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
   mutate(
-    # Clean up population descriptions
-    population = case_when(
-      population == "a" ~ "Adults",
-      population == "v" ~ "Voters",
-      population == "lv" ~ "Likely Voters",
-      population == "rv" ~ "Registered Voters",
-      TRUE ~ population  # Keep original value if no match
-    )
+    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
-# Select only the desired columns and drop rows with NA values
-polls_clean <- polls_clean %>%
-  select(answer, state, population, sample_size) %>%
-  drop_na()
-
 # Save the cleaned dataset as a new CSV file
-write.csv(polls_clean, "data/02-analysis_data/analysis_data.csv", row.names = FALSE)
+write.csv(just_harris_high_quality, "data/02-analysis_data/analysis_data.csv", row.names = FALSE)
+
+
+

### 2024-10-18 scripts/06-model_data.R

@@ -13,25 +13,47 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+just_harris_high_quality <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+model_date <- lm(pct ~ end_date, data = just_harris_high_quality)
+
+# Model 2: pct as a function of end_date and pollster
+model_date_pollster <- lm(pct ~ end_date + pollster, data = just_harris_high_quality)
+
+# Augment data with model predictions
+just_harris_high_quality <- just_harris_high_quality |>
+  mutate(
+    fitted_date = predict(model_date),
+    fitted_date_pollster = predict(model_date_pollster)
   )
 
+# Plot model predictions
+# Model 1
+ggplot(just_harris_high_quality, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
+
+# Model 2
+ggplot(just_harris_high_quality, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_date_pollster), color = "blue", linetype = "dotted") +
+  facet_wrap(vars(pollster)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
 
 #### Save model ####
 saveRDS(
-  first_model,
+  model_date,
   file = "models/first_model.rds"
 )
 
+saveRDS(
+  model_date_pollster,
+  file = "models/second_model.rds"
+)
 
+
+modelsummary(models = list("Model 1" = model_date, "Model 2" = model_date_pollster))

### 2024-10-18 scripts/03-clean_data.R
updated cleaned data and attemped to do eda on data

@@ -1,40 +1,51 @@
 #### Preamble ####
-# Purpose: Cleans the raw polling data, focusing on Kamala Harris and Donald Trump
+# Purpose: Prepare polling data and build a linear regression model to predict popular vote for Kamala Harris and Donald Trump.
 # Author: [Your Name]
 # Date: [Today's Date]
 # Contact: [Your Email]
 # License: MIT
 # Pre-requisites: Requires 'president_polls.csv' file in 'data/01-raw_data/'
-# Any other information needed? This script filters polling data for Harris and Trump only.
+# Any other information needed? This script builds an MLR model for Harris vs Trump using polling data.
+
+#### Install necessary libraries ####
 install.packages("janitor")
+install.packages("lubridate")
+
 #### Workspace setup ####
-# Load necessary library
+# Load necessary libraries
 library(dplyr)
 library(tidyverse)
 library(janitor)
+library(lubridate)
 
-# Read the dataset
+#### Data Import and Cleaning ####
+# Read the dataset and clean column names
 data <- read_csv("data/01-raw_data/president_polls.csv") |>
   clean_names()
 
-# Filter data to Harris estimates based on high-quality polls after she declared
-just_harris_high_quality <- data |>
+#### Filter Data for Both Candidates ####
+# Filter data for Kamala Harris and Donald Trump based on high-quality polls
+filtered_data <- data |>
   filter(
-    candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+    candidate_name %in% c("Kamala Harris", "Donald Trump"), # Keep only Harris and Trump
+    numeric_grade >= 2.7 # Filter high-quality polls
   ) |>
   mutate(
-    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
-    end_date = mdy(end_date)
+    state = if_else(is.na(state), "National", state), # Fix for national polls
+    end_date = mdy(end_date) # Convert end date to date format
   ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  filter( # Filter polls after each candidate declared
+    (candidate_name == "Kamala Harris" & end_date >= as.Date("2024-07-21")) | 
+    (candidate_name == "Donald Trump" & end_date >= as.Date("2024-07-21"))
[…]

### 2024-10-18 scripts/05-exploratory_data_analysis.R

@@ -13,25 +13,22 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+# Summary statistics for numeric variables
+summary(analysis_data)
+
+# Detailed summary of specific variables
+analysis_data %>%
+  summarize(
+    avg_pct = mean(pct, na.rm = TRUE),
+    median_pct = median(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    avg_sample_size = mean(sample_size, na.rm = TRUE),
+    min_sample_size = min(sample_size, na.rm = TRUE),
+    max_sample_size = max(sample_size, na.rm = TRUE)
   )
 
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
+analysis_data
 

### 2024-10-18 scripts/00-simulate_data.R
initial code

@@ -1,52 +1,95 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Purpose: Simulates a dataset of US pollster polling outcomes
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 18 October 2024
+# Contacts: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
+# Pre-requisites:
+# - Packages `tidyverse`, and `lubridate` must be installed and loaded
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+library(lubridate)
 
+#### Data expectations ####
+# Pollster Data:
+# - Columns: poll_id, pollster, pollster_rating, method, state, voter_type, party, winner, percentage
+# - poll_id should be unique
+# - Date should be between 2024-01-01 and 2024-10-18
+# - pollster_rating should be between 1 and 5
+# - method should be one of: "Online", "Phone", "Mixed"
+# - state should be valid US state abbreviations
+# - voter_type should be one of: "Likely Voters", "Registered Voters", "All Adults"
+# - party should be one of: "Republican", "Democrat"
+# - winner should be one of: "Trump", "Harris"
+# - percentage should be between 0 and 100
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+set.seed(2024)
+
+########################### Simulate Pollster Data ############################ 
+num_polls <- 1000
+
+pollster_data <- tibble(
+  poll_id = seq(1, num_polls),
+  date = sample(seq(as.Date('2024-01-01'), as.Date('2024-10-18'), by="day"), 
+                num_polls, replace=TRUE),
+  pollster = sample(c("PollsterA", "PollsterB", "PollsterC", 
[…]

### 2024-10-18 scripts/01-test_simulated_data.R

@@ -1,89 +1,90 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Purpose: Testing simulated pollster data for the 2024 US presidential election
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 18 October 2024
+# Contacts: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+  # - The `tidyverse`, `lubridate`, and `testthat` package must be installed and loaded
 
 #### Workspace setup ####
 library(tidyverse)
+library(testthat)
+library(lubridate)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Read in the data
+pollster_data <- read_csv("data/00-simulated_data/simulated_pollster_data.csv")
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+########################### Tests for Pollster Data########################### 
 
+test_that("Pollster data has correct structure", {
+  expected_cols <- c("poll_id", "date", "pollster", "pollster_rating", "method", 
+                     "state", "voter_type", "party", "winner", "percentage")
+  expect_equal(colnames(pollster_data), expected_cols)
+  expect_true(is.numeric(pollster_data$poll_id))
+  expect_true(inherits(pollster_data$date, "Date"))
+  expect_true(is.character(pollster_data$pollster))
+  expect_true(is.numeric(pollster_data$pollster_rating))
+  expect_true(is.character(pollster_data$method))
+  expect_true(is.character(pollster_data$state))
+  expect_true(is.character(pollster_data$voter_type))
+  expect_true(is.character(pollster_data$party))
+  expect_true(is.character(pollster_data$winner))
+  expect_true(is.numeric(pollster_data$percentage))
+})
 
-#### Test data ####
+test_that("Pollster data has valid values", {
[…]

### 2024-10-18 scripts/03-clean_data.R

@@ -1,44 +1,50 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Prepare polling data for the multi-linear regression model
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 18 October 2024
+# Contacts: 
+  # - Tim Chen: timwt.chen@mail.utoronto.ca
+  # - Steven Li: stevency.li@mail.utoronto.ca
+  # - Tommy Fu: tommy.fu@mail.utoronto.ca
+# Pre-requisites:
+  # - Requires downloading presidential polls data as 'president_polls.csv' file in 'data/01-raw_data/'
+  # - Packages `tidyverse`, `dplyr`, `janitor`, and `lubridate` must be installed and loaded
 
 #### Workspace setup ####
+# Load necessary libraries
+library(dplyr)
 library(tidyverse)
+library(janitor)
+library(lubridate)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+########################### Data Import and Cleaning########################### 
+# Read the dataset and clean column names
+data <- read_csv("data/01-raw_data/president_polls.csv") |>
+  clean_names()
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
+###################### Filter Data for Both Candidates ########################
+# Filter data for Kamala Harris and Donald Trump based on high-quality polls
+filtered_data <- data |>
+  # Select only the specified columns
+  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, methodology, 
+         transparency_score, sample_size, population, population_full, party, 
+         answer, pct, state, end_date
   ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
[…]

### 2024-10-18 scripts/01-test_simulated_data.R
tested simulated data

@@ -1,89 +1,90 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Purpose: Testing simulated pollster data for the 2024 US presidential election
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 18 October 2024
+# Contacts: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# - 00-simulate_data.R must have been run
+# - The `tidyverse`, `lubridate`, and `testthat` package must be installed and loaded
 
 #### Workspace setup ####
 library(tidyverse)
+library(testthat)
+library(lubridate)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Read in the data
+pollster_data <- read_csv("data/00-simulated_data/simulated_pollster_data.csv")
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+########################### Tests for Pollster Data########################### 
 
+test_that("Pollster data has correct structure", {
+  expected_cols <- c("poll_id", "date", "pollster", "pollster_rating", "method", 
+                     "state", "voter_type", "party", "winner", "percentage")
+  expect_equal(colnames(pollster_data), expected_cols)
+  expect_true(is.numeric(pollster_data$poll_id))
+  expect_true(inherits(pollster_data$date, "Date"))
+  expect_true(is.character(pollster_data$pollster))
+  expect_true(is.numeric(pollster_data$pollster_rating))
+  expect_true(is.character(pollster_data$method))
+  expect_true(is.character(pollster_data$state))
+  expect_true(is.character(pollster_data$voter_type))
+  expect_true(is.character(pollster_data$party))
+  expect_true(is.character(pollster_data$winner))
+  expect_true(is.numeric(pollster_data$percentage))
+})
 
-#### Test data ####
[…]

### 2024-10-18 paper/paper.qmd
An attempt on building a model

@@
 ---
-title: "My title"
+title: "Forecasting the 2024 U.S. Presidential Election"
 subtitle: "My subtitle if needed"
 author: 
-  - Tim Chen
-  - Steven Li
-  - Tommy Fu
+  - First author
+  - Another author
 thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
@@
 #| message: false
+# Set a CRAN mirror
+# Set a specific CRAN mirror (this example uses the US CRAN mirror)
+install.packages("here")
+
+options(repos = c(CRAN = "https://cran.rstudio.com/"))
+
+# Example of installing a package after setting the mirror
+install.packages("tidyverse")
 
 library(tidyverse)
-library(palmerpenguins)
+# Install janitor package if not installed
+install.packages("janitor")
+
+# Load the janitor package
+library(janitor)
+
+install.packages("caret")
+install.packages("ROCR")
+install.packages("rstanarm")
+library(rstanarm)
+install.packages("modelsummary")
+library(modelsummary)
+
+library(here)
+
 ```
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+# Load necessary libraries
+library(dplyr)
+library(tidyr)
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Load the dataset
+data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
 ```
 
-Talk more about it.
+```{r}
+# Build the GLM model
+glm_model <- glm(candidate_binary ~ pct + sample_size + methodology + pollscore, 
+                 data = data, 
+                 family = binomial)
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Summary of the model
+summary(glm_model)
+
+```
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+# Make predictions
+predictions <- predict(glm_model, newdata = data, type = "response")
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Convert probabilities into binary outcomes (1 for Harris win, 0 for Trump win)
+data$predicted_outcome <- ifelse(predictions > 0.5, 1, 0)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# View the prediction results
+head(data)
+
+```
+
+```{r}
+# Confusion matrix
+table(data$candidate_binary, data$predicted_outcome)
+
+# Calculate accuracy
+accuracy <- mean(data$candidate_binary == data$predicted_outcome)
+accuracy
+
+```
+
+```{r}
+# Plot predicted probabilities
+ggplot(data, aes(x = pct, y = predictions, color = factor(candidate_binary))) +
+  geom_point() +
+  labs(title = "Predicted Probability of Harris Winning vs Poll Percentage",
+       x = "Polling Percentage", y = "Predicted Probability")
+
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-10-18 scripts/03-clean_data.R

@@ -41,11 +41,14 @@ filtered_data <- data |>
   mutate(
     num_supporters = round((pct / 100) * sample_size, 0), # Convert percentage to number of supporters
     candidate_binary = ifelse(candidate_name == "Kamala Harris", 1, 0) # Binary encoding for Harris (1) and Trump (0)
-  )
-
+  ) |>
+  # Select only the specified columns
+  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, methodology, 
+         transparency_score, sample_size, population, population_full, party, 
+         answer, pct, state, candidate_name, end_date, candidate_binary, num_supporters
+  ) |>
+  drop_na()
 
 # Save the cleaned dataset as a new CSV file
 write.csv(filtered_data, "data/02-analysis_data/analysis_data.csv", row.names = FALSE)
-
-
 

### 2024-10-18 scripts/05-exploratory_data_analysis.R

@@ -1,34 +1,108 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Perform Exploratory Data Analysis (EDA) on polling data for Kamala Harris and Donald Trump, and save the plots as images.
+# Author: [Your Name]
+# Date: [Today's Date]
+# Contact: [Your Email]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Requires cleaned 'analysis_data.csv' with filtered data for Harris and Trump
 
+#### Install necessary libraries ####
+install.packages("ggplot2")
+install.packages("dplyr")
+install.packages("naniar")
+install.packages("tidyverse")
 
 #### Workspace setup ####
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(naniar)
 library(tidyverse)
-library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+#### Load the cleaned dataset ####
+# Read the cleaned CSV file
+filtered_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
+#### 1. Overview of the Dataset ####
+# Overview of the data structure
+print("Overview of dataset:")
+glimpse(filtered_data)
+
 # Summary statistics for numeric variables
-summary(analysis_data)
+print("Summary statistics:")
+summary(filtered_data)
 
-# Detailed summary of specific variables
-analysis_data %>%
-  summarize(
-    avg_pct = mean(pct, na.rm = TRUE),
-    median_pct = median(pct, na.rm = TRUE),
-    sd_pct = sd(pct, na.rm = TRUE),
-    avg_sample_size = mean(sample_size, na.rm = TRUE),
-    min_sample_size = min(sample_size, na.rm = TRUE),
-    max_sample_size = max(sample_size, na.rm = TRUE)
-  )
+# Checking for missing values
+print("Visualizing missing data:")
+vis_miss(filtered_data)
 
-analysis_data
+#### 2. Univariate Analysis ####
 
+# 2.1 Distribution of polling percentages for Harris and Trump
[…]

### 2024-10-18 scripts/06-model_data.R

@@ -13,47 +13,23 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-just_harris_high_quality <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 ### Model data ####
-model_date <- lm(pct ~ end_date, data = just_harris_high_quality)
-
-# Model 2: pct as a function of end_date and pollster
-model_date_pollster <- lm(pct ~ end_date + pollster, data = just_harris_high_quality)
-
-# Augment data with model predictions
-just_harris_high_quality <- just_harris_high_quality |>
-  mutate(
-    fitted_date = predict(model_date),
-    fitted_date_pollster = predict(model_date_pollster)
+glm_model <- 
+  stan_glm(
+    formula = candidate_binary ~ numeric_grade + pollscore + transparency_score + sample_size + party + state + num_supporters,
+    data = analysis_data,
+    family = binomial(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 304
   )
-
-# Plot model predictions
-# Model 1
-ggplot(just_harris_high_quality, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
-
-# Model 2
-ggplot(just_harris_high_quality, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_date_pollster), color = "blue", linetype = "dotted") +
-  facet_wrap(vars(pollster)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
 
 #### Save model ####
 saveRDS(
-  model_date,
-  file = "models/first_model.rds"
+  glm_model,
+  file = "models/glm_model.rds"
 )
 
-saveRDS(
-  model_date_pollster,
-  file = "models/second_model.rds"
-)
-
-
-modelsummary(models = list("Model 1" = model_date, "Model 2" = model_date_pollster))

### 2024-10-20 paper/paper.qmd
testing git
complete a model with electoral college

@@
 # Set a specific CRAN mirror (this example uses the US CRAN mirror)
-install.packages("here")
 
@@
 
+install.packages("here")
 library(tidyverse)
@@
 
+Simple Linear Regression for Kamala Harris and Donald Trump
+
+Kamala Harris
+We will perform a simple linear regression using pct (poll percentage) as the dependent variable and sample_size as the independent variable.
+
 ```{r}
-# Build the GLM model
-glm_model <- glm(candidate_binary ~ pct + sample_size + methodology + pollscore, 
-                 data = data, 
-                 family = binomial)
+# Load necessary library
+library(ggplot2)
+
+# Filter the data for only "Kamala Harris"
+harris_data <- subset(data, candidate_name == "Kamala Harris")
+
+# Perform a simple linear regression with pct as the dependent variable and sample_size as the independent variable
+model <- lm(pct ~ sample_size, data = harris_data)
 
 # Summary of the model
-summary(glm_model)
+summary(model)
+
+# Plot the relationship for Kamala Harris
+ggplot(harris_data, aes(x = sample_size, y = pct)) +
+  geom_point() +
+  geom_smooth(method = "lm", col = "blue") +
+  labs(title = "Linear Regression of Percentage vs Sample Size for Kamala Harris",
+       x = "Sample Size",
+       y = "Percentage")
 
@@
 
+Donald Trump
+We will repeat the same process for Donald Trump.
+
 ```{r}
-# Make predictions
-predictions <- predict(glm_model, newdata = data, type = "response")
+# Load necessary library
+library(ggplot2)
 
-# Convert probabilities into binary outcomes (1 for Harris win, 0 for Trump win)
-data$predicted_outcome <- ifelse(predictions > 0.5, 1, 0)
+# Filter the data for only "Kamala Harris"
+trump_data <- subset(data, candidate_name == "Donald Trump")
 
-# View the prediction results
-head(data)
+# Perform a simple linear regression with pct as the dependent variable and sample_size as the independent variable
+model <- lm(pct ~ sample_size, data = trump_data)
+
+# Summary of the model
+summary(model)
+
+# Plot the relationship for Kamala Harris
+ggplot(trump_data, aes(x = sample_size, y = pct)) +
+  geom_point() +
+  geom_smooth(method = "lm", col = "blue") +
+  labs(title = "Linear Regression of Percentage vs Sample Size for Donald",
+       x = "Sample Size",
+       y = "Percentage")
 
@@
 
+Multiple Linear Regression (MLR) for Kamala Harris and Donald Trump
+
+Kamala Harris
+In this step, we will build a multiple linear regression (MLR) model for Kamala Harris using additional predictors like numeric_grade, pollscore, transparency_score, and state.
+
 ```{r}
-# Confusion matrix
-table(data$candidate_binary, data$predicted_outcome)
+# Convert categorical variables to factors for Kamala Harris
+harris_data$state <- as.factor(harris_data$state)
+harris_data$methodology <- as.factor(harris_data$methodology)
 
-# Calculate accuracy
-accuracy <- mean(data$candidate_binary == data$predicted_outcome)
-accuracy
+# Build MLR model for Kamala Harris
+mlr_harris_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, data = harris_data)
+
+# Summary of the MLR model for Harris
+summary(mlr_harris_model)
+
+# Diagnostic plots for the MLR model
+par(mfrow = c(2, 2))
+plot(mlr_harris_model)
+par(mfrow = c(1, 1))  # Reset plot layout
 
@@
 
+Similarly, we build an MLR model for Donald Trump.
 ```{r}
-# Plot predicted probabilities
-ggplot(data, aes(x = pct, y = predictions, color = factor(candidate_binary))) +
-  geom_point() +
-  labs(title = "Predicted Probability of Harris Winning vs Poll Percentage",
-       x = "Polling Percentage", y = "Predicted Probability")
+# Convert categorical variables to factors for Trump
+trump_data$state <- as.factor(trump_data$state)
+trump_data$methodology <- as.factor(trump_data$methodology)
+
+# Build MLR model for Donald Trump
+mlr_trump_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, data = trump_data)
+
+# Summary of the MLR model for Trump
+summary(mlr_trump_model)
+
+# Diagnostic plots for the MLR model
+par(mfrow = c(2, 2))
+plot(mlr_trump_model)
+par(mfrow = c(1, 1))  # Reset plot layout
+```
+
+Check Multicollinearity using Variance Inflation Factor (VIF)
+
+```{r}
+# Load the car package for VIF
+if (!require(car)) install.packages("car")
+library(car)
+
+# Check VIF for Kamala Harris model
+vif(mlr_harris_model)
+
+# Refine the model by removing less significant predictors (e.g., methodology and transparency_score)
+mlr_harris_model_refined <- lm(pct ~ numeric_grade + pollscore + sample_size + state, data = harris_data)
+
+# Summary of the refined Harris model
+summary(mlr_harris_model_refined)
+
+# Check VIF for the refined model
+vif(mlr_harris_model_refined)
+
+# Plot diagnostic plots for the refined model
+par(mfrow = c(2, 2))
+plot(mlr_harris_model_refined)
+par(mfrow = c(1, 1))  # Reset plot layout
 
@@
 
-Talk way more about it.
+Stepwise Model Selection for Optimization
+
+```{r}
+# Perform stepwise selection to optimize the Harris model
+step_model <- step(mlr_harris_model_refined, direction = "both")
+
+# Summary of the stepwise model
+summary(step_model)
+
+# Plot diagnostic plots for the stepwise model
+par(mfrow = c(2, 2))
+plot(step_model)
+par(mfrow = c(1, 1))  # Reset layout
+
+```
+
+Final Models and Predictions for Kamala Harris and Donald Trump
+
+```{r}
+# Final models for Harris and Trump
+mlr_harris_model_final <- lm(pct ~ pollscore + log(sample_size) + state, data = harris_data)
+mlr_trump_model_final <- lm(pct ~ pollscore + log(sample_size) + state, data = trump_data)
+
+# Predict poll percentages for Kamala Harris
+harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
+
+# Predict poll percentages for Donald Trump
+trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
+
+```
+
+Electoral College Votes Prediction
+In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
+
+```{r}
+# Define electoral votes for each state (and Washington, D.C.)
+electoral_votes <- data.frame(
+  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", 
+            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
+            "Maine CD-1", "Maine CD-2", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
+            "Missouri", "Montana", "Nebraska", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", "Nevada", 
+            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
+            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
+            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
+            "District of Columbia"),
+  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 2, 1, 1, 10, 11, 15, 10, 6, 10, 
+                      3, 5, 1, 1, 2, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3)
+)
+
+# Aggregate predictions by state for Harris and Trump
+harris_state_avg <- aggregate(predicted_pct_harris ~ state, data = harris_data, FUN = mean)
+trump_state_avg <- aggregate(predicted_pct_trump ~ state, data = trump_data, FUN = mean)
+
+# Merge predictions for both candidates
+prediction_comparison <- merge(harris_state_avg, trump_state_avg, by = "state", all.x = TRUE)
+
+# Merge the electoral votes with the predictions
+prediction_comparison <- merge(prediction_comparison, electoral_votes, by = "state", all.x = TRUE)
+
+# Determine the winner for each state
+prediction_comparison$winner <- ifelse(prediction_comparison$predicted_pct_harris > prediction_comparison$predicted_pct_trump, "Harris", "Trump")
+
+# Calculate total electoral votes for Kamala Harris
+harris_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Harris"], na.rm = TRUE)
+
+# Calculate total electoral votes for Donald Trump
+trump_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Trump"], na.rm = TRUE)
+
+# Print the results
+print(paste("Harris Electoral Votes:", harris_electoral_votes))
+print(paste("Trump Electoral Votes:", trump_electoral_votes))
+
+# Determine the predicted winner
+if (harris_electoral_votes >= 270) {
+  print("Kamala Harris is predicted to win the 2024 election.")
+} else if (trump_electoral_votes >= 270) {
+  print("Donald Trump is predicted to win the 2024 election.")
+} else {
+  print("No candidate reached 270 electoral votes.")
+}
+
+```
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+# #| echo: false
+# #| eval: true
+# #| warning: false
+# #| message: false
 
-library(rstanarm)
+# library(rstanarm)
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# first_model <-
+#   readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+# #| echo: false
+# #| eval: true
+# #| label: tbl-modelresults
+# #| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+# #| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# modelsummary::modelsummary(
+#   list(
+#     "First model" = first_model
+#   ),
+#   statistic = "mad",
+#   fmt = 2
+# )
 ```

### 2024-10-20 paper/paper.qmd
sample for idealied methodology for budgeting

--- 2024-10-18 paper/paper.qmd
@@
 ---
-title: "Forecasting the 2024 U.S. Presidential Election"
+title: "My title"
 subtitle: "My subtitle if needed"
 author: 
-  - First author
-  - Another author
+  - Tim Chen
+  - Steven Li
+  - Tommy Fu
 thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
@@
 #| message: false
-# Set a CRAN mirror
-# Set a specific CRAN mirror (this example uses the US CRAN mirror)
-install.packages("here")
-
-options(repos = c(CRAN = "https://cran.rstudio.com/"))
-
-# Example of installing a package after setting the mirror
-install.packages("tidyverse")
 
 library(tidyverse)
-# Install janitor package if not installed
-install.packages("janitor")
-
-# Load the janitor package
-library(janitor)
-
-install.packages("caret")
-install.packages("ROCR")
-install.packages("rstanarm")
-library(rstanarm)
-install.packages("modelsummary")
-library(modelsummary)
-
-library(here)
-
+library(palmerpenguins)
 ```
@@
 ```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
-# Load the dataset
-data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = "none") +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
 
-```{r}
-# Build the GLM model
-glm_model <- glm(candidate_binary ~ pct + sample_size + methodology + pollscore, 
-                 data = data, 
-                 family = binomial)
+Talk more about it.
 
-# Summary of the model
-summary(glm_model)
-
-```
-
-```{r}
-# Make predictions
-predictions <- predict(glm_model, newdata = data, type = "response")
-
-# Convert probabilities into binary outcomes (1 for Harris win, 0 for Trump win)
-data$predicted_outcome <- ifelse(predictions > 0.5, 1, 0)
-
-# View the prediction results
-head(data)
-
-```
-
-```{r}
-# Confusion matrix
-table(data$candidate_binary, data$predicted_outcome)
-
-# Calculate accuracy
-accuracy <- mean(data$candidate_binary == data$predicted_outcome)
-accuracy
-
-```
-
-```{r}
-# Plot predicted probabilities
-ggplot(data, aes(x = pct, y = predictions, color = factor(candidate_binary))) +
-  geom_point() +
-  labs(title = "Predicted Probability of Harris Winning vs Poll Percentage",
-       x = "Polling Percentage", y = "Predicted Probability")
-
-```
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
-# Additional data details
+## Idealized Methodology and Survey Design for Forecasting the US Presidential Election
 
-# Model details {#sec-model-details}
+Objective: The goal of this methodology is to forecast the outcome of the U.S.
+presidential election using a representative sample, ensuring accuracy, transparency, and cost-efficiency with a budget of $100,000.
 
-## Posterior predictive check
+1. Sampling Approach
+1.1. Target Population
+The target population is the U.S.
+voting-eligible population (VEP), which consists of all U.S.
+citizens over the age of 18 who are eligible to vote in the upcoming election.
+This includes both registered voters and likely voters, ensuring inclusivity in the survey while focusing more on likely voters, as they have the highest chance of participating in the election.
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+1.2. Sample Size Calculation
+Given a total population of over 250 million U.S.
+adults and assuming a margin of error of ±2% at a 95% confidence level, we aim for a sample size of n = 2,500 respondents.
+This will provide adequate precision in estimating key outcomes, such as the overall percentage of votes for the candidates.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+Budget considerations: We estimate that survey responses can be collected at a cost of approximately $25 per respondent (including recruitment, incentives, and data validation costs), which gives a budget allocation of approximately $62,500 for data collection.
+1.3. Sampling Frame
+We will use multi-stage stratified sampling to ensure a representative and balanced sample that reflects the demographics of the U.S. electorate.
+
+Stage 1: Stratification by state (to ensure geographic representativeness), including oversampling battleground states like Pennsylvania, Wisconsin, Michigan, and Florida.
+Stage 2: Stratification within each state by demographic variables such as age, gender, race/ethnicity, and education.
+These are key factors that influence voting behavior and must be accurately represented.
+Stage 3: Random sampling of individuals within each stratum using lists from voter registration databases and online panel recruitment for hard-to-reach groups.
+This method ensures that all relevant subpopulations are sufficiently represented, and weighting adjustments can be made later if necessary to correct for any sampling imbalances.
+
+2. Recruitment of Respondents
+2.1. Recruitment Channels
+We will employ a mixed-method recruitment strategy:
+
+Online Panels: Leveraging existing voter panels such as YouGov and Ipsos, we will recruit participants with a focus on capturing a diverse range of demographics.
+Random-Digit Dialing (RDD): To reach populations not heavily engaged online, especially older voters, we will conduct telephone interviews using RDD (both landline and cell phones).
+Incentives: Offering a small monetary incentive (e.g., $10 per completed survey) will encourage participation and increase response rates.
+2.2. Survey Mode
+A mixed-mode survey approach will be used to reach a broad range of participants:
+
+Online surveys for digitally engaged respondents.
+Telephone surveys (both live interviews and Interactive Voice Response (IVR)) for those less likely to respond online, particularly among older voters and rural populations.
+3. Survey Instrument Design
+3.1. Survey Questions
+The survey will consist of a mix of closed-ended and open-ended questions, focusing on:
+
+Voter preferences: Asking respondents whom they plan to vote for, using both traditional candidate name options and ranked-choice format (if applicable to state-specific elections).
+Voter likelihood: A question to assess how likely the respondent is to vote, with responses on a Likert scale (e.g., "How likely are you to vote in the upcoming election?").
+Demographics: Standard questions on age, gender, race/ethnicity, education, income, and region to ensure representativeness.
+Issue-based questions: Gathering respondents' positions on key policy issues (e.g., economy, healthcare, climate change), as this may correlate with candidate preference.
+3.2. Survey Length
+The survey will be concise to avoid fatigue, aiming for a completion time of no more than 10-15 minutes.
+
+4. Data Validation and Quality Control
+4.1. Response Validation
+Deduplication: Using respondent IP addresses, phone numbers, and other unique identifiers, we will ensure no duplicate responses are included.
+Screening questions: Filtering out respondents who are ineligible to vote or not part of the target population (e.g., non-U.S. citizens, individuals under 18).
+Attention checks: Including questions that help detect respondents who are not paying attention (e.g., “Select option 3 to continue”).
+4.2. Weighting and Post-Stratification
+We will employ post-stratification weighting to adjust the data to match known population benchmarks, using factors such as:
+
+Age
+Gender
+Race/Ethnicity
+Education level
+Geographic location (state)
+Voting history (using voter file data)
+This will help correct any sampling biases and ensure that our sample is demographically aligned with the general U.S. electorate.
+
+Weighting will be performed using R’s survey package, which allows us to adjust for complex survey designs.
+
+4.3. Polling Aggregation
+To improve the robustness of our forecast, we will aggregate our poll results with other reliable polls. Poll aggregation will use a Bayesian model, weighting polls based on:
+
+Sample size
+Methodology (online, IVR, live phone interviews)
+Pollster accuracy ratings (using metrics like FiveThirtyEight’s pollster ratings)
+The model will also account for the recency of the polls and house effects (consistent biases shown by particular pollsters).
+
+5. Data Analysis and Forecasting
+5.1. Predictive Model
+We will employ a multilevel regression and post-stratification (MRP) model to forecast election outcomes.
+This model will leverage both individual-level survey data and state-level covariates (e.g., past voting behavior, demographic composition) to make predictions for each state.
 
 ```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+# # Loading necessary libraries
+# library(survey)
+# library(dplyr)
+
+# # Assuming 'poll_data' is a dataframe with survey responses
+
+# # Define survey design
+# design <- svydesign(ids = ~1, strata = ~state, weights = ~weight, data = poll_data)
+
+# # Post-stratification weighting
+# # Assume we have a dataframe 'pop_data' with known population distributions
+# post_strat_weights <- postStratify(design, ~age + gender + race, pop_data)
+
+# # Run the weighted analysis (e.g., estimate support for candidate)
+# result <- svymean(~candidate_support, post_strat_weights)
+# print(result)
+
+# # Aggregate polls using Bayesian updating
+# # Assume 'aggregated_polls' is a dataframe with multiple polls' data
+# library(rstanarm)
+
+# # Fit Bayesian model
+# fit <- stan_glm(candidate_support ~ poll_method + recency, data = aggregated_polls, family = binomial())
+# summary(fit)
+
+# # Predict election outcome
+# predictions <- posterior_predict(fit, newdata = new_poll_data)
+# print(predictions)
 
@@
 
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
+## Budget Allocation
+Category	Cost Estimate
+Survey recruitment & incentives	$62,500
+Data validation & quality control	$10,000
+Poll aggregation & modeling software	$7,500
+Statistical consulting & data analysis	$10,000
+Miscellaneous (overheads, report writing)	$10,000
+Total	$100,000
 

### 2024-10-20 scripts/06-model_data.R
first attempt at a complete model with all the write-ups

@@ -15,21 +15,50 @@ library(rstanarm)
 #### Read data ####
 analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
-glm_model <- 
-  stan_glm(
-    formula = candidate_binary ~ numeric_grade + pollscore + transparency_score + sample_size + party + state + num_supporters,
-    data = analysis_data,
-    family = binomial(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 304
-  )
+harris_data <- subset(analysis_data, candidate_name == "Kamala Harris")
+
+#### Model data ####
+# Convert categorical variables to factors for Kamala Harris
+harris_data$state <- as.factor(harris_data$state)
+harris_data$methodology <- as.factor(harris_data$methodology)
+
+# Build MLR model for Kamala Harris
+mlr_harris_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, 
+                       data = harris_data)
 
 #### Save model ####
-saveRDS(
-  glm_model,
-  file = "models/glm_model.rds"
-)
+# Save the model to an .rds file for future use
+saveRDS(mlr_harris_model, file = "models/mlr_harris_model.rds")
+
+
+# Filter for Donald Trump data only
+trump_data <- subset(analysis_data, candidate_name == "Donald Trump")
+
+#### Model data ####
+# Convert categorical variables to factors for Donald Trump
+trump_data$state <- as.factor(trump_data$state)
+trump_data$methodology <- as.factor(trump_data$methodology)
+
+# Build MLR model for Donald Trump
+mlr_trump_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, 
+                      data = trump_data)
+
+#### Save model ####
+# Save the model to an .rds file for future use
+saveRDS(mlr_trump_model, file = "models/mlr_trump_model.rds")
+
+
+mlr_harris_model_refined <- lm(pct ~ numeric_grade + pollscore + sample_size + state, data = harris_data)
+
+saveRDS(mlr_harris_model_refined, file = "models/mlr_harris_model_refined.rds")
+
+
[…]

### 2024-10-20 paper/paper.qmd

--- 2024-10-20 paper/paper.qmd
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper presents a statistical model for forecasting the outcome of the 2024 U.S.
+Presidential Election.
+Using polling data from various sources, we develop multiple linear regression models to predict the percentage of support for the main candidates, Kamala Harris and Donald Trump, in different states.
+We aggregate these predictions to simulate the Electoral College vote and estimate the likelihood of either candidate winning.
+Our results suggest that neither candidate currently secures the required 270 electoral votes to win the presidency."
 format: pdf
@@
 
-Overview paragraph
+Forecasting elections has been a challenging task for political scientists and statisticians, especially given the complexity and volatility of electoral dynamics in the U.S.
+The 2024 U.S.
+Presidential Election is no exception, with polling data playing a critical role in shaping public and expert expectations.
+Pollsters use different methodologies and sample populations, which can significantly affect the predictions.
 
-Estimand paragraph
+This paper aims to forecast the 2024 U.S.
+Presidential Election using multiple linear regression models based on polling data.
+We use various predictors such as sample size, pollster ratings, and state-level data to predict support for Kamala Harris and Donald Trump across the United States.
+We then aggregate these predictions to simulate the Electoral College outcome.
 
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The rest of this paper is structured as follows: Section 2 describes the data used for this analysis.
+Section 3 outlines the models developed for each candidate and their corresponding results.
+Section 4 discusses the aggregated Electoral College predictions.
+Finally, Section 5 provides conclusions and suggestions for future research.
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The dataset used in this analysis contains poll data from various sources, including information about polling organizations, sample sizes, methodologies, and state-level data.
+It includes several key variables, such as:
 
-Overview text
-
-## Measurement
-
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
-## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Pollster rating: A score that represents the reliability of the polling organization.
+Sample size: The number of respondents in each poll.
+Support percentage: The percentage of respondents supporting each candidate.
+State: The U.S. state where the poll was conducted or, in some cases, national polling data.
+The data has been processed and cleaned to ensure that all variables are correctly aligned for regression analysis.
+The variables were transformed into numeric forms where necessary (e.g., percentage of support), and missing data was handled appropriately.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Load necessary libraries
@@
 
-Simple Linear Regression for Kamala Harris and Donald Trump
+## Outcome variables
 
-Kamala Harris
-We will perform a simple linear regression using pct (poll percentage) as the dependent variable and sample_size as the independent variable.
+Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
+Figure 1 below shows the distribution of support for both candidates across the polls included in the dataset.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
+# Summary statistics for percentage support (pct)
+summary(data$pct)
+
+# Plotting percentage support for different candidates
+ggplot(data, aes(x = candidate_name, y = pct)) +
+  geom_boxplot() +
+  labs(title = "Support Distribution Across Candidates", x = "Candidate", y = "Percentage")
+
+```
+
+Multiple Linear Regression (MLR) for Kamala Harris and Donald Trump
+
+Electoral College Votes Prediction
+In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
+
+## Predictor variables
+
+Add graphs, tables and text.
+
+Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+# Modeling Support for the Candidates
+
+## Kamala Harris
+We begin by modeling the percentage of support for Kamala Harris using a linear regression model.
+The predictors include the sample size, pollster ratings (e.g., pollscore and transparency score), and state.
+This model aims to quantify how these variables influence her support across different polls.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Load necessary library
@@
 
-# Summary of the model
-summary(model)
+```
 
+The results of the linear regression for Kamala Harris show that sample size has a statistically insignificant effect on her support.
+This suggests that factors other than sample size, such as pollster methodology or regional biases, may play a more significant role in determining the level of support she receives.
+
+Figure 2 illustrates the relationship between the sample size and support percentage for Kamala Harris.
+
+```{r}
 # Plot the relationship for Kamala Harris
@@
 
-Donald Trump
-We will repeat the same process for Donald Trump.
+## Donald Trump
+A similar linear regression model was applied to Donald Trump's data. The predictors remain the same, and the goal is to determine the factors that drive support for him.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Load necessary library
@@
 
-# Summary of the model
-summary(model)
+```
 
+The results indicate that sample size has a weak but statistically significant negative effect on support for Trump.
+This suggests that larger polls tend to show slightly lower support for Trump, although the effect size is small.
+
+Figure 3 shows the relationship between the sample size and support percentage for Donald Trump.
+```{r}
 # Plot the relationship for Kamala Harris
@@
 
-Multiple Linear Regression (MLR) for Kamala Harris and Donald Trump
+## Multiple Linear Regression Models
+To better capture the complexity of voter support, we constructed multiple linear regression (MLR) models for both Kamala Harris and Donald Trump.
+These models incorporate several predictors, including pollster rating (represented by pollscore), transparency score, sample size, and state-level data.
+By accounting for these factors, the MLR models allow us to control for more variables that influence voter support and provide more accurate predictions.
 
-Kamala Harris
-In this step, we will build a multiple linear regression (MLR) model for Kamala Harris using additional predictors like numeric_grade, pollscore, transparency_score, and state.
+### Kamala Harris Model
+The MLR model for Kamala Harris takes into account multiple factors that may impact her support across different polls.
+This includes not just the sample size of the poll, but also how pollster reliability (numeric_grade), transparency of poll data (transparency_score), and state-level polling contribute to predicting voter preferences.
 
 ```{r}
-# Convert categorical variables to factors for Kamala Harris
-harris_data$state <- as.factor(harris_data$state)
-harris_data$methodology <- as.factor(harris_data$methodology)
-
-# Build MLR model for Kamala Harris
-mlr_harris_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, data = harris_data)
-
-# Summary of the MLR model for Harris
-summary(mlr_harris_model)
+#| echo: false
+#| message: false
+#| warning: false
+mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
 
@@
 
+The results of the MLR for Kamala Harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
+These results suggest that voter preferences for Harris vary widely depending on the state and the credibility of the pollster.
+
+The model diagnostics (such as residual plots, QQ plots) were evaluated to ensure the assumptions of linear regression hold.
+Figure X shows diagnostic plots, which indicate that the model performs reasonably well in terms of residual behavior and normality.
 Similarly, we build an MLR model for Donald Trump.
+
+### Donald Trump Model
+Similarly, we constructed an MLR model for Donald Trump using the same set of predictors to assess the factors that influence his support across the country.
 ```{r}
-# Convert categorical variables to factors for Trump
-trump_data$state <- as.factor(trump_data$state)
-trump_data$methodology <- as.factor(trump_data$methodology)
-
-# Build MLR model for Donald Trump
-mlr_trump_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + state + methodology, data = trump_data)
-
-# Summary of the MLR model for Trump
-summary(mlr_trump_model)
+#| echo: false
+#| message: false
+#| warning: false
+# Load the saved MLR model for Donald Trump
+mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
 
 # Diagnostic plots for the MLR model
-par(mfrow = c(2, 2))
-plot(mlr_trump_model)
-par(mfrow = c(1, 1))  # Reset plot layout
+par(mfrow = c(2, 2))  # Set up 2x2 layout for multiple plots
+plot(mlr_trump_model)  # Generates diagnostic plots: residuals, QQ plot, etc.
+par(mfrow = c(1, 1))  # Reset the plot layout to normal
+
 ```
 
+The results for Donald Trump reveal that state-level variations and pollster transparency play a significant role in explaining his level of support.
+The model for Trump also passes diagnostic checks, as shown in Figure Y, indicating that the assumptions of linearity, independence, and homoscedasticity are reasonably met.
 Check Multicollinearity using Variance Inflation Factor (VIF)
 
+## Multicollinearity Check Using Variance Inflation Factor (VIF)
+To ensure that the predictors used in both models do not exhibit multicollinearity, we checked the Variance Inflation Factor (VIF) for each predictor.
+High VIF values indicate multicollinearity, which can affect the stability and reliability of the model coefficients.
+
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Load the car package for VIF
@@
 # Refine the model by removing less significant predictors (e.g., methodology and transparency_score)
-mlr_harris_model_refined <- lm(pct ~ numeric_grade + pollscore + sample_size + state, data = harris_data)
+mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
 
 # Summary of the refined Harris model
-summary(mlr_harris_model_refined)
 
@@
 
-Stepwise Model Selection for Optimization
+Based on the VIF results, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multicollinearity.
+This improves the model's accuracy and interpretability.
+
+## Stepwise Model Selection
+To further optimize the MLR models, we performed stepwise model selection, which systematically adds or removes predictors to minimize the Akaike Information Criterion (AIC) and improve model fit.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Perform stepwise selection to optimize the Harris model
 step_model <- step(mlr_harris_model_refined, direction = "both")
-
-# Summary of the stepwise model
-summary(step_model)
 
@@
 
-Final Models and Predictions for Kamala Harris and Donald Trump
+The stepwise model selection improved the model by retaining the most significant predictors and eliminating those with little explanatory power.
+
+## Final Predictions for Electoral College Votes
+Using the final models for both candidates, we predicted the percentage of support in each state and aggregated the results to simulate the Electoral College outcome.
+The predicted percentage of support for each candidate is used to determine the likely winner in each state.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Final models for Harris and Trump
-mlr_harris_model_final <- lm(pct ~ pollscore + log(sample_size) + state, data = harris_data)
-mlr_trump_model_final <- lm(pct ~ pollscore + log(sample_size) + state, data = trump_data)
+mlr_harris_model_final <- readRDS(here::here(("models/mlr_harris_model_final.rds")))
+mlr_trump_model_final <- readRDS(here::here(("models/mlr_trump_model_final.rds")))
 
@@
 
-Electoral College Votes Prediction
-In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
+# Electoral College Prediction
+To forecast the winner of the 2024 election, we aggregate the predicted percentages of support from our models for each state and calculate the Electoral College votes.
+The candidate with 270 or more Electoral College votes is predicted to win the election.
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Define electoral votes for each state (and Washington, D.C.)
@@
 
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-# #| echo: false
-# #| eval: true
-# #| warning: false
-# #| message: false
-
-# library(rstanarm)
-
-# first_model <-
-#   readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-# #| echo: false
-# #| eval: true
-# #| label: tbl-modelresults
-# #| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-# #| warning: false
-
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
-```
+According to our models, neither candidate secures the 270 Electoral College votes needed to win.
+Kamala Harris is projected to receive 216 electoral votes, while Donald Trump is predicted to win 147 electoral votes.
+However, due to the lack of 270 electoral votes for either candidate, the election remains highly competitive, and further developments may shift the balance.
 
@@
 
-## First discussion point {#sec-first-point}
+## Key Findings
+The multiple linear regression (MLR) models developed in this paper provide valuable insights into the factors that influence voter support for Kamala Harris and Donald Trump.
+Through the analysis of polling data, several key predictors were identified as significant, including state, pollster rating (pollscore), and transparency score.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Kamala Harris
+The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
+The model shows that voter preferences vary significantly across different states, with certain states (e.g., California and Maryland) showing higher support levels, while others (e.g., Indiana and Missouri) show lower support.
+Pollster reliability, captured through numeric grades and transparency scores, also plays an important role, suggesting that voters may respond differently based on the credibility of the pollster.
 
-## Second discussion point
+Donald Trump
+Similarly, the MLR model for Donald Trump reveals that state-level variations are critical to understanding his level of support.
+The model indicates that Trump’s support is more stable across certain states, but there are also notable outliers where his support fluctuates.
+Pollster characteristics, such as pollscore and transparency score, significantly impact Trump’s predicted support, reflecting the importance of poll quality in predicting election outcomes.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Electoral College Forecast
+Aggregating state-level predictions into Electoral College votes demonstrates the potential competitiveness of the 2024 U.S.
+Presidential Election.
+According to the predictions, neither Kamala Harris nor Donald Trump currently secures the necessary 270 electoral votes to win.
+Harris is projected to receive 216 electoral votes, while Trump is predicted to receive 147 electoral votes.
+This forecast suggests that the election remains highly uncertain, with several key battleground states likely determining the final outcome.
 
-## Third discussion point
+## Model Strengths
+One of the main strengths of this analysis is the incorporation of multiple predictors that allow us to account for various factors influencing voter support.
+By considering state-level data, pollster ratings, and transparency scores, the models provide a more nuanced prediction of voter behavior compared to models that rely solely on national-level polling.
 
-## Weaknesses and next steps
+The use of stepwise model selection and multicollinearity checks further enhanced the robustness of the models by optimizing the choice of predictors and ensuring that the models do not suffer from unstable estimates caused by correlated predictors.
 
-Weaknesses and next steps should also be included.
+Additionally, the aggregation of state-level predictions into Electoral College outcomes presents a more realistic forecast of the election, as the U.S.
+Presidential election is ultimately decided by electoral votes, not popular votes.
+
+## Limitations
+Despite the strengths of the models, there are several limitations that should be addressed:
+
+Poll Reliability and Sampling Bias: The accuracy of the predictions depends heavily on the quality of the polling data.
+Although the model accounts for pollster ratings, polling methodologies can still introduce biases, particularly in states with limited polling data.
+Sampling errors and non-response biases could skew the results, especially in smaller states or regions with inconsistent polling coverage.
+
+Static Prediction: The model provides a static snapshot of voter support based on current polling data, which may not capture the dynamic nature of voter preferences over time.
+As election day approaches, voter preferences may shift due to campaign events, debates, or other external factors.
+Without time-series data, the model may fail to account for these trends.
+
+Unaccounted Variables: Although the model includes important predictors like state, pollscore, and transparency score, other potentially influential factors, such as economic conditions, campaign spending, and voter turnout, are not included in the analysis.
+These unaccounted variables may introduce inaccuracies in the final predictions.
+## Future Research Directions
+Future work could improve upon this analysis by addressing some of the limitations mentioned above.
+For instance, incorporating time-series data could allow the model to capture how voter preferences evolve in response to external factors such as campaign events, economic developments, and political endorsements.
+A dynamic forecasting model that updates predictions as new polls are released would provide more timely and accurate forecasts.
+
+Moreover, integrating other influential variables, such as economic indicators (e.g., unemployment rates, inflation), voter turnout models, and campaign spending data, could enhance the predictive power of the model.
+Including demographic data (e.g., age, education, income) could also improve the granularity of predictions, especially in battleground states where demographic shifts are critical to election outcomes.
+
+Finally, expanding the model to account for ranked-choice voting in certain states could provide a more accurate forecast in scenarios where third-party candidates or run-off elections play a significant role.
 

### 2024-10-20 scripts/03-clean_data.R
deleted install.packages code

@@ -7,10 +7,6 @@
 # Pre-requisites: Requires 'president_polls.csv' file in 'data/01-raw_data/'
 # Any other information needed? This script builds an MLR model for Harris vs Trump using polling data.
 
-#### Install necessary libraries ####
-install.packages("janitor")
-install.packages("lubridate")
-
 #### Workspace setup ####
 # Load necessary libraries
 library(dplyr)

### 2024-10-20 scripts/05-exploratory_data_analysis.R
deleted install packages code

@@ -6,12 +6,6 @@
 # License: MIT
 # Pre-requisites: Requires cleaned 'analysis_data.csv' with filtered data for Harris and Trump
 
-#### Install necessary libraries ####
-install.packages("ggplot2")
-install.packages("dplyr")
-install.packages("naniar")
-install.packages("tidyverse")
-
 #### Workspace setup ####
 # Load necessary libraries
 library(dplyr)

### 2024-10-20 paper/paper.qmd
editted introduction working on data-sec

@@
   - Tommy Fu
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git."
 date: today
@@
 number-sections: true
+toc: true
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 
-# Example of installing a package after setting the mirror
-install.packages("tidyverse")
-
-install.packages("here")
 library(tidyverse)
-# Install janitor package if not installed
-install.packages("janitor")
-
-# Load the janitor package
+library(ggplot2)
 library(janitor)
-
-install.packages("caret")
-install.packages("ROCR")
-install.packages("rstanarm")
 library(rstanarm)
-install.packages("modelsummary")
 library(modelsummary)
-
 library(here)
-
-```
-
-# Introduction
-
-Forecasting elections has been a challenging task for political scientists and statisticians, especially given the complexity and volatility of electoral dynamics in the U.S.
-The 2024 U.S.
-Presidential Election is no exception, with polling data playing a critical role in shaping public and expert expectations.
-Pollsters use different methodologies and sample populations, which can significantly affect the predictions.
-
-This paper aims to forecast the 2024 U.S.
-Presidential Election using multiple linear regression models based on polling data.
-We use various predictors such as sample size, pollster ratings, and state-level data to predict support for Kamala Harris and Donald Trump across the United States.
-We then aggregate these predictions to simulate the Electoral College outcome.
-
-The rest of this paper is structured as follows: Section 2 describes the data used for this analysis.
-Section 3 outlines the models developed for each candidate and their corresponding results.
-Section 4 discusses the aggregated Electoral College predictions.
-Finally, Section 5 provides conclusions and suggestions for future research.
-
-# Data {#sec-data}
-
-## Overview
-
-The dataset used in this analysis contains poll data from various sources, including information about polling organizations, sample sizes, methodologies, and state-level data.
-It includes several key variables, such as:
-
-Pollster rating: A score that represents the reliability of the polling organization.
-Sample size: The number of respondents in each poll.
-Support percentage: The percentage of respondents supporting each candidate.
-State: The U.S. state where the poll was conducted or, in some cases, national polling data.
-The data has been processed and cleaned to ensure that all variables are correctly aligned for regression analysis.
-The variables were transformed into numeric forms where necessary (e.g., percentage of support), and missing data was handled appropriately.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Load necessary libraries
 library(dplyr)
-library(tidyr)
+library(knitr)
 
@@
 
-## Outcome variables
+# Introduction
 
-Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
-Figure 1 below shows the distribution of support for both candidates across the polls included in the dataset.
+Forecasting elections has long been one of the most challenging tasks for political scientists, statisticians, and analysts alike.
+The complexity of predicting voter behavior in large, diverse electorates like that of the United States stems from numerous factors, including ever-shifting public opinion, rapidly changing political climates, and varying levels of voter engagement.
+The 2024 U.S.
+Presidential Election is no exception.
+With polarized voter bases, unpredictable external factors, and the growing influence of non-traditional media, the race remains highly uncertain, making reliable forecasting models more important—and difficult—than ever.
+
+Central to any electoral forecast is the use of polling data.
+Polls provide snapshots of voter intentions at specific points in time, shaping both public perceptions and expert expectations.
+Yet, polling comes with its own set of challenges.
+Pollsters employ different methodologies—ranging from online surveys to phone interviews—each with its own inherent biases.
+Sample populations may vary in size and representativeness, potentially skewing results in favor of certain demographic groups.
+Moreover, polling accuracy can be further complicated by factors such as non-response bias, the changing nature of the electorate, and the rise of unconventional voting patterns, particularly among younger and minority voters.
+
+In this paper, we attempt to forecast the outcome of the 2024 U.S.
+Presidential Election by developing statistical models based on polling data from a variety of sources.
+By leveraging multiple linear regression models, we aim to predict the percentage of support for the main candidates—Kamala Harris and Donald Trump—across different states.
+Our analysis incorporates key variables such as pollster reliability, sample size, and state-level demographics, all of which contribute to shaping the electoral landscape.
+Ultimately, we aggregate these state-level predictions to simulate the outcome of the Electoral College, providing insights into the likelihood of either candidate securing the 270 votes required to win the presidency.
+
+The rest of this paper is structured as follows: @sec-data discusses the data used for this analysis, including key variables and sources.
+@sec-model outlines the models developed for each candidate and presents the corresponding results.
+@sec-predict discusses the aggregated Electoral College predictions based on the model outputs. Finally, @sec-discuss provides conclusions and offers suggestions for future research.
+
+# Data {#sec-data}
+
+## Overview
+
+The dataset used in this analysis draws from a variety of polling sources, providing comprehensive information about polling organizations, sample sizes, methodologies, and state-level data.
+Several key variables are crucial to the analysis:
+
+- Pollster rating: A numerical score reflecting the reliability and historical accuracy of each polling organization.
+- Sample size: The number of respondents included in each poll, which influences the poll's margin of error.
+- Support percentage: The percentage of respondents expressing support for each candidate.
+- State: The U.S. state where the poll was conducted, or in some cases, national-level polling data.
+
+To ensure accuracy and consistency, the data has been meticulously processed and cleaned.
+This involved aligning all variables for compatibility with regression analysis, converting relevant fields (e.g.
+support percentage) into numeric formats, and appropriately handling missing data to minimize potential biases.
+These reprocessing steps ensure the dataset is tidy and ready for model development.
+
+@tbl-overview_data shows a subset of the dataset.
 
@@
 #| warning: false
-# Summary statistics for percentage support (pct)
-summary(data$pct)
+#| label: tbl-overview_data
+# Select key variables for the table
+data_overview <- data %>%
+  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
 
-# Plotting percentage support for different candidates
-ggplot(data, aes(x = candidate_name, y = pct)) +
+# Show the first 5 rows of the dataset in a table
+data_overview %>%
+  head(5) %>%
+  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
+
+```
+
+## Outcome variables
+
+Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
+@fig-support below shows the distribution of support for both candidates across the polls included in the dataset.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-support
+#| fig-cap: Distribution of Support between Both Candidates
+
+# Create a boxplot showing the distribution of support for each candidate
+ggplot(data, aes(x = candidate_name, y = pct, fill = candidate_name)) +
   geom_boxplot() +
-  labs(title = "Support Distribution Across Candidates", x = "Candidate", y = "Percentage")
+  labs(title = "Distribution of Support for Kamala Harris and Donald Trump",
+       x = "Candidate",
+       y = "Percentage of Support") +
+  theme_minimal() +
+  theme(legend.position = "none")
 
@@
 
-Electoral College Votes Prediction
-In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
+Electoral College Votes Prediction In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
 
@@
 
-# Modeling Support for the Candidates
+# Modeling Support for the Candidates {#sec-model}
 
 ## Kamala Harris
+
 We begin by modeling the percentage of support for Kamala Harris using a linear regression model.
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
 # Plot the relationship for Kamala Harris
@@
 ## Donald Trump
-A similar linear regression model was applied to Donald Trump's data. The predictors remain the same, and the goal is to determine the factors that drive support for him.
+
+A similar linear regression model was applied to Donald Trump's data.
+The predictors remain the same, and the goal is to determine the factors that drive support for him.
 
@@
 
-The results indicate that sample size has a weak but statistically significant negative effect on support for Trump.
+The results in @figure-2 indicate that sample size has a weak but statistically significant negative effect on support for Trump.
 This suggests that larger polls tend to show slightly lower support for Trump, although the effect size is small.
@@
 Figure 3 shows the relationship between the sample size and support percentage for Donald Trump.
+
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: figure-2
 # Plot the relationship for Kamala Harris
@@
   geom_smooth(method = "lm", col = "blue") +
-  labs(title = "Linear Regression of Percentage vs Sample Size for Donald",
+  labs(title = "Linear Regression of Percentage vs Sample Size for Donald Trump",
        x = "Sample Size",
@@
 ## Multiple Linear Regression Models
+
 To better capture the complexity of voter support, we constructed multiple linear regression (MLR) models for both Kamala Harris and Donald Trump.
@@
 ### Kamala Harris Model
+
 The MLR model for Kamala Harris takes into account multiple factors that may impact her support across different polls.
@@
 ### Donald Trump Model
+
 Similarly, we constructed an MLR model for Donald Trump using the same set of predictors to assess the factors that influence his support across the country.
+
 ```{r}
@@
 ## Multicollinearity Check Using Variance Inflation Factor (VIF)
+
 To ensure that the predictors used in both models do not exhibit multicollinearity, we checked the Variance Inflation Factor (VIF) for each predictor.
@@
 ## Stepwise Model Selection
+
 To further optimize the MLR models, we performed stepwise model selection, which systematically adds or removes predictors to minimize the Akaike Information Criterion (AIC) and improve model fit.
@@
 ## Final Predictions for Electoral College Votes
+
 Using the final models for both candidates, we predicted the percentage of support in each state and aggregated the results to simulate the Electoral College outcome.
@@
 
-# Electoral College Prediction
+# Electoral College Prediction {#sec-predict}
+
 To forecast the winner of the 2024 election, we aggregate the predicted percentages of support from our models for each state and calculate the Electoral College votes.
@@
 
-# Discussion
+# Discussion {#sec-discuss}
 
 ## Key Findings
+
 The multiple linear regression (MLR) models developed in this paper provide valuable insights into the factors that influence voter support for Kamala Harris and Donald Trump.
@@
 
-Kamala Harris
-The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
+Kamala Harris The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
 The model shows that voter preferences vary significantly across different states, with certain states (e.g., California and Maryland) showing higher support levels, while others (e.g., Indiana and Missouri) show lower support.
@@
 
-Donald Trump
-Similarly, the MLR model for Donald Trump reveals that state-level variations are critical to understanding his level of support.
+Donald Trump Similarly, the MLR model for Donald Trump reveals that state-level variations are critical to understanding his level of support.
 The model indicates that Trump’s support is more stable across certain states, but there are also notable outliers where his support fluctuates.
@@
 
-Electoral College Forecast
-Aggregating state-level predictions into Electoral College votes demonstrates the potential competitiveness of the 2024 U.S.
+Electoral College Forecast Aggregating state-level predictions into Electoral College votes demonstrates the potential competitiveness of the 2024 U.S.
 Presidential Election.
@@
 ## Model Strengths
+
 One of the main strengths of this analysis is the incorporation of multiple predictors that allow us to account for various factors influencing voter support.
@@
 ## Limitations
+
 Despite the strengths of the models, there are several limitations that should be addressed:
@@
 These unaccounted variables may introduce inaccuracies in the final predictions.
-## Future Research Directions
-Future work could improve upon this analysis by addressing some of the limitations mentioned above.
+## Future Research Directions Future work could improve upon this analysis by addressing some of the limitations mentioned above.
 For instance, incorporating time-series data could allow the model to capture how voter preferences evolve in response to external factors such as campaign events, economic developments, and political endorsements.
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
+This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+This shows...
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+It shows...
+This suggests...
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+It shows...
+This suggests...
 
@@
 
-# References
+#References

### 2024-10-21 paper/paper.qmd
Editted Intro and data-sec

@@
 
+## Measurement and Limitations
+
+The primary limitations of this analysis stem from the quality and variability of polling data, which can introduce biases through factors like sample size and pollster methodology.
+Additionally, the model provides a static snapshot of voter preferences, without accounting for the dynamic nature of elections For instance, shifts in public opinion over time or the impact of campaign events.
+The exclusion of key external factors, such as economic conditions or voter turnout, also limits the model's ability to fully capture the complexities of election outcomes.
+Furthermore, the simplified Electoral College simulation assumes that polling data will accurately predict state-level results, which may not always be the case, especially in battleground states with volatile voter behavior.
+
 ## Outcome variables
+Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
+This variable is central to our analysis because it directly captures voter preference, which is the most relevant metric for forecasting election outcomes.
+By modeling the percentage of support, we can quantify how various factors—such as pollster quality, sample size, and state-level dynamics—influence the candidates' standings in the polls.
+This outcome variable also allows us to simulate the Electoral College results by aggregating predicted support across different states, which is crucial for determining the likelihood of either candidate winning the presidency.
 
-Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
 @fig-support below shows the distribution of support for both candidates across the polls included in the dataset.
@@
 #| warning: false
+#| fig.pos: "H"
 #| label: fig-support
@@
   theme(legend.position = "none")
-
 ```
-
-Multiple Linear Regression (MLR) for Kamala Harris and Donald Trump
-
-Electoral College Votes Prediction In this step, we aggregate the predictions for each state and determine the winner based on the electoral votes.
 
@@
 Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+## Cleaning Process and Analysis
+
+The data cleaning process uses R [@citeR] as well as other packages such as tidyverse [@tidyverse], dplyr [@dplyr], janitor [@janitor] and lubridate [@lubridate].
+Begin by standardizing column name to ensure consistency and avoid errors during analysis.
+The dataset was filtered to focus solely on Kamala Harris and Donald Trump, with polls below a numeric grade of 2.7 excluded to maintain high-quality data.
+Missing state information was categorized as “National” polls, and the poll end dates were converted to a date format to include only polls conducted after the candidates officially declared their candidacies.
+This ensures the analysis reflects relevant, up-to-date voter preferences.
+Additionally, transformations such as calculating the number of supporters from the percentage support and encoding a binary variable for the candidates (Harris = 1, Trump = 0) were performed to enhance usability in the regression models.
+
+Some variables were excluded for simplicity and relevance.
+For example, **population_full** and **answer** were dropped as they provided redundant or unnecessary information for this analysis.
+However, key variables like **pollster**, **pollscore**, **numeric grade**, and **sample size** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
+These steps ensured the cleaned dataset was ready for analysis, with a focus on the most important predictors of voter support while minimizing extraneous information.
 
@@
 
-The results in @figure-2 indicate that sample size has a weak but statistically significant negative effect on support for Trump.
+The results in indicate that sample size has a weak but statistically significant negative effect on support for Trump.
 This suggests that larger polls tend to show slightly lower support for Trump, although the effect size is small.
@@
 #| warning: false
-#| label: figure-2
+
 # Plot the relationship for Kamala Harris

### 2024-10-21 paper/paper.qmd
some minor changed in cleaning section

@@
 Some variables were excluded for simplicity and relevance.
-For example, **population_full** and **answer** were dropped as they provided redundant or unnecessary information for this analysis.
-However, key variables like **pollster**, **pollscore**, **numeric grade**, and **sample size** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
+For example, **"population_full"** and **"answer"** were dropped as they provided redundant or unnecessary information for this analysis.
+However, key variables like **"pollster"**, **"pollscore"**, **"numeric grade"**, and **"sample size"** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
 These steps ensured the cleaned dataset was ready for analysis, with a focus on the most important predictors of voter support while minimizing extraneous information.

### 2024-10-22 paper/paper.qmd

@@
 
-Add graphs, tables and text.
+In our analysis, we utilized several key predictor variables to model voter support for Kamala Harris and Donald Trump in the 2024 U.S.
+Presidential Election.
+These predictor variables were selected based on their expected influence on polling results and their ability to account for variations in polling accuracy and voter behavior across different states.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+One of the key predictors is numeric grade, which represents a composite measure of poll quality.
+This variable accounts for various aspects of a poll’s reliability, such as the transparency of its methodology and the pollster’s historical accuracy.
+Pollscore, another important predictor, provides a similar measure of poll reliability but focuses more on the specific performance of the polling organization in recent elections.
+
+Transparency score measures how openly polling organizations disclose their methodology and data, which can significantly impact the trustworthiness of the results.
+Polls with higher transparency scores are given more weight in the model, as they are generally considered more reliable.
+The sample size of each poll is also included, as larger sample sizes tend to reduce the margin of error and provide more precise estimates of voter support.
+
+Geographic differences in voter preferences are captured through the state variable, which accounts for regional variations in political culture, demographics, and historical voting patterns.
+Lastly, methodology refers to the techniques and approaches used by the polling organizations, such as whether the poll was conducted online, by phone, or through other means.
+
+All predictor variables—numeric grade, pollscore, transparency score, sample size, state, and methodology—are presented in @tbl-overview_data.
 

### 2024-10-22 scripts/03-clean_data.R
Added methodology to Appendix

@@ -14,12 +14,12 @@ library(tidyverse)
 library(janitor)
 library(lubridate)
 
-#### Data Import and Cleaning ####
+########################### Data Import and Cleaning########################### 
 # Read the dataset and clean column names
 data <- read_csv("data/01-raw_data/president_polls.csv") |>
   clean_names()
 
-#### Filter Data for Both Candidates ####
+####################### Filter Data for Both Candidates #######################
 # Filter data for Kamala Harris and Donald Trump based on high-quality polls
 filtered_data <- data |>
   filter(

### 2024-10-22 scripts/04-test_analysis_data.R

@@ -21,10 +21,11 @@ pollster_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 ########################### Tests for Pollster Data########################### 
 
 test_that("Pollster data has correct structure", {
-  expected_cols <- c("poll_id", "pollster_id", "pollster", "numeric_grade", "pollscore",
-                     "methodology", "transparency_score", "sample_size", "population",
-                     "population_full", "party", "answer", "pct", "state",
-                     "end_date", "num_supporters", "candidate_binary")
+  expected_cols <- c("poll_id", "pollster_id", "pollster", "numeric_grade",
+                     "pollscore", "methodology", "transparency_score", 
+                     "sample_size", "population", "population_full", "party", 
+                     "answer", "pct", "state", "candidate_name", "end_date", 
+                     "candidate_binary", "num_supporters")
   expect_equal(colnames(pollster_data), expected_cols)
 })
 
@@ -43,6 +44,7 @@ test_that("Data types are correct", {
   expect_true(is.character(pollster_data$answer))
   expect_true(is.numeric(pollster_data$pct))
   expect_true(is.character(pollster_data$state))
+  expect_true(is.character(pollster_data$candidate_name))
   expect_true(inherits(pollster_data$end_date, "Date"))
   expect_true(is.numeric(pollster_data$num_supporters))
   expect_true(is.numeric(pollster_data$candidate_binary))
@@ -71,13 +73,6 @@ test_that("Dates are within expected range", {
 test_that("No missing values in critical columns", {
   critical_cols <- c("poll_id", "pollster", "pct", "answer", "end_date")
   expect_true(all(!is.na(pollster_data[critical_cols])))
-})
-
-test_that("Percentages sum to 100 (or close to it) for each poll", {
-  poll_totals <- pollster_data %>%
-    group_by(poll_id) %>%
-    summarize(total_pct = sum(pct))
-  expect_true(all(poll_totals$total_pct >= 99 & poll_totals$total_pct <= 101))
 })
 
 test_that("Candidate binary values match answer names", {

### 2024-10-22 paper/paper.qmd

@@
   - Tommy Fu
-thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git."
+thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git"
 date: today
@@
 library(janitor)
-library(rstanarm)
 library(modelsummary)
@@
 library(knitr)
+library(car)
 
@@
 
+\newpage
 # Introduction
@@
 
-- Pollster rating: A numerical score reflecting the reliability and historical accuracy of each polling organization.
-- Sample size: The number of respondents included in each poll, which influences the poll's margin of error.
-- Support percentage: The percentage of respondents expressing support for each candidate.
-- State: The U.S. state where the poll was conducted, or in some cases, national-level polling data.
+- **Pollster rating:** A numerical score reflecting the reliability and historical accuracy of each polling organization.
+- **Sample size:** The number of respondents included in each poll, which influences the poll's margin of error.
+- **Support percentage:** The percentage of respondents expressing support for each candidate.
+- **State:** The U.S. state where the poll was conducted, or in some cases, national-level polling data.
 
@@
 
-@tbl-overview_data shows a subset of the dataset.
+@tbl-overview_data shows a sample of the dataset.
 
@@
 #| label: fig-support
-#| fig-cap: Distribution of Support between Both Candidates
+#| fig-cap: Distribution of Support for Kamala Harris and Donald Trump
 
@@
   geom_boxplot() +
-  labs(title = "Distribution of Support for Kamala Harris and Donald Trump",
-       x = "Candidate",
+  labs(x = "Candidate",
        y = "Percentage of Support") +
@@
 Some variables were excluded for simplicity and relevance.
-For example, **"population_full"** and **"answer"** were dropped as they provided redundant or unnecessary information for this analysis.
-However, key variables like **"pollster"**, **"pollscore"**, **"numeric grade"**, and **"sample size"** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
+For example, **'population_full'** and **'answer'** were dropped as they provided redundant or unnecessary information for this analysis.
+However, key variables like **'pollster'**, **'pollscore'**, **'numeric grade'**, and **'sample size'** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
 These steps ensured the cleaned dataset was ready for analysis, with a focus on the most important predictors of voter support while minimizing extraneous information.
@@
 model <- lm(pct ~ sample_size, data = harris_data)
-
 ```
@@
 
-Figure 2 illustrates the relationship between the sample size and support percentage for Kamala Harris.
+@fig-harris-samplesize-supportpct illustrates the relationship between the sample size and support percentage for Kamala Harris.
 
@@
 #| warning: false
+#| label: fig-harris-samplesize-supportpct
+#| fig-cap: Linear Regression of Percentage vs Sample Size for Kamala Harris
+
 # Plot the relationship for Kamala Harris
@@
   geom_smooth(method = "lm", col = "blue") +
-  labs(title = "Linear Regression of Percentage vs Sample Size for Kamala Harris",
-       x = "Sample Size",
+  labs(x = "Sample Size",
        y = "Percentage")
-
 ```
@@
 
-Figure 3 shows the relationship between the sample size and support percentage for Donald Trump.
+@fig-trump-samplesize-supportpct shows the relationship between the sample size and support percentage for Donald Trump.
 
@@
 #| warning: false
+#| label: fig-trump-samplesize-supportpct
+#| fig-cap: Linear Regression of Percentage vs Sample Size for Donald Trump
 
-# Plot the relationship for Kamala Harris
+# Plot the relationship for Trump
 ggplot(trump_data, aes(x = sample_size, y = pct)) +
@@
   geom_smooth(method = "lm", col = "blue") +
-  labs(title = "Linear Regression of Percentage vs Sample Size for Donald Trump",
-       x = "Sample Size",
+  labs(x = "Sample Size",
        y = "Percentage")
@@
 #| warning: false
+#| label: fig-mlr-harris
+#| fig-cap: Multi-Linear Regression model for Kamala Harris
+
 mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
@@
 par(mfrow = c(1, 1))  # Reset plot layout
-
 ```
 
-The results of the MLR for Kamala Harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
+The results in @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
 These results suggest that voter preferences for Harris vary widely depending on the state and the credibility of the pollster.
@@
 #| warning: false
+#| label: fig-mlr-trump
+#| fig-cap: Multi-Linear Regression model for Donald Trump
+
 # Load the saved MLR model for Donald Trump
@@
 
-The results for Donald Trump reveal that state-level variations and pollster transparency play a significant role in explaining his level of support.
+The results for Donald Trump in @fig-mlr-trump reveal that state-level variations and pollster transparency play a significant role in explaining his level of support.
 The model for Trump also passes diagnostic checks, as shown in Figure Y, indicating that the assumptions of linearity, independence, and homoscedasticity are reasonably met.
-Check Multicollinearity using Variance Inflation Factor (VIF)
+Check Multi-collinearity using Variance Inflation Factor (VIF)
 
@@
 
-To ensure that the predictors used in both models do not exhibit multicollinearity, we checked the Variance Inflation Factor (VIF) for each predictor.
-High VIF values indicate multicollinearity, which can affect the stability and reliability of the model coefficients.
+To ensure that the predictors used in both models do not exhibit multi-collinearity, we checked the Variance Inflation Factor (VIF) for each predictor.
+High VIF values indicate multi-collinearity, which can affect the stability and reliability of the model coefficients.
 
@@
 #| warning: false
-# Load the car package for VIF
-if (!require(car)) install.packages("car")
-library(car)
+#| label: tbl-harris-vif
+#| tbl-cap: Harris MLR model Variance Inflation Factor (VIF)
 
 # Check VIF for Kamala Harris model
-vif(mlr_harris_model)
+kable(vif(mlr_harris_model), digits=3)
+```
 
-# Refine the model by removing less significant predictors (e.g., methodology and transparency_score)
+Based on the VIF results in @tbl-harris-vif, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multi-collinearity.This improves the model's accuracy and interpretability.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-harris-vif
+#| fig-cap: Harris Refined Multi-Linear Regression Model
+
+# Refine the model by removing less significant predictors 
+# (e.g., methodology and transparency_score)
 mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
-
-# Summary of the refined Harris model
-
-# Check VIF for the refined model
-vif(mlr_harris_model_refined)
 
@@
 par(mfrow = c(1, 1))  # Reset plot layout
-
 ```
 
-Based on the VIF results, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multicollinearity.
-This improves the model's accuracy and interpretability.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-harris-refined-vif
+#| tbl-cap: Harris Refined MLR model Variance Inflation Factor (VIF)
+
+# Refine the model by removing less significant predictors 
+# (e.g., methodology and transparency_score)
+
+# Check VIF for the refined model
+kable(vif(mlr_harris_model_refined), digits=3)
+```
 
@@
 #| warning: false
+#| label: fig-bbb
+#| fig-cap: bbb
+
 # Perform stepwise selection to optimize the Harris model
-step_model <- step(mlr_harris_model_refined, direction = "both")
+step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
 
@@
 par(mfrow = c(1, 1))  # Reset layout
-
 ```
@@
 #| warning: false
+#| label: fig-ccc
+#| fig-cap: ccc
+
 # Final models for Harris and Trump
@@
 #| warning: false
+#| label: fig-ddd
+#| fig-cap: ddd
+
 # Define electoral votes for each state (and Washington, D.C.)
@@
 
-Kamala Harris The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
+The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
 The model shows that voter preferences vary significantly across different states, with certain states (e.g., California and Maryland) showing higher support levels, while others (e.g., Indiana and Missouri) show lower support.
@@
 
-Poll Reliability and Sampling Bias: The accuracy of the predictions depends heavily on the quality of the polling data.
+- **Poll Reliability and Sampling Bias:** The accuracy of the predictions depends heavily on the quality of the polling data.
 Although the model accounts for pollster ratings, polling methodologies can still introduce biases, particularly in states with limited polling data.
 Sampling errors and non-response biases could skew the results, especially in smaller states or regions with inconsistent polling coverage.
-
-Static Prediction: The model provides a static snapshot of voter support based on current polling data, which may not capture the dynamic nature of voter preferences over time.
+- **Static Prediction:** The model provides a static snapshot of voter support based on current polling data, which may not capture the dynamic nature of voter preferences over time.
 As election day approaches, voter preferences may shift due to campaign events, debates, or other external factors.
 Without time-series data, the model may fail to account for these trends.
+- **Unaccounted Variables:** Although the model includes important predictors like state, pollscore, and transparency score, other potentially influential factors, such as economic conditions, campaign spending, and voter turnout, are not included in the analysis.These unaccounted variables may introduce inaccuracies in the final predictions.
 
-Unaccounted Variables: Although the model includes important predictors like state, pollscore, and transparency score, other potentially influential factors, such as economic conditions, campaign spending, and voter turnout, are not included in the analysis.
-These unaccounted variables may introduce inaccuracies in the final predictions.
-## Future Research Directions Future work could improve upon this analysis by addressing some of the limitations mentioned above.
+## Future Research 
+Directions Future work could improve upon this analysis by addressing some of the limitations mentioned above.
 For instance, incorporating time-series data could allow the model to capture how voter preferences evolve in response to external factors such as campaign events, economic developments, and political endorsements.
@@
 
-# Appendix {.unnumbered}
+# Appendix
 
-# Additional data details
+## YouGov Pollster Methodology Overview and Evaluation
+YouGov conducts online surveys through their proprietary panel of U.S.
+adults, using nonprobability sampling methods combined with sophisticated weighting procedures to achieve representative results.
+Their approach balances speed and cost-effectiveness with statistical rigor through careful sample selection and data quality controls.
 
-# Model details {#sec-model-details}
+### Survey Population and Sampling
+YouGov's target population typically comprises all U.S.
+adults or adult citizens, with their sampling frame consisting of their opt-in online panel covering approximately 95% of Americans.
+For general population surveys, they aim for 1,000-2,000 respondents, selected based on demographic and political characteristics to match the target population.
 
-## Posterior predictive check
+### Panel Recruitment and Participation
+Panel members are recruited through advertising and website partnerships, with surveys offered in multiple languages including Spanish to ensure broad representation.
+Participants receive points exchangeable for small monetary rewards, though many report being motivated by the desire to contribute to research.
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
-This shows...
+### Quality Control
+YouGov employs several measures to maintain data quality:
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
-This shows...
+- Verification of panelist identity through email and IP checks
+- Response quality surveys to gauge reliability
+- Monitoring of response times and patterns
+- Removal of respondents who fail quality checks
+- Question randomization to reduce bias
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+### Non-response and Weighting
+To address potential biases, YouGov applies statistical weighting based on demographics (age, gender, race, education) and political factors (voting behavior, party identification).
+Their weighting process considers multiple characteristics simultaneously to better reflect real-world demographic intersections.
 
-```
+### Strengths and Limitations
+The methodology's primary strengths include rapid data collection, cost-effectiveness, and the ability to track opinions over time.
+However, the nonprobability sampling approach may introduce biases, and the online-only format could underrepresent certain populations.
+While weighting helps address these limitations, it cannot fully account for all potential sources of bias.
 
-## Diagnostics
+## Idealized Survey Methodology
+This idealized survey methodology outlines a comprehensive plan for forecasting the US presidential election within a budget of $100,000.
+The approach is designed to be statistically sound, practical, and capable of accurately predicting election outcomes by considering both the popular vote and electoral college implications.
 
-@fig-stanareyouokay-1 is a trace plot.
-It shows...
-This suggests...
+### Sampling Strategy
+The target population for this survey is eligible voters across the United States who are likely to participate in the upcoming presidential election. To achieve a representative sample:
 
-@fig-stanareyouokay-2 is a Rhat plot.
-It shows...
-This suggests...
+- **Sampling Frame:** Utilize a combination of registered voter lists and demographic data from reputable sources such as the US Census Bureau.
+- **Sampling Method:** Implement stratified random sampling to ensure representation across key demographics, including age, gender, race, education level, and geographic location.
+- **Sample Size Calculation:** Aim for a sample size of approximately 10,000 respondents to achieve a margin of error of ±1% at a 95% confidence level.
+- **Geographical Distribution:** Allocate samples proportionally across all 50 states and the District of Columbia, with oversampling in swing states to better predict electoral college outcomes.
+- **Addressing Sampling Biases:** Apply weighting adjustments to account for underrepresented groups and ensure that the sample mirrors the overall voter population.
 
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+### Recruitment Plan
+To recruit respondents effectively, we will leverage online panels, social media advertising, and partnerships with community organizations to reach a diverse audience.
+Offering modest incentives, such as $5 digital gift cards, encourages participation while managing costs.
+Quota sampling within strata maintains demographic balance, and follow-up reminders along with mobile-friendly survey formats help reduce non-response bias.
+The data collection will occur over a two-week period to capture timely opinions without introducing temporal biases.
 
-```
+### Survey Design Elements
+The survey is crafted to elicit accurate and meaningful responses:
+
+- **Question Types and Formats:** Use a mix of closed-ended questions and multiple-choice options for clarity and ease of analysis.
+- **Response Options:** Include balanced and neutral response choices, with options for "Undecided" or "Prefer not to say."
+- **Question Order and Flow:** Begin with general questions to build rapport, followed by more specific vote intention queries, and conclude with demographic questions.
+- **Demographic Information:** Collect data on age, gender, race, education, income, and geographic location.
+- **Political Affiliation and History:** Ask about party affiliation, past voting behavior, and political engagement.
+- **Likely Voter Screens:** Include questions to gauge voting likelihood, such as past voting frequency and intention to vote in the upcoming election.
+- **Vote Intention Questions:** Directly ask which candidate the respondent intends to vote for, ensuring confidentiality and anonymity.
+
+### Quality Control
+To maintain data integrity, we implement several quality control measures.
+Real-time validation checks within the survey prevent inconsistent or illogical responses.
+Attention-check questions identify disengaged respondents.
+We use unique survey links and track IP addresses to prevent duplicate submissions, while CAPTCHA verification deters automated responses.
+Incomplete or suspicious responses are excluded during data cleaning to ensure the final dataset is robust and reliable.
+
+### Data Processing
+These data processing steps will be taken to ensure accurate analysis:
+
+- **Weighting Methodology:** Adjust survey results using weighting factors based on demographic proportions in the voting population.
+- **Handling Missing Data:** Employ imputation techniques or exclude cases with significant missing information.
+- **Outlier Detection:** Identify and review outliers that may skew results, determining whether to retain or discard them.
+- **Response Validation:** Cross-check responses for consistency and plausibility.
+- **Poll Aggregation Approach:** Combine survey data with other reputable polls using meta-analytic techniques to enhance prediction accuracy.
+
+### Budget Allocation
+A budget allocation of $100,000 ensures all aspects are adequately funded:
+
+- **Recruitment Costs:** $40,000 for advertising and partnerships to reach potential respondents.
+- **Incentive Payments:** \$50,000 allocated for participant incentives ($5 x 10,000 respondents).
+- **Survey Platform Fees:** $2,000 for premium features on a survey platform like Google Forms or an equivalent.
+- **Data Analysis Tools:** $3,000 for statistical software licenses and data processing tools.
+- **Quality Control Measures:** $3,000 for implementing validation systems and CAPTCHA services.
+- **Administrative Costs:** $2,000 for project management and miscellaneous expenses.
+
+### Conclusion
+This methodology presents a feasible and thorough plan to forecast the US presidential election within the specified budget.
+By adhering to best practices in survey design and execution, and by carefully considering both the popular vote and electoral college implications, the survey aims to provide accurate and reliable insights into voter intentions.
 
@@
 
-#References
+# References

### 2024-10-22 paper/paper.qmd
moved regression model graphs to appendix

@@
 
+mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
+mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
 ```
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-mlr-harris
-#| fig-cap: Multi-Linear Regression model for Kamala Harris
-
-mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
-
-# Diagnostic plots for the MLR model
-par(mfrow = c(2, 2))
-plot(mlr_harris_model)
-par(mfrow = c(1, 1))  # Reset plot layout
-```
-
 The results in @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
@@
 Similarly, we constructed an MLR model for Donald Trump using the same set of predictors to assess the factors that influence his support across the country.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-mlr-trump
-#| fig-cap: Multi-Linear Regression model for Donald Trump
-
-# Load the saved MLR model for Donald Trump
-mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
-
-# Diagnostic plots for the MLR model
-par(mfrow = c(2, 2))  # Set up 2x2 layout for multiple plots
-plot(mlr_trump_model)  # Generates diagnostic plots: residuals, QQ plot, etc.
-par(mfrow = c(1, 1))  # Reset the plot layout to normal
-
-```
 
@@
 kable(vif(mlr_harris_model), digits=3)
+
+mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
 ```
@@
 Based on the VIF results in @tbl-harris-vif, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multi-collinearity.This improves the model's accuracy and interpretability.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-harris-vif
-#| fig-cap: Harris Refined Multi-Linear Regression Model
-
-# Refine the model by removing less significant predictors 
-# (e.g., methodology and transparency_score)
-mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
-
-# Plot diagnostic plots for the refined model
-par(mfrow = c(2, 2))
-plot(mlr_harris_model_refined)
-par(mfrow = c(1, 1))  # Reset plot layout
-```
 
@@
 kable(vif(mlr_harris_model_refined), digits=3)
+
+step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
 ```
@@
 To further optimize the MLR models, we performed stepwise model selection, which systematically adds or removes predictors to minimize the Akaike Information Criterion (AIC) and improve model fit.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-bbb
-#| fig-cap: bbb
-
-# Perform stepwise selection to optimize the Harris model
-step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
-
-# Plot diagnostic plots for the stepwise model
-par(mfrow = c(2, 2))
-plot(step_model)
-par(mfrow = c(1, 1))  # Reset layout
-```
 
@@
 
+# Appendix
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-mlr-harris
+#| fig-cap: Multi-Linear Regression model for Kamala Harris
+#| fig.width: 10
+#| fig.height: 4
+
+mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
+
+# For Harris model
+par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
+plot(mlr_harris_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
+par(mfrow = c(1, 1))  # Reset plot layout
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-mlr-trump
+#| fig-cap: Multi-Linear Regression model for Donald Trump
+#| fig.width: 10
+#| fig.height: 4
+
+# Load the saved MLR model for Donald Trump
+mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
+
+# For Trump model
+par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
+plot(mlr_trump_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
+par(mfrow = c(1, 1))  # Reset plot layout
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-harris-vif
+#| fig-cap: Harris Refined Multi-Linear Regression Model
+#| fig.width: 10
+#| fig.height: 4
+
+# Refine the model by removing less significant predictors 
+# (e.g., methodology and transparency_score)
+mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
+
+# For Harris refined model
+par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
+plot(mlr_harris_model_refined, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
+par(mfrow = c(1, 1))  # Reset plot layout
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-harris-step
+#| fig-cap: Harris Final Multi-Linear Regression Model
+#| fig.width: 10
+#| fig.height: 4
+
+# Perform stepwise selection to optimize the Harris model
+step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
+
+# For Harris stepwise model
+par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
+plot(step_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
+par(mfrow = c(1, 1))  # Reset plot layout
+
+```
+
+\newpage
+
 # References

### 2024-10-24 paper/paper.qmd

@@
 
-# Create a boxplot showing the distribution of support for each candidate
-ggplot(data, aes(x = candidate_name, y = pct, fill = candidate_name)) +
-  geom_boxplot() +
-  labs(x = "Candidate",
-       y = "Percentage of Support") +
+ggplot(data, aes(x = pct, fill = candidate_name)) +
+  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
+  facet_wrap(~ candidate_name) +
+  labs(x = "Percentage of Support",
+       y = "Count") +
+  scale_fill_discrete(name = "Candidates") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(legend.position = "bottom")
+
 ```

### 2024-10-24 paper/paper.qmd
Edits to intro and data

--- 2024-10-22 paper/paper.qmd
@@
 Presidential Election.
-Using polling data from various sources, we develop multiple linear regression models to predict the percentage of support for the main candidates, Kamala Harris and Donald Trump, in different states.
+Using polling data from FiveThirtyEight, we develop multiple linear regression models to predict the percentage of support for the main candidates, Kamala Harris and Donald Trump, in different states.
 We aggregate these predictions to simulate the Electoral College vote and estimate the likelihood of either candidate winning.
@@
 
-Forecasting elections has long been one of the most challenging tasks for political scientists, statisticians, and analysts alike.
-The complexity of predicting voter behavior in large, diverse electorates like that of the United States stems from numerous factors, including ever-shifting public opinion, rapidly changing political climates, and varying levels of voter engagement.
-The 2024 U.S.
-Presidential Election is no exception.
-With polarized voter bases, unpredictable external factors, and the growing influence of non-traditional media, the race remains highly uncertain, making reliable forecasting models more important—and difficult—than ever.
+The 2024 United States presidential election presents unprecedented challenges for electoral forecasting.
+As the country navigates increasing political polarization and evolving voting patterns, the reliability of traditional polling methods has come under intense scrutiny [@viala2024].
+The task of predicting voter behavior in America's diverse electorate is complicated by numerous factors, including shifting public opinion, rapidly changing political landscapes, and varying levels of voter engagement across different demographic groups.
 
-Central to any electoral forecast is the use of polling data.
-Polls provide snapshots of voter intentions at specific points in time, shaping both public perceptions and expert expectations.
-Yet, polling comes with its own set of challenges.
-Pollsters employ different methodologies—ranging from online surveys to phone interviews—each with its own inherent biases.
-Sample populations may vary in size and representativeness, potentially skewing results in favor of certain demographic groups.
-Moreover, polling accuracy can be further complicated by factors such as non-response bias, the changing nature of the electorate, and the rise of unconventional voting patterns, particularly among younger and minority voters.
+Recent history has highlighted the complexities of election forecasting.
+The polling failures in 2016 and 2020—where polls significantly underestimated Republican support in key states—have prompted a fundamental reassessment of polling methodologies [@keeter2024].
+These challenges are particularly acute in swing states, where margins of victory are often razor-thin and can determine the outcome of the entire election.
+The American Association for Public Opinion Research (AAPOR) identified several critical factors contributing to these polling errors, including the underrepresentation of Republican voters and difficulties in predicting voter turnout patterns [@viala2024].
 
-In this paper, we attempt to forecast the outcome of the 2024 U.S.
-Presidential Election by developing statistical models based on polling data from a variety of sources.
-By leveraging multiple linear regression models, we aim to predict the percentage of support for the main candidates—Kamala Harris and Donald Trump—across different states.
-Our analysis incorporates key variables such as pollster reliability, sample size, and state-level demographics, all of which contribute to shaping the electoral landscape.
-Ultimately, we aggregate these state-level predictions to simulate the outcome of the Electoral College, providing insights into the likelihood of either candidate securing the 270 votes required to win the presidency.
+Survey methodology plays a crucial role in addressing these challenges.
+Well-designed surveys require careful consideration of sampling strategies, questionnaire design, and data collection methods to ensure representative results.
+As Keeter [-@keeter2024] emphasizes, pollsters must now employ sophisticated weighting procedures and rigorous quality controls to overcome declining response rates and potential partisan non-response bias.
+Understanding the strengths and limitations of different polling approaches—from traditional probability sampling to newer online panels—is essential for accurate electoral forecasting.
 
-The rest of this paper is structured as follows: @sec-data discusses the data used for this analysis, including key variables and sources.
-@sec-model outlines the models developed for each candidate and presents the corresponding results.
-@sec-predict discusses the aggregated Electoral College predictions based on the model outputs. Finally, @sec-discuss provides conclusions and offers suggestions for future research.
+This paper develops statistical models to forecast the outcome of the 2024 presidential election between Kamala Harris and Donald Trump.
+By leveraging multi-linear regression models, we predict the percentage of support for each candidate across different states, incorporating key variables such as pollster rating, sample size, and state-level demographics.
+Through aggregating these state-level predictions, we simulate Electoral College outcomes to provide insights into each candidate's probability of securing the required 270 electoral votes.
+Our analysis also includes a detailed examination of YouGov's polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
+
+The remainder of this paper is structured as follows.
+@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
+@sec-model outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
+@sec-predict presents our Electoral College predictions based on the model outputs.
+@sec-discuss discusses the implications of our findings and suggests directions for future research.
+Finally, @sec-appendix evaluates YouGov's polling methodology, and our idealized survey methodology.
 
@@
 ## Overview
+Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news.
+This database compiles and standardizes polling results from various organizations, applying quality metrics and assessments to each poll.
+Several key variables from this dataset are crucial to our analysis:
 
-The dataset used in this analysis draws from a variety of polling sources, providing comprehensive information about polling organizations, sample sizes, methodologies, and state-level data.
-Several key variables are crucial to the analysis:
-
-- **Pollster rating:** A numerical score reflecting the reliability and historical accuracy of each polling organization.
+- **Pollster rating (Pollscore):** A numerical score reflecting the reliability and historical accuracy of each polling organization.
 - **Sample size:** The number of respondents included in each poll, which influences the poll's margin of error.
-- **Support percentage:** The percentage of respondents expressing support for each candidate.
+- **Support percentage (pct):** The percentage of respondents expressing support for each candidate.
 - **State:** The U.S. state where the poll was conducted, or in some cases, national-level polling data.
 
-To ensure accuracy and consistency, the data has been meticulously processed and cleaned.
-This involved aligning all variables for compatibility with regression analysis, converting relevant fields (e.g.
-support percentage) into numeric formats, and appropriately handling missing data to minimize potential biases.
-These reprocessing steps ensure the dataset is tidy and ready for model development.
-
-@tbl-overview_data shows a sample of the dataset.
+@tbl-overview_data presents a sample of the processed dataset, showcasing the key variables necessary for the analysis.
 
@@
 #| label: tbl-overview_data
-# Select key variables for the table
+#| tbl-cap: Sample Overview of Selected Variables in the Polling Dataset
+
+set.seed(123)
+# Select key variables and sample for the table
 data_overview <- data %>%
-  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
+  select(poll_id, pollster, pollscore, sample_size, pct, state, 
+         candidate_name, end_date) %>% sample_n(5)
 
-# Show the first 5 rows of the dataset in a table
-data_overview %>%
-  head(5) %>%
-  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
+colnames(data_overview) <- c("Poll ID", "Pollster", "Pollscore", "Sample Size",
+                             "pct", "State", "Candidate", "End Date")
 
+kable(data_overview)
 ```
@@
 ## Measurement and Limitations
+There are several measurement and limitation considerations for our dataset:
 
-The primary limitations of this analysis stem from the quality and variability of polling data, which can introduce biases through factors like sample size and pollster methodology.
-Additionally, the model provides a static snapshot of voter preferences, without accounting for the dynamic nature of elections For instance, shifts in public opinion over time or the impact of campaign events.
-The exclusion of key external factors, such as economic conditions or voter turnout, also limits the model's ability to fully capture the complexities of election outcomes.
-Furthermore, the simplified Electoral College simulation assumes that polling data will accurately predict state-level results, which may not always be the case, especially in battleground states with volatile voter behavior.
+- **Poll Quality:** While we filter for high-quality polls using numeric grades, differences in polling methodologies may introduce systematic biases.
+The inclusion of pollster ratings helps account for historical accuracy but cannot completely eliminate these potential biases.
+- **Temporal Dynamics:** Our dataset provides a snapshot of voter preferences during a specific timeframe.
+This static nature means we cannot capture the full dynamics of voter preference evolution over the campaign period.
+- **Geographic Coverage:** Although we have national and state-level polling data, coverage varies by state.
+Battleground states typically have more frequent polling, while safer states may have sparse data, potentially affecting our state-level predictions.
+- **Response Bias:** Despite careful methodology by pollsters, self-selection bias in survey participation and social desirability bias in responses remain potential concerns.
 
 ## Outcome variables
-Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll.
-This variable is central to our analysis because it directly captures voter preference, which is the most relevant metric for forecasting election outcomes.
-By modeling the percentage of support, we can quantify how various factors—such as pollster quality, sample size, and state-level dynamics—influence the candidates' standings in the polls.
-This outcome variable also allows us to simulate the Electoral College results by aggregating predicted support across different states, which is crucial for determining the likelihood of either candidate winning the presidency.
+Our primary outcome variable is the percentage of support (*pct*) for Kamala Harris and Donald Trump in each poll.
+This measurement represents the proportion of respondents who indicate they would vote for each candidate if the election were held on the day of the poll.
+The variable directly captures voter preferences and serves as the foundation for our electoral predictions.
 
-@fig-support below shows the distribution of support for both candidates across the polls included in the dataset.
+@fig-support below illustrates the distribution of support for both candidates, revealing several notable patterns.
+First, the support percentages cluster between 40% and 60%, reflecting the competitive nature of the race.
+Second, the distributions show slight differences between candidates, with Harris's support displaying more variation than Trump's.
+This pattern might reflect differences in voter certainty or polling methodology across different states and time periods.
 
@@
 #| warning: false
-#| fig.pos: "H"
 #| label: fig-support
@@
 
-# Create a boxplot showing the distribution of support for each candidate
-ggplot(data, aes(x = candidate_name, y = pct, fill = candidate_name)) +
-  geom_boxplot() +
-  labs(x = "Candidate",
-       y = "Percentage of Support") +
+ggplot(data, aes(x = pct, fill = candidate_name)) +
+  geom_histogram(bins = 30, alpha = 0.8, position = "identity") +
+  facet_wrap(~ candidate_name) +
+  labs(x = "Percentage of Support",
+       y = "Count") +
+  scale_fill_discrete(name = "Candidates") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(legend.position = "bottom")
 ```
@@
 
-In our analysis, we utilized several key predictor variables to model voter support for Kamala Harris and Donald Trump in the 2024 U.S.
-Presidential Election.
-These predictor variables were selected based on their expected influence on polling results and their ability to account for variations in polling accuracy and voter behavior across different states.
+Our model incorporates several key predictor variables, each chosen for its theoretical importance in explaining polling variations and electoral outcomes:
 
-One of the key predictors is numeric grade, which represents a composite measure of poll quality.
-This variable accounts for various aspects of a poll’s reliability, such as the transparency of its methodology and the pollster’s historical accuracy.
-Pollscore, another important predictor, provides a similar measure of poll reliability but focuses more on the specific performance of the polling organization in recent elections.
+- **Numeric Grade:** A composite measure (scale: 0-4) incorporating factors such as methodology rigor and historical accuracy
+- **Pollscore:** A measure of historical polling accuracy (range: -4 to +4, where negative scores indicate better performance)
+- **Transparency Score:** Quantifies the openness of polling methodology (scale: 0-10)
+- **Sample Size:** Number of respondents, typically ranging from 500 to 3000
+- **State:** State-level indicators capturing regional political variations
+- **Methodology:** Survey approach (e.g., online panel, phone interviews)
 
-Transparency score measures how openly polling organizations disclose their methodology and data, which can significantly impact the trustworthiness of the results.
-Polls with higher transparency scores are given more weight in the model, as they are generally considered more reliable.
-The sample size of each poll is also included, as larger sample sizes tend to reduce the margin of error and provide more precise estimates of voter support.
+The relationship between these predictors and polling outcomes is complex and often interconnected.
+For example, while larger sample sizes generally reduce sampling error, this effect may be moderated by the poll's methodology and quality metrics.
+Similarly, geographic variations in polling accuracy suggest that the relationship between predictors and outcomes may vary systematically across states.
 
-Geographic differences in voter preferences are captured through the state variable, which accounts for regional variations in political culture, demographics, and historical voting patterns.
-Lastly, methodology refers to the techniques and approaches used by the polling organizations, such as whether the poll was conducted online, by phone, or through other means.
+These variables were selected based on both theoretical foundations in electoral polling literature and practical considerations of data availability and quality.
+We focused on these core variables due to their consistent availability across polls and demonstrated importance in previous electoral forecasting efforts.
 
-All predictor variables—numeric grade, pollscore, transparency score, sample size, state, and methodology—are presented in @tbl-overview_data.
+**Edit note: Add a table in appendix to show sample with all variables used in model**
 
@@
 
-The data cleaning process uses R [@citeR] as well as other packages such as tidyverse [@tidyverse], dplyr [@dplyr], janitor [@janitor] and lubridate [@lubridate].
-Begin by standardizing column name to ensure consistency and avoid errors during analysis.
-The dataset was filtered to focus solely on Kamala Harris and Donald Trump, with polls below a numeric grade of 2.7 excluded to maintain high-quality data.
-Missing state information was categorized as “National” polls, and the poll end dates were converted to a date format to include only polls conducted after the candidates officially declared their candidacies.
-This ensures the analysis reflects relevant, up-to-date voter preferences.
-Additionally, transformations such as calculating the number of supporters from the percentage support and encoding a binary variable for the candidates (Harris = 1, Trump = 0) were performed to enhance usability in the regression models.
+The data cleaning process employed R [@citeR] along with several specialized packages: tidyverse [@tidyverse] for data manipulation, dplyr [@dplyr] for data transformation, janitor [@janitor] for consistent naming conventions, and lubridate [@lubridate] for date handling.
 
-Some variables were excluded for simplicity and relevance.
-For example, **'population_full'** and **'answer'** were dropped as they provided redundant or unnecessary information for this analysis.
-However, key variables like **'pollster'**, **'pollscore'**, **'numeric grade'**, and **'sample size'** were retained because they provide essential insights into the quality and precision of the polls, directly influencing the reliability of the models.
-These steps ensured the cleaned dataset was ready for analysis, with a focus on the most important predictors of voter support while minimizing extraneous information.
+Our cleaning process followed several key steps:
+
+1. Filtered for high-quality polls using a minimum threshold (*numeric_grade* $≥$ 2.7)
+2. Limited temporal coverage to post-campaign announcement period (after July 21, 2024)
+3. Standardized geographic data:
+
+  > i. Coded missing state information as "National" polls
+  > ii. Verified state names for consistency
+
+4. Transformed key variables:
+
+  > i. Converted end dates to standardized date format
+  > ii. Calculated absolute supporter numbers from percentages and sample sizes
+  > iii. Created binary candidate indicators (Harris = 1, Trump = 0)
+
+Our variable selection process prioritized measures essential for polling accuracy and electoral prediction while eliminating redundant or non-informative fields.
+We retained key poll quality metrics including *pollscore* and *numeric_grade*, which provide crucial information about the reliability of each survey.
+Sample characteristics such as *sample size* and *methodology* were preserved to account for differences in polling precision.
+*State* and polling *end_date* information were maintained to capture regional variations and time-dependent patterns in voter preferences.
+
+We excluded several variables that offered little additional analytical value, such as *population_full*, which duplicated information available in other fields, and administrative data that did not directly influence predictions.
+This focused approach to variable selection allowed us to preserve all information necessary for accurate electoral forecasting.
 
@@
 
-The results in @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
+The results @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
 These results suggest that voter preferences for Harris vary widely depending on the state and the credibility of the pollster.
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 
@@
 
+\newpage
 ## Idealized Survey Methodology
@@
 
-# Appendix
+## Multi-Linear Regression Models
 

### 2024-10-24 paper/paper.qmd
Abstract edit

@@
 date-format: long
-abstract: "This paper presents a statistical model for forecasting the outcome of the 2024 U.S.
-Presidential Election.
-Using polling data from FiveThirtyEight, we develop multiple linear regression models to predict the percentage of support for the main candidates, Kamala Harris and Donald Trump, in different states.
-We aggregate these predictions to simulate the Electoral College vote and estimate the likelihood of either candidate winning.
-Our results suggest that neither candidate currently secures the required 270 electoral votes to win the presidency."
+abstract: "Analysis of polling data from July to October 2024 reveals significant challenges in forecasting the U.S.
+presidential election between Kamala Harris and Donald Trump, with state-level variations and pollster methodology playing crucial roles in prediction accuracy.
+Using multi-linear regression models incorporating pollster ratings, sample sizes, and state-level data, we find that Harris is projected to receive 216 electoral votes while Trump is predicted to secure 147 electoral votes.
+State-level polling variations and methodological differences among pollsters significantly impact predicted outcomes, with pollster reliability metrics explaining substantial variance in candidate support levels.
+These findings highlight the continued importance of refining polling methodologies and expanding state-level coverage to improve electoral forecasting accuracy, particularly in an era of increasing political polarization."
 format: pdf
@@
 - **State:** The U.S. state where the poll was conducted, or in some cases, national-level polling data.
+- ...
 
@@
 data_overview <- data %>%
-  select(poll_id, pollster, pollscore, sample_size, pct, state, 
-         candidate_name, end_date) %>% sample_n(5)
+  select(poll_id, pollster, pollscore, numeric_grade, transparency_score,
+         sample_size, pct, state, methodology, candidate_name, end_date
+         ) %>% sample_n(5)
 
-colnames(data_overview) <- c("Poll ID", "Pollster", "Pollscore", "Sample Size",
-                             "pct", "State", "Candidate", "End Date")
+colnames(data_overview) <- 
+  c("Poll ID", "Pollster", "Pollscore", "Numeric Grade", "Transparency Score", 
+    "Sample Size", "pct", "State", "Methodology", "Candidate", "End Date")
 
@@
 
-According to our models, neither candidate secures the 270 Electoral College votes needed to win.
-Kamala Harris is projected to receive 216 electoral votes, while Donald Trump is predicted to win 147 electoral votes.
-However, due to the lack of 270 electoral votes for either candidate, the election remains highly competitive, and further developments may shift the balance.
+According to our models, Kamala Harris is projected to receive 216 electoral votes, while Donald Trump is predicted to win 147 electoral votes.
 
@@
 Presidential Election.
-According to the predictions, neither Kamala Harris nor Donald Trump currently secures the necessary 270 electoral votes to win.
 Harris is projected to receive 216 electoral votes, while Trump is predicted to receive 147 electoral votes.

### 2024-10-24 paper/paperV2.qmd
version 2 plan

--- 2024-10-24 paper/paper.qmd
@@
 date-format: long
-abstract: "Analysis of polling data from July to October 2024 reveals significant challenges in forecasting the U.S.
-presidential election between Kamala Harris and Donald Trump, with state-level variations and pollster methodology playing crucial roles in prediction accuracy.
-Using multi-linear regression models incorporating pollster ratings, sample sizes, and state-level data, we find that Harris is projected to receive 216 electoral votes while Trump is predicted to secure 147 electoral votes.
-State-level polling variations and methodological differences among pollsters significantly impact predicted outcomes, with pollster reliability metrics explaining substantial variance in candidate support levels.
-These findings highlight the continued importance of refining polling methodologies and expanding state-level coverage to improve electoral forecasting accuracy, particularly in an era of increasing political polarization."
+abstract: "Do Abstract after finalizing everything"
 format: pdf
@@
 These challenges are particularly acute in swing states, where margins of victory are often razor-thin and can determine the outcome of the entire election.
-The American Association for Public Opinion Research (AAPOR) identified several critical factors contributing to these polling errors, including the underrepresentation of Republican voters and difficulties in predicting voter turnout patterns [@viala2024].
+The American Association for Public Opinion Research (AAPOR) identified several critical factors contributing to these polling errors, including the under representation of Republican voters and difficulties in predicting voter turnout patterns [@viala2024].
 
@@
 By leveraging multi-linear regression models, we predict the percentage of support for each candidate across different states, incorporating key variables such as pollster rating, sample size, and state-level demographics.
-Through aggregating these state-level predictions, we simulate Electoral College outcomes to provide insights into each candidate's probability of securing the required 270 electoral votes.
-Our analysis also includes a detailed examination of YouGov's polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
+Through aggregating these state-level predictions, we simulate Electoral College outcomes to provide insights into each candidate's probability of securing the required 270 electoral votes.Our analysis also includes a detailed examination of YouGov's polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
+
+**Edit note: will need to expand/edit after getting final results**
 
@@
 
+**Steven will do data section**
+
 ## Overview
@@
 - **State:** The U.S. state where the poll was conducted, or in some cases, national-level polling data.
-- ...
+- ...other variables
 
-@tbl-overview_data presents a sample of the processed dataset, showcasing the key variables necessary for the analysis.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-overview_data
-#| tbl-cap: Sample Overview of Selected Variables in the Polling Dataset
-
-set.seed(123)
-# Select key variables and sample for the table
-data_overview <- data %>%
-  select(poll_id, pollster, pollscore, numeric_grade, transparency_score,
-         sample_size, pct, state, methodology, candidate_name, end_date
-         ) %>% sample_n(5)
-
-colnames(data_overview) <- 
-  c("Poll ID", "Pollster", "Pollscore", "Numeric Grade", "Transparency Score", 
-    "Sample Size", "pct", "State", "Methodology", "Candidate", "End Date")
-
-kable(data_overview)
-```
+@tbl-overview_data in the appendix presents a sample of the processed dataset, showcasing the key variables necessary for the analysis.
 
@@
 
-**Edit note: Add a table in appendix to show sample with all variables used in model**
-
 ## Cleaning Process and Analysis
@@
 
-# Modeling Support for the Candidates {#sec-model}
+# Model {#sec-model}
 
-## Kamala Harris
+## Model Overview and Theoretical Framework
 
-We begin by modeling the percentage of support for Kamala Harris using a linear regression model.
-The predictors include the sample size, pollster ratings (e.g., pollscore and transparency score), and state.
-This model aims to quantify how these variables influence her support across different polls.
+- Justify the choice of multiple linear regression for election forecasting
+- Discuss why this approach is appropriate given the data structure
+- Explain how the model addresses the unique challenges of electoral prediction
+- Define the mathematical notation for the model
+- Explain each component of the model specification
+- Discuss underlying assumptions and their implications
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Load necessary library
-library(ggplot2)
+## Baseline Model Development
 
-# Filter the data for only "Kamala Harris"
-harris_data <- subset(data, candidate_name == "Kamala Harris")
+- Present the initial model specification with all variables
+- Explain the 80-20 train-test split methodology
+- Detail the model validation approach
+- Present mathematical notation:
 
-# Perform a simple linear regression with pct as the dependent variable and sample_size as the independent variable
-model <- lm(pct ~ sample_size, data = harris_data)
-```
+Report key metrics such as, but not all required. U decide:
 
-The results of the linear regression for Kamala Harris show that sample size has a statistically insignificant effect on her support.
-This suggests that factors other than sample size, such as pollster methodology or regional biases, may play a more significant role in determining the level of support she receives.
+- R-squared values
+- Mean Squared Error (MSE)
+- Root Mean Squared Error (RMSE)
+- Out-of-sample testing accuracy
 
-@fig-harris-samplesize-supportpct illustrates the relationship between the sample size and support percentage for Kamala Harris.
+## Model Refinement Using VIF Analysis
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-harris-samplesize-supportpct
-#| fig-cap: Linear Regression of Percentage vs Sample Size for Kamala Harris
-
-# Plot the relationship for Kamala Harris
-ggplot(harris_data, aes(x = sample_size, y = pct)) +
-  geom_point() +
-  geom_smooth(method = "lm", col = "blue") +
-  labs(x = "Sample Size",
-       y = "Percentage")
-```
-
-## Donald Trump
-
-A similar linear regression model was applied to Donald Trump's data.
-The predictors remain the same, and the goal is to determine the factors that drive support for him.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Load necessary library
-library(ggplot2)
-
-# Filter the data for only "Kamala Harris"
-trump_data <- subset(data, candidate_name == "Donald Trump")
-
-# Perform a simple linear regression with pct as the dependent variable and sample_size as the independent variable
-model <- lm(pct ~ sample_size, data = trump_data)
-
-```
-
-The results in indicate that sample size has a weak but statistically significant negative effect on support for Trump.
-This suggests that larger polls tend to show slightly lower support for Trump, although the effect size is small.
-
-@fig-trump-samplesize-supportpct shows the relationship between the sample size and support percentage for Donald Trump.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-trump-samplesize-supportpct
-#| fig-cap: Linear Regression of Percentage vs Sample Size for Donald Trump
-
-# Plot the relationship for Trump
-ggplot(trump_data, aes(x = sample_size, y = pct)) +
-  geom_point() +
-  geom_smooth(method = "lm", col = "blue") +
-  labs(x = "Sample Size",
-       y = "Percentage")
-
-mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
-mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
-```
-
-## Multiple Linear Regression Models
-
-To better capture the complexity of voter support, we constructed multiple linear regression (MLR) models for both Kamala Harris and Donald Trump.
-These models incorporate several predictors, including pollster rating (represented by pollscore), transparency score, sample size, and state-level data.
-By accounting for these factors, the MLR models allow us to control for more variables that influence voter support and provide more accurate predictions.
-
-### Kamala Harris Model
-
-The MLR model for Kamala Harris takes into account multiple factors that may impact her support across different polls.
-This includes not just the sample size of the poll, but also how pollster reliability (numeric_grade), transparency of poll data (transparency_score), and state-level polling contribute to predicting voter preferences.
-
-The results @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
-These results suggest that voter preferences for Harris vary widely depending on the state and the credibility of the pollster.
-
-The model diagnostics (such as residual plots, QQ plots) were evaluated to ensure the assumptions of linear regression hold.
-Figure X shows diagnostic plots, which indicate that the model performs reasonably well in terms of residual behavior and normality.
-Similarly, we build an MLR model for Donald Trump.
-
-### Donald Trump Model
-
-Similarly, we constructed an MLR model for Donald Trump using the same set of predictors to assess the factors that influence his support across the country.
-
-The results for Donald Trump in @fig-mlr-trump reveal that state-level variations and pollster transparency play a significant role in explaining his level of support.
-The model for Trump also passes diagnostic checks, as shown in Figure Y, indicating that the assumptions of linearity, independence, and homoscedasticity are reasonably met.
-Check Multi-collinearity using Variance Inflation Factor (VIF)
-
-## Multicollinearity Check Using Variance Inflation Factor (VIF)
-
-To ensure that the predictors used in both models do not exhibit multi-collinearity, we checked the Variance Inflation Factor (VIF) for each predictor.
-High VIF values indicate multi-collinearity, which can affect the stability and reliability of the model coefficients.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-harris-vif
-#| tbl-cap: Harris MLR model Variance Inflation Factor (VIF)
-
-# Check VIF for Kamala Harris model
-kable(vif(mlr_harris_model), digits=3)
-
-mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
-```
-
-Based on the VIF results in @tbl-harris-vif, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multi-collinearity.This improves the model's accuracy and interpretability.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-harris-refined-vif
-#| tbl-cap: Harris Refined MLR model Variance Inflation Factor (VIF)
-
-# Refine the model by removing less significant predictors 
-# (e.g., methodology and transparency_score)
-
-# Check VIF for the refined model
-kable(vif(mlr_harris_model_refined), digits=3)
-
-step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
-```
+- Explain the importance of addressing multicollinearity
+- Present VIF analysis methodology and thresholds used
+- Show which variables were retained/removed and why
+- Compare refined model performance to baseline
+- Present updated mathematical notation
+- Report comparative metrics showing improvement
 
@@
 
-To further optimize the MLR models, we performed stepwise model selection, which systematically adds or removes predictors to minimize the Akaike Information Criterion (AIC) and improve model fit.
+- Explain the stepwise selection methodology
+- Justify the selection criteria (AIC/BIC)
+- Present the final model specification
+- Compare performance metrics across all three models
+- Discuss why this represents the optimal model
 
-The stepwise model selection improved the model by retaining the most significant predictors and eliminating those with little explanatory power.
+## Model Validation and Diagnostics
 
-## Final Predictions for Electoral College Votes
+- Present residual analysis
+- Discuss normality assumptions
+- Address heteroscedasticity
+- Present cross-validation results
+- Discuss model limitations
 
-Using the final models for both candidates, we predicted the percentage of support in each state and aggregated the results to simulate the Electoral College outcome.
-The predicted percentage of support for each candidate is used to determine the likely winner in each state.
+# Results {#sec-result}
+## Model Performance Comparison
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ccc
-#| fig-cap: ccc
+- Present comprehensive comparison table of all three models
+- Include all relevant statistical measures
+- Visualize performance metrics
+- Discuss which model performs best and why
 
-# Final models for Harris and Trump
-mlr_harris_model_final <- readRDS(here::here(("models/mlr_harris_model_final.rds")))
-mlr_trump_model_final <- readRDS(here::here(("models/mlr_trump_model_final.rds")))
+## State-Level Predictions
 
-# Predict poll percentages for Kamala Harris
-harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
+- Present predictions for key battleground states
+- Compare model predictions with recent polling averages
+- Discuss uncertainty in predictions
+- Present confidence intervals
 
-# Predict poll percentages for Donald Trump
-trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
+## Electoral College Projections
 
-```
+- Aggregate state-level predictions into electoral votes
+- Present probability distributions for different outcomes
+- Discuss implications for election outcome
+- Address uncertainty in projections
 
-# Electoral College Prediction {#sec-predict}
+## Model Insights
 
-To forecast the winner of the 2024 election, we aggregate the predicted percentages of support from our models for each state and calculate the Electoral College votes.
-The candidate with 270 or more Electoral College votes is predicted to win the election.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ddd
-#| fig-cap: ddd
-
-# Define electoral votes for each state (and Washington, D.C.)
-electoral_votes <- data.frame(
-  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", 
-            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
-            "Maine CD-1", "Maine CD-2", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
-            "Missouri", "Montana", "Nebraska", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", "Nevada", 
-            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
-            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
-            "District of Columbia"),
-  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 2, 1, 1, 10, 11, 15, 10, 6, 10, 
-                      3, 5, 1, 1, 2, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3)
-)
-
-# Aggregate predictions by state for Harris and Trump
-harris_state_avg <- aggregate(predicted_pct_harris ~ state, data = harris_data, FUN = mean)
-trump_state_avg <- aggregate(predicted_pct_trump ~ state, data = trump_data, FUN = mean)
-
-# Merge predictions for both candidates
-prediction_comparison <- merge(harris_state_avg, trump_state_avg, by = "state", all.x = TRUE)
-
-# Merge the electoral votes with the predictions
-prediction_comparison <- merge(prediction_comparison, electoral_votes, by = "state", all.x = TRUE)
-
-# Determine the winner for each state
-prediction_comparison$winner <- ifelse(prediction_comparison$predicted_pct_harris > prediction_comparison$predicted_pct_trump, "Harris", "Trump")
-
-# Calculate total electoral votes for Kamala Harris
-harris_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Harris"], na.rm = TRUE)
-
-# Calculate total electoral votes for Donald Trump
-trump_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Trump"], na.rm = TRUE)
-
-# Print the results
-print(paste("Harris Electoral Votes:", harris_electoral_votes))
-print(paste("Trump Electoral Votes:", trump_electoral_votes))
-
-# Determine the predicted winner
-if (harris_electoral_votes >= 270) {
-  print("Kamala Harris is predicted to win the 2024 election.")
-} else if (trump_electoral_votes >= 270) {
-  print("Donald Trump is predicted to win the 2024 election.")
-} else {
-  print("No candidate reached 270 electoral votes.")
-}
-
-```
-
-According to our models, Kamala Harris is projected to receive 216 electoral votes, while Donald Trump is predicted to win 147 electoral votes.
+- Discuss which variables are most influential
+- Present effect sizes for key predictors
+- Analyze regional variations in model performance
+- Identify any surprising or counterintuitive findings
 
 # Discussion {#sec-discuss}
-
 ## Key Findings
 
-The multiple linear regression (MLR) models developed in this paper provide valuable insights into the factors that influence voter support for Kamala Harris and Donald Trump.
-Through the analysis of polling data, several key predictors were identified as significant, including state, pollster rating (pollscore), and transparency score.
-
-The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
-The model shows that voter preferences vary significantly across different states, with certain states (e.g., California and Maryland) showing higher support levels, while others (e.g., Indiana and Missouri) show lower support.
-Pollster reliability, captured through numeric grades and transparency scores, also plays an important role, suggesting that voters may respond differently based on the credibility of the pollster.
-
-Donald Trump Similarly, the MLR model for Donald Trump reveals that state-level variations are critical to understanding his level of support.
-The model indicates that Trump’s support is more stable across certain states, but there are also notable outliers where his support fluctuates.
-Pollster characteristics, such as pollscore and transparency score, significantly impact Trump’s predicted support, reflecting the importance of poll quality in predicting election outcomes.
-
-Electoral College Forecast Aggregating state-level predictions into Electoral College votes demonstrates the potential competitiveness of the 2024 U.S.
-Presidential Election.
-Harris is projected to receive 216 electoral votes, while Trump is predicted to receive 147 electoral votes.
-This forecast suggests that the election remains highly uncertain, with several key battleground states likely determining the final outcome.
+- Summarize main results and their implications
+- Compare findings with existing literature
+- Discuss how results align with or differ from previous election modeling efforts
+- Address practical significance of findings
 
@@
 
-One of the main strengths of this analysis is the incorporation of multiple predictors that allow us to account for various factors influencing voter support.
-By considering state-level data, pollster ratings, and transparency scores, the models provide a more nuanced prediction of voter behavior compared to models that rely solely on national-level polling.
-
-The use of stepwise model selection and multicollinearity checks further enhanced the robustness of the models by optimizing the choice of predictors and ensuring that the models do not suffer from unstable estimates caused by correlated predictors.
-
-Additionally, the aggregation of state-level predictions into Electoral College outcomes presents a more realistic forecast of the election, as the U.S.
-Presidential election is ultimately decided by electoral votes, not popular votes.
+- Discuss advantages of the final model
+- Compare with other approaches in literature
+- Highlight unique contributions
+- Explain practical applications
 
@@
 
-Despite the strengths of the models, there are several limitations that should be addressed:
+- Address data limitations
+- Discuss model assumptions and their impact
+- Consider external validity
+- Acknowledge potential sources of error
 
-- **Poll Reliability and Sampling Bias:** The accuracy of the predictions depends heavily on the quality of the polling data.
-Although the model accounts for pollster ratings, polling methodologies can still introduce biases, particularly in states with limited polling data.
-Sampling errors and non-response biases could skew the results, especially in smaller states or regions with inconsistent polling coverage.
-- **Static Prediction:** The model provides a static snapshot of voter support based on current polling data, which may not capture the dynamic nature of voter preferences over time.
-As election day approaches, voter preferences may shift due to campaign events, debates, or other external factors.
-Without time-series data, the model may fail to account for these trends.
-- **Unaccounted Variables:** Although the model includes important predictors like state, pollscore, and transparency score, other potentially influential factors, such as economic conditions, campaign spending, and voter turnout, are not included in the analysis.These unaccounted variables may introduce inaccuracies in the final predictions.
+## Future Research Directions
 
-## Future Research 
-Directions Future work could improve upon this analysis by addressing some of the limitations mentioned above.
-For instance, incorporating time-series data could allow the model to capture how voter preferences evolve in response to external factors such as campaign events, economic developments, and political endorsements.
-A dynamic forecasting model that updates predictions as new polls are released would provide more timely and accurate forecasts.
-
-Moreover, integrating other influential variables, such as economic indicators (e.g., unemployment rates, inflation), voter turnout models, and campaign spending data, could enhance the predictive power of the model.
-Including demographic data (e.g., age, education, income) could also improve the granularity of predictions, especially in battleground states where demographic shifts are critical to election outcomes.
-
-Finally, expanding the model to account for ranked-choice voting in certain states could provide a more accurate forecast in scenarios where third-party candidates or run-off elections play a significant role.
+- Suggest methodological improvements
+- Propose additional variables to consider
+- Discuss potential extensions of the model
+- Recommend practical applications
 
@@
 
-## Multi-Linear Regression Models
-
+## Sample Data
 ```{r}
@@
 #| warning: false
-#| label: fig-mlr-harris
-#| fig-cap: Multi-Linear Regression model for Kamala Harris
-#| fig.width: 10
-#| fig.height: 4
+#| label: tbl-overview_data
+#| tbl-cap: Sample Overview of Selected Variables in the Polling Dataset
 
-mlr_harris_model <- readRDS(here::here(("models/mlr_harris_model.rds")))
+set.seed(123)
+# Select key variables and sample for the table
+data_overview <- data %>%
+  select(poll_id, pollster, pollscore, numeric_grade, transparency_score,
+         sample_size, pct, state, methodology, candidate_name, end_date
+         ) %>% sample_n(5)
 
-# For Harris model
-par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
-plot(mlr_harris_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
-par(mfrow = c(1, 1))  # Reset plot layout
+colnames(data_overview) <- 
+  c("Poll ID", "Pollster", "Pollscore", "Numeric Grade", "Transparency Score", 
+    "Sample Size", "pct", "State", "Methodology", "Candidate", "End Date")
 
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-mlr-trump
-#| fig-cap: Multi-Linear Regression model for Donald Trump
-#| fig.width: 10
-#| fig.height: 4
-
-# Load the saved MLR model for Donald Trump
-mlr_trump_model <- readRDS(here::here(("models/mlr_trump_model.rds")))
-
-# For Trump model
-par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
-plot(mlr_trump_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
-par(mfrow = c(1, 1))  # Reset plot layout
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-harris-vif
-#| fig-cap: Harris Refined Multi-Linear Regression Model
-#| fig.width: 10
-#| fig.height: 4
-
-# Refine the model by removing less significant predictors 
-# (e.g., methodology and transparency_score)
-mlr_harris_model_refined <- readRDS(here::here(("models/mlr_harris_model_refined.rds")))
-
-# For Harris refined model
-par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
-plot(mlr_harris_model_refined, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
-par(mfrow = c(1, 1))  # Reset plot layout
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-harris-step
-#| fig-cap: Harris Final Multi-Linear Regression Model
-#| fig.width: 10
-#| fig.height: 4
-
-# Perform stepwise selection to optimize the Harris model
-step_model <- step(mlr_harris_model_refined, direction = "both", trace=0)
-
-# For Harris stepwise model
-par(mfrow = c(1, 2))  # Set layout for 1 row and 2 columns
-plot(step_model, which = c(1, 2))  # 1: Residuals vs Fitted, 2: Normal Q-Q
-par(mfrow = c(1, 1))  # Reset plot layout
-
+kable(t(data_overview))
 ```
@@
 
+## Multi-Linear Regression Models
+
+\newpage
+
 # References

### 2024-10-31 paper/paper.qmd
model has been verified with test and train data, moved graphs to appendix, and changed writing

--- 2024-10-24 paper/paper.qmd
@@
 title: "Forecasting the 2024 U.S. Presidential Election"
-subtitle: "My subtitle if needed"
+subtitle: "Modeling State-Level Polling Data to Forecast the Electoral College Outcome"
 author: 
@@
   - Tommy Fu
-thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git"
+thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git."
 date: today
 date-format: long
-abstract: "This paper presents a statistical model for forecasting the outcome of the 2024 U.S.
-Presidential Election.
-Using polling data from various sources, we develop multiple linear regression models to predict the percentage of support for the main candidates, Kamala Harris and Donald Trump, in different states.
-We aggregate these predictions to simulate the Electoral College vote and estimate the likelihood of either candidate winning.
-Our results suggest that neither candidate currently secures the required 270 electoral votes to win the presidency."
+abstract: "We forecast the outcome of the 2024 U.S.
+Presidential Election between Kamala Harris and Donald Trump by developing multiple linear regression models based on comprehensive polling data collected throughout the election cycle.
+Incorporating variables such as state-level demographics, pollster reliability scores, transparency scores, and sample sizes, we applied the same statistical model to both candidates to ensure consistent comparison.
+Our analysis predicts that Kamala Harris will receive 216 electoral votes, while Donald Trump is projected to secure 147 electoral votes.
+Neither candidate achieves the 270 electoral votes required to win the presidency, highlighting the potential for a closely contested election.
+These findings underscore the significant impact of state-specific factors on voter support and suggest that neither candidate currently holds a decisive advantage.
+We recommend that future research includes dynamic modeling techniques and additional predictive variables, such as economic indicators and voter turnout rates, to enhance the accuracy of election forecasts.
+Our study emphasizes the complexities involved in electoral predictions and the necessity of balancing multiple factors in policy design and electoral analysis."
 format: pdf
@@
 
-The dataset used in this analysis draws from a variety of polling sources, providing comprehensive information about polling organizations, sample sizes, methodologies, and state-level data.
-Several key variables are crucial to the analysis:
+The dataset employed in this analysis encompasses polling data from various reputable sources, capturing critical variables related to polling organizations, sample sizes, candidate support levels, and polling coverage (state-level or national).
+The primary variables of interest are outlined as follows:
 
-- **Pollster rating:** A numerical score reflecting the reliability and historical accuracy of each polling organization.
-- **Sample size:** The number of respondents included in each poll, which influences the poll's margin of error.
-- **Support percentage:** The percentage of respondents expressing support for each candidate.
-- **State:** The U.S. state where the poll was conducted, or in some cases, national-level polling data.
+- **Poll ID**: A unique identifier for each poll entry, facilitating data tracking and management (e.g., 88590).
+- **Pollster**: The organization conducting the poll, which serves as a primary indicator of the poll’s methodological quality (e.g., YouGov, Ipsos).
+- **Pollscore**: A quantitative measure of the pollster’s reliability, where lower (negative) values suggest higher predictive accuracy (e.g., -1.1 for YouGov).
+- **Sample Size**: The number of respondents included in each poll, influencing the statistical precision and margin of error (e.g., 1414).
+- **Support Percentage (pct)**: The proportion of respondents expressing support for each candidate, serving as the dependent variable in subsequent analyses (e.g., 47% for Kamala Harris).
+- **State**: Indicates the geographical focus of the poll, either at the state level or nationwide (e.g., National).
+- **Candidate Name**: The candidate assessed in the poll, providing context for support levels and enabling candidate-specific analyses (e.g., Kamala Harris, Donald Trump).
+- **End Date**: The date on which the poll concluded, offering temporal alignment for longitudinal analyses (e.g., 2024-10-07).
 
-To ensure accuracy and consistency, the data has been meticulously processed and cleaned.
-This involved aligning all variables for compatibility with regression analysis, converting relevant fields (e.g.
-support percentage) into numeric formats, and appropriately handling missing data to minimize potential biases.
-These reprocessing steps ensure the dataset is tidy and ready for model development.
+To ensure analytical rigor, the dataset underwent extensive preprocessing.
+This involved standardizing variable formats for compatibility with regression models, converting support percentages into numerical values, and systematically addressing missing data to minimize bias.
+These preprocessing steps ensured that the dataset adhered to the standards of a "tidy" dataset, enabling robust model development and enhancing interpretability of subsequent statistical analyses.
 
 @tbl-overview_data shows a sample of the dataset.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-overview_data
-# Select key variables for the table
-data_overview <- data %>%
-  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
-
-# Show the first 5 rows of the dataset in a table
-data_overview %>%
-  head(5) %>%
-  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
-
-```
 
@@
 
-## Kamala Harris
-
-We begin by modeling the percentage of support for Kamala Harris using a linear regression model.
-The predictors include the sample size, pollster ratings (e.g., pollscore and transparency score), and state.
-This model aims to quantify how these variables influence her support across different polls.
+To analyze factors influencing candidate support, we use a linear regression model with sample size as the primary predictor.
+This model examines the relationship between sample size and the percentage of support for Kamala Harris and Donald Trump across different polls.
+By including sample size as a predictor, we can assess whether larger or smaller sample sizes impact reported support levels for each candidate.
 
@@
 
-The results of the linear regression for Kamala Harris show that sample size has a statistically insignificant effect on her support.
-This suggests that factors other than sample size, such as pollster methodology or regional biases, may play a more significant role in determining the level of support she receives.
+For Kamala Harris, the linear regression results show that sample size has a statistically insignificant effect on her support percentage.
+This implies that other factors, such as pollster methodology or regional biases, may play a more substantial role in shaping her reported support.
 
-@fig-harris-samplesize-supportpct illustrates the relationship between the sample size and support percentage for Kamala Harris.
+In contrast, the results for Donald Trump indicate a weak but statistically significant negative effect of sample size on his support percentage.
+This suggests that, on average, larger sample sizes tend to show slightly lower support for Trump, though the effect size remains small.
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-harris-samplesize-supportpct
-#| fig-cap: Linear Regression of Percentage vs Sample Size for Kamala Harris
-
-# Plot the relationship for Kamala Harris
-ggplot(harris_data, aes(x = sample_size, y = pct)) +
-  geom_point() +
-  geom_smooth(method = "lm", col = "blue") +
-  labs(x = "Sample Size",
-       y = "Percentage")
-```
-
-## Donald Trump
-
-A similar linear regression model was applied to Donald Trump's data.
-The predictors remain the same, and the goal is to determine the factors that drive support for him.
+Figures @fig-harris-samplesize-supportpct and @fig-trump-samplesize-supportpct illustrate the relationship between sample size and support percentage for Kamala Harris and Donald Trump, respectively.
+These plots provide a visual representation of the trends observed in the linear regression models for each candidate.
 
@@
 
-The results in indicate that sample size has a weak but statistically significant negative effect on support for Trump.
-This suggests that larger polls tend to show slightly lower support for Trump, although the effect size is small.
-
-@fig-trump-samplesize-supportpct shows the relationship between the sample size and support percentage for Donald Trump.
-
 ```{r}
@@
 #| warning: false
-#| label: fig-trump-samplesize-supportpct
-#| fig-cap: Linear Regression of Percentage vs Sample Size for Donald Trump
-
-# Plot the relationship for Trump
-ggplot(trump_data, aes(x = sample_size, y = pct)) +
-  geom_point() +
-  geom_smooth(method = "lm", col = "blue") +
-  labs(x = "Sample Size",
-       y = "Percentage")
 
@@
 To better capture the complexity of voter support, we constructed multiple linear regression (MLR) models for both Kamala Harris and Donald Trump.
-These models incorporate several predictors, including pollster rating (represented by pollscore), transparency score, sample size, and state-level data.
-By accounting for these factors, the MLR models allow us to control for more variables that influence voter support and provide more accurate predictions.
+These models incorporate key predictors, including pollster rating (represented by pollscore), transparency score, sample size, and state-level data.
+By including these variables, the MLR models aim to control for multiple factors influencing voter support and to enhance the accuracy of our predictions.
 
-### Kamala Harris Model
+The MLR model for Kamala Harris indicates that factors such as state and pollster rating significantly impact her predicted level of support across polls.
+The results, shown in @fig-mlr-harris, suggest that Harris's support varies substantially by state and depends on the credibility of the pollster.
+Diagnostic checks, including residual and QQ plots, confirm that the model reasonably meets linear regression assumptions.
 
-The MLR model for Kamala Harris takes into account multiple factors that may impact her support across different polls.
-This includes not just the sample size of the poll, but also how pollster reliability (numeric_grade), transparency of poll data (transparency_score), and state-level polling contribute to predicting voter preferences.
-
-The results in @fig-mlr-harris show that several predictors, including state and pollster rating, significantly impact her predicted percentage of voter support.
-These results suggest that voter preferences for Harris vary widely depending on the state and the credibility of the pollster.
-
-The model diagnostics (such as residual plots, QQ plots) were evaluated to ensure the assumptions of linear regression hold.
-Figure X shows diagnostic plots, which indicate that the model performs reasonably well in terms of residual behavior and normality.
-Similarly, we build an MLR model for Donald Trump.
-
-### Donald Trump Model
-
-Similarly, we constructed an MLR model for Donald Trump using the same set of predictors to assess the factors that influence his support across the country.
-
-The results for Donald Trump in @fig-mlr-trump reveal that state-level variations and pollster transparency play a significant role in explaining his level of support.
-The model for Trump also passes diagnostic checks, as shown in Figure Y, indicating that the assumptions of linearity, independence, and homoscedasticity are reasonably met.
-Check Multi-collinearity using Variance Inflation Factor (VIF)
+Similarly, the MLR model for Donald Trump highlights that state-level differences and pollster transparency play a significant role in explaining his support.
+Figure @fig-mlr-trump illustrates these findings, and diagnostic evaluations confirm that the model assumptions (linearity, independence, and homoscedasticity) are sufficiently met.
 
@@
 
-To further optimize the MLR models, we performed stepwise model selection, which systematically adds or removes predictors to minimize the Akaike Information Criterion (AIC) and improve model fit.
+To refine the multi-linear regression model, I applied a stepwise selection method to optimize the choice of predictor variables.
+Stepwise selection evaluates the model iteratively, adding or removing variables based on predefined criteria (typically AIC) to find a more parsimonious model.
+Using both forward and backward selection, stepwise method in the step() function assesses each predictor's contribution and determines if adding or removing it improves model fit.
+By allowing the function to examine variables in both directions, this process systematically removed non-significant variables that did not contribute meaningfully to the model's explanatory power, leaving only the most impactful predictors in the final refined model.
+The resulting plots, showing the "Residuals vs Fitted" and "Normal Q-Q" diagnostic views, allow for visual inspection of model assumptions and residual patterns, supporting the quality of the refined model.
 
-The stepwise model selection improved the model by retaining the most significant predictors and eliminating those with little explanatory power.
+## National-Level Model Evaluation and Predictive Accuracy
 
-## Final Predictions for Electoral College Votes
+After finalizing the model selection process, we evaluated the predictive performance of our models for Kamala Harris and Donald Trump using national-level polling data.
+Focusing on national polls, we split the dataset into training and test sets, maintaining consistency with a fixed random seed to ensure reproducibility.
+For each candidate, we developed a multiple linear regression model using pollscore and a log-transformed sample_size as predictors, capturing factors pertinent to national-level polling dynamics.
 
-Using the final models for both candidates, we predicted the percentage of support in each state and aggregated the results to simulate the Electoral College outcome.
-The predicted percentage of support for each candidate is used to determine the likely winner in each state.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ccc
-#| fig-cap: ccc
-
-# Final models for Harris and Trump
-mlr_harris_model_final <- readRDS(here::here(("models/mlr_harris_model_final.rds")))
-mlr_trump_model_final <- readRDS(here::here(("models/mlr_trump_model_final.rds")))
-
-# Predict poll percentages for Kamala Harris
-harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
-
-# Predict poll percentages for Donald Trump
-trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
-
-```
+We then trained the models on the training subset and generated predictions for the test subset, evaluating model accuracy using the Root Mean Squared Error (RMSE).
+The RMSE metric quantifies the average prediction error in the test set, offering an indication of each model’s reliability in predicting future national poll outcomes.
+The Harris model yielded an RMSE of 3.12, while the Trump model’s RMSE was 2.40, indicating that both models provide reasonably accurate predictions, with the Trump model demonstrating slightly lower average error in predicting national polling support.
+These results reflect the models’ robustness and highlight their utility for assessing national-level candidate support.
 
@@
 
-To forecast the winner of the 2024 election, we aggregate the predicted percentages of support from our models for each state and calculate the Electoral College votes.
-The candidate with 270 or more Electoral College votes is predicted to win the election.
+To forecast the winner of the 2024 U.S.
+Presidential election between Kamala Harris and Donald Trump, we performed the following series of steps, focusing on the distribution of electoral votes across states based on predicted polling percentages.
+This approach allows for an estimation of each candidate's electoral college support under the U.S.
+voting system.
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ddd
-#| fig-cap: ddd
+1.
+Predicting State-Level Polling Percentages: We used final multiple linear regression models for both candidates, Harris and Trump.
+These models generated predicted polling percentages for each candidate across different states, capturing their expected levels of support based on the poll data used.
 
-# Define electoral votes for each state (and Washington, D.C.)
-electoral_votes <- data.frame(
-  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", 
-            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
-            "Maine CD-1", "Maine CD-2", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
-            "Missouri", "Montana", "Nebraska", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", "Nevada", 
-            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
-            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
-            "District of Columbia"),
-  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 2, 1, 1, 10, 11, 15, 10, 6, 10, 
-                      3, 5, 1, 1, 2, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3)
-)
+2.
+Averaging Predicted Support by State: For both Harris and Trump, we calculated the average predicted polling percentage within each state.
+This step provides a state-level summary of each candidate's support based on our predictions, simplifying comparisons between them in each state.
 
-# Aggregate predictions by state for Harris and Trump
-harris_state_avg <- aggregate(predicted_pct_harris ~ state, data = harris_data, FUN = mean)
-trump_state_avg <- aggregate(predicted_pct_trump ~ state, data = trump_data, FUN = mean)
+3.
+Assigning Electoral Votes: We then merged our state-level predictions with the distribution of electoral votes, as assigned by state in the Electoral College.
+This dataset provided the electoral vote count for each state, including special allocations for districts in Maine and Nebraska, which can split their electoral votes by congressional district.
 
-# Merge predictions for both candidates
-prediction_comparison <- merge(harris_state_avg, trump_state_avg, by = "state", all.x = TRUE)
+4.
+Determining State Winners: Using the averaged predictions, we identified the winner in each state by comparing the predicted support percentages.
+Harris or Trump was deemed the winner in a state if their predicted polling percentage exceeded that of their opponent.
 
-# Merge the electoral votes with the predictions
-prediction_comparison <- merge(prediction_comparison, electoral_votes, by = "state", all.x = TRUE)
+5.
+Aggregating Electoral Votes: Finally, we summed the electoral votes for each candidate across all states where they were predicted to win.
+This yielded total electoral vote counts for both Harris and Trump, which are essential for determining the predicted winner under the U.S.
+system, where 270 electoral votes are required to secure the presidency.
 
-# Determine the winner for each state
-prediction_comparison$winner <- ifelse(prediction_comparison$predicted_pct_harris > prediction_comparison$predicted_pct_trump, "Harris", "Trump")
+6.
+Interpreting the Results: With Harris predicted to receive 216 electoral votes compared to Trump’s 147, Harris demonstrates a stronger position in the Electoral College, though neither candidate reaches the 270-vote threshold required for a decisive victory.
+This outcome suggests that, while a clear electoral win is not forecasted, Harris holds a better chance of winning the election based on her lead in projected electoral votes.
+This lead places her closer to the threshold and may indicate a competitive advantage should undecided states or close races lean in her favor.
 
-# Calculate total electoral votes for Kamala Harris
-harris_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Harris"], na.rm = TRUE)
-
-# Calculate total electoral votes for Donald Trump
-trump_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Trump"], na.rm = TRUE)
-
-# Print the results
-print(paste("Harris Electoral Votes:", harris_electoral_votes))
-print(paste("Trump Electoral Votes:", trump_electoral_votes))
-
-# Determine the predicted winner
-if (harris_electoral_votes >= 270) {
-  print("Kamala Harris is predicted to win the 2024 election.")
-} else if (trump_electoral_votes >= 270) {
-  print("Donald Trump is predicted to win the 2024 election.")
-} else {
-  print("No candidate reached 270 electoral votes.")
-}
-
-```
-
-According to our models, neither candidate secures the 270 Electoral College votes needed to win.
-Kamala Harris is projected to receive 216 electoral votes, while Donald Trump is predicted to win 147 electoral votes.
-However, due to the lack of 270 electoral votes for either candidate, the election remains highly competitive, and further developments may shift the balance.
+This electoral vote aggregation based on state-level predictions provides an interpretation of how polling data (@fig-predicted-vote-pct), analyzed through multiple linear regression models, translates into likely electoral support for each candidate.
+It effectively models potential election outcomes while respecting the structure of the Electoral College​​.
 
@@
 
-The multiple linear regression (MLR) models developed in this paper provide valuable insights into the factors that influence voter support for Kamala Harris and Donald Trump.
-Through the analysis of polling data, several key predictors were identified as significant, including state, pollster rating (pollscore), and transparency score.
+This study set out to forecast the outcome of the 2024 U.S.
+Presidential Election between Kamala Harris and Donald Trump by developing a multiple linear regression (MLR) model based on comprehensive polling data.
+Importantly, the same MLR model was applied to both candidates, ensuring consistency in how predictors influenced the predicted support percentages for each.
 
-The MLR model for Kamala Harris highlights the importance of state-level factors in determining her support.
-The model shows that voter preferences vary significantly across different states, with certain states (e.g., California and Maryland) showing higher support levels, while others (e.g., Indiana and Missouri) show lower support.
-Pollster reliability, captured through numeric grades and transparency scores, also plays an important role, suggesting that voters may respond differently based on the credibility of the pollster.
+The MLR model incorporated several significant predictors of voter support, including state-level factors, pollster ratings (pollscore), transparency scores, and sample size.
+By using the same model for both Harris and Trump, we were able to directly compare how these variables impacted each candidate's predicted support across different states.
 
-Donald Trump Similarly, the MLR model for Donald Trump reveals that state-level variations are critical to understanding his level of support.
-The model indicates that Trump’s support is more stable across certain states, but there are also notable outliers where his support fluctuates.
-Pollster characteristics, such as pollscore and transparency score, significantly impact Trump’s predicted support, reflecting the importance of poll quality in predicting election outcomes.
+For both Kamala Harris and Donald Trump, the model highlighted the substantial impact of state-level variations on their predicted support.
+Certain states showed higher predicted support for Harris, such as California and Maryland, which historically lean Democratic.
+Conversely, states like Indiana and Missouri exhibited higher predicted support for Trump, aligning with their Republican-leaning tendencies.
+The consistency of the model across both candidates allowed for a clear comparison of how state demographics and historical voting patterns influenced voter preferences.
 
-Electoral College Forecast Aggregating state-level predictions into Electoral College votes demonstrates the potential competitiveness of the 2024 U.S.
-Presidential Election.
-According to the predictions, neither Kamala Harris nor Donald Trump currently secures the necessary 270 electoral votes to win.
-Harris is projected to receive 216 electoral votes, while Trump is predicted to receive 147 electoral votes.
-This forecast suggests that the election remains highly uncertain, with several key battleground states likely determining the final outcome.
+Pollster reliability indicators, such as numeric grades and transparency scores, were significant predictors for both candidates.
+This suggests that polls conducted by more credible organizations tended to report more accurate and potentially higher support levels for each candidate.
+The sample size also played a role, although its impact varied slightly between the two candidates.
+By using the same model, we ensured that these predictors were weighted equally in the analysis of both Harris's and Trump's support levels.
+
+When aggregating the state-level predictions to simulate the Electoral College outcomes, the findings suggest a competitive race.
+Harris is projected to receive 216 electoral votes, while Trump is predicted to secure 147 electoral votes.
+Neither candidate reaches the 270-vote threshold required to win the presidency, emphasizing the potential for a closely contested election.
+The remaining electoral votes are likely concentrated in battleground states, where voter preferences are more volatile and could ultimately determine the election outcome.
 
@@
 
-One of the main strengths of this analysis is the incorporation of multiple predictors that allow us to account for various factors influencing voter support.
-By considering state-level data, pollster ratings, and transparency scores, the models provide a more nuanced prediction of voter behavior compared to models that rely solely on national-level polling.
+A primary strength of this analysis lies in the consistent application of the same MLR model to both candidates.
+This approach ensures that the comparison between Harris and Trump is based on the same criteria and that any differences in predicted support are attributable to variations in the data rather than differences in the modeling approach.
 
-The use of stepwise model selection and multicollinearity checks further enhanced the robustness of the models by optimizing the choice of predictors and ensuring that the models do not suffer from unstable estimates caused by correlated predictors.
+By incorporating multiple predictors—including state-level data, pollster ratings, transparency scores, and sample size—the model provides a nuanced understanding of the factors influencing voter preferences.
+This multifaceted approach allows for more accurate predictions compared to models relying solely on national-level polling data or a single set of predictors.
 
-Additionally, the aggregation of state-level predictions into Electoral College outcomes presents a more realistic forecast of the election, as the U.S.
-Presidential election is ultimately decided by electoral votes, not popular votes.
+The use of stepwise model selection and checks for multicollinearity enhanced the robustness of the MLR model.
+Stepwise selection optimized the predictor variables, ensuring that only those contributing significantly to the model were included.
+Checking for multicollinearity through the Variance Inflation Factor (VIF) helped prevent unreliable estimates due to correlated predictors, thereby improving the model's validity for both candidates.
+
+Moreover, translating the predicted support percentages into Electoral College projections adds practical value to the analysis.
+Since the U.S.
+Presidential Election is determined by electoral votes rather than the popular vote, this approach offers a more realistic forecast of the election outcome.
+Using the same model for both candidates ensures that the Electoral College projections are directly comparable.
 
@@
 
-Despite the strengths of the models, there are several limitations that should be addressed:
+Despite the strengths, several limitations need to be acknowledged:
 
-- **Poll Reliability and Sampling Bias:** The accuracy of the predictions depends heavily on the quality of the polling data.
-Although the model accounts for pollster ratings, polling methodologies can still introduce biases, particularly in states with limited polling data.
-Sampling errors and non-response biases could skew the results, especially in smaller states or regions with inconsistent polling coverage.
-- **Static Prediction:** The model provides a static snapshot of voter support based on current polling data, which may not capture the dynamic nature of voter preferences over time.
-As election day approaches, voter preferences may shift due to campaign events, debates, or other external factors.
-Without time-series data, the model may fail to account for these trends.
-- **Unaccounted Variables:** Although the model includes important predictors like state, pollscore, and transparency score, other potentially influential factors, such as economic conditions, campaign spending, and voter turnout, are not included in the analysis.These unaccounted variables may introduce inaccuracies in the final predictions.
+- Reliance on Polling Data Quality: The accuracy of the model is inherently dependent on the quality of the polling data used.
+While the analysis accounted for pollster ratings and transparency scores, inherent biases and methodological differences between polls could still affect the results.
+Sampling errors, non-response biases, and underrepresentation of certain demographic groups are persistent issues in polling data that can skew predictions for both candidates.
+
+- Static Snapshot of Voter Preferences: The model provides a static view based on the most recent polling data available.
+It does not account for the dynamic nature of election campaigns, where voter preferences can shift rapidly due to various factors such as political events, debates, or emerging issues.
+This limitation means the model may not accurately predict future changes in voter sentiment leading up to the election for either candidate.
+
+- Exclusion of Other Influential Factors: The analysis did not incorporate other variables that can significantly impact election outcomes, such as economic indicators (e.g., unemployment rates, GDP growth), social movements, campaign strategies, or voter turnout efforts.
+The absence of these factors may limit the model's ability to fully capture the complexities of voter behavior for both Harris and Trump.
+
+- Electoral College Simplifications: The Electoral College simulation assumes that the candidate with the higher predicted support in a state wins all its electoral votes (except for Maine and Nebraska).
+This winner-takes-all approach does not account for the proportional allocation of electoral votes in those two states or potential variations in voter turnout that could affect the actual results.
 
 ## Future Research 
-Directions Future work could improve upon this analysis by addressing some of the limitations mentioned above.
-For instance, incorporating time-series data could allow the model to capture how voter preferences evolve in response to external factors such as campaign events, economic developments, and political endorsements.
-A dynamic forecasting model that updates predictions as new polls are released would provide more timely and accurate forecasts.
+To enhance the predictive power and accuracy of election forecasting models, future research could consider the following approaches:
 
-Moreover, integrating other influential variables, such as economic indicators (e.g., unemployment rates, inflation), voter turnout models, and campaign spending data, could enhance the predictive power of the model.
-Including demographic data (e.g., age, education, income) could also improve the granularity of predictions, especially in battleground states where demographic shifts are critical to election outcomes.
+- Incorporate Time-Series Analysis: Introducing a time-series component would allow the model to account for trends and shifts in voter preferences over time.
+This dynamic approach could capture the impact of events such as debates, policy announcements, or external factors on voter sentiment for both candidates.
 
-Finally, expanding the model to account for ranked-choice voting in certain states could provide a more accurate forecast in scenarios where third-party candidates or run-off elections play a significant role.
+- Expand Predictive Variables: Including additional variables such as economic indicators, demographic data, and social media sentiment analysis could provide a more comprehensive view of the factors influencing voter behavior.
+Integrating data on unemployment rates or consumer confidence might reveal economic influences on candidate support.
+
+- Voter Turnout Models: Developing models to predict voter turnout could significantly improve election forecasts.
+Turnout can vary widely between elections and is often influenced by voter enthusiasm, mobilization efforts, and barriers to voting.
+Incorporating turnout probabilities could adjust the predicted support percentages to reflect more realistic electoral scenarios.
+
+- Advanced Modeling Techniques: Exploring advanced statistical methods or machine learning algorithms might capture complex nonlinear relationships between predictors and voter support.
+Techniques such as random forests, gradient boosting, or neural networks could uncover patterns not detectable through linear regression.
+
+- Account for Electoral Nuances: With variations in how electoral votes are allocated and the potential impact of third-party candidates or ranked-choice voting in some states, future models should consider these electoral nuances.
+Simulating different voting systems could provide insights into how alternative methods might affect election outcomes.
+
+- Cross-Validation with Alternative Data Sources: Validating the model against alternative data sources, such as exit polls or historical voting patterns, could test its robustness.
+Cross-validation helps ensure that the model is not overfitted to the polling data used and can generalize to different datasets.
 
@@
 #| warning: false
+#| label: tbl-overview_data
+
+data_overview <- data %>%
+  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
+
+# Transpose the first 5 rows to display them vertically
+data_overview_transposed <- data_overview %>%
+  head(5) %>%
+  t()
+
+# Display the transposed table with a caption
+data_overview_transposed %>%
+  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-harris-samplesize-supportpct
+#| fig-cap: Linear Regression of Percentage vs Sample Size for Kamala Harris
+
+# Plot the relationship for Kamala Harris
+ggplot(harris_data, aes(x = sample_size, y = pct)) +
+  geom_point() +
+  geom_smooth(method = "lm", col = "blue") +
+  labs(x = "Sample Size",
+       y = "Percentage")
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-trump-samplesize-supportpct
+#| fig-cap: Linear Regression of Percentage vs Sample Size for Donald Trump
+
+# Plot the relationship for Trump
+ggplot(trump_data, aes(x = sample_size, y = pct)) +
+  geom_point() +
+  geom_smooth(method = "lm", col = "blue") +
+  labs(x = "Sample Size",
+       y = "Percentage")
+```
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
 #| label: fig-mlr-harris
@@
 
+```{r, out.width="90%", out.height="70%", fig.align="center"}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-predicted-vote-pct
+#| fig-cap: Predicted Vote Percentages by State
+knitr::include_graphics(here::here(("improved_vote_percentages_by_state.png")))
+
+```
+
 \newpage

### 2024-10-31 paper/paper.qmd

@@
 ## Overview
+Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news.
+This database compiles and standardizes polling results from various organizations, applying quality metrics and assessments to each poll.
+Several key variables from this dataset are crucial to our analysis:
 
@@
 ggplot(data, aes(x = pct, fill = candidate_name)) +
-  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
+  geom_histogram(bins = 30, alpha = 0.8, position = "identity") +
   facet_wrap(~ candidate_name) +
@@
   theme(legend.position = "bottom")
-
 ```

### 2024-11-01 scripts/00-simulate_data.R
Improved Sim and Testing

@@ -1,7 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of US pollster polling outcomes
+# Purpose: Simulates polling data for the 2024 US presidential election between
+# Kamala Harris and Donald Trump, incorporating realistic polling features
 # Authors: Tim Chen, Steven Li, Tommy Fu
-# Date: 18 October 2024
+# Date: 31 October 2024
 # Contacts: 
 # - Tim Chen: timwt.chen@mail.utoronto.ca
 # - Steven Li: stevency.li@mail.utoronto.ca
@@ -13,83 +14,123 @@
 #### Workspace setup ####
 library(tidyverse)
 library(lubridate)
+set.seed(2024) # For reproducibility
 
-#### Data expectations ####
-# Pollster Data:
-# - Columns: poll_id, pollster, pollster_rating, method, state, voter_type, party, winner, percentage
-# - poll_id should be unique
-# - Date should be between 2024-01-01 and 2024-10-18
-# - pollster_rating should be between 1 and 5
-# - method should be one of: "Online", "Phone", "Mixed"
-# - state should be valid US state abbreviations
-# - voter_type should be one of: "Likely Voters", "Registered Voters", "All Adults"
-# - party should be one of: "Republican", "Democrat"
-# - winner should be one of: "Trump", "Harris"
-# - percentage should be between 0 and 100
+######################## Define simulation parameters ########################
+# Key polling parameters based on historical patterns
+BASE_HARRIS_SUPPORT <- 51.0  # Starting point for Harris support
+POLL_ERROR <- 2.5           # Standard deviation for polling errors
+DEM_BIAS <- -1.0           # Historical pro-Democrat bias in polls
 
-set.seed(2024)
-
-########################### Simulate Pollster Data ############################ 
-num_polls <- 1000
-
-pollster_data <- tibble(
-  poll_id = seq(1, num_polls),
-  date = sample(seq(as.Date('2024-01-01'), as.Date('2024-10-18'), by="day"), 
-                num_polls, replace=TRUE),
-  pollster = sample(c("PollsterA", "PollsterB", "PollsterC", 
-                      "PollsterD", "PollsterE"), num_polls, replace=TRUE),
-  pollster_rating = round(runif(num_polls, 1, 5), 1),
-  method = sample(c("Online", "Phone", "Mixed"), num_polls, replace=TRUE),
[…]

### 2024-11-01 scripts/01-test_simulated_data.R

@@ -1,90 +1,110 @@
 #### Preamble ####
 # Purpose: Testing simulated pollster data for the 2024 US presidential election
 # Authors: Tim Chen, Steven Li, Tommy Fu
-# Date: 18 October 2024
+# Date: 31 October 2024
 # Contacts: 
 # - Tim Chen: timwt.chen@mail.utoronto.ca
 # - Steven Li: stevency.li@mail.utoronto.ca
 # - Tommy Fu: tommy.fu@mail.utoronto.ca
 # Pre-requisites: 
 # - 00-simulate_data.R must have been run
-# - The `tidyverse`, `lubridate`, and `testthat` package must be installed and loaded
+# - The `tidyverse`, `lubridate`, and `testthat` packages must be installed
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
 library(lubridate)
 
-# Read in the data
-pollster_data <- read_csv("data/00-simulated_data/simulated_pollster_data.csv")
+# Read in the simulated data
+pollster_data <- read_csv(
+  here::here("data/00-simulated_data/simulated_pollster_data.csv"))
 
-########################### Tests for Pollster Data########################### 
-
-test_that("Pollster data has correct structure", {
-  expected_cols <- c("poll_id", "date", "pollster", "pollster_rating", "method", 
-                     "state", "voter_type", "party", "winner", "percentage")
-  expect_equal(colnames(pollster_data), expected_cols)
-  expect_true(is.numeric(pollster_data$poll_id))
-  expect_true(inherits(pollster_data$date, "Date"))
-  expect_true(is.character(pollster_data$pollster))
-  expect_true(is.numeric(pollster_data$pollster_rating))
-  expect_true(is.character(pollster_data$method))
-  expect_true(is.character(pollster_data$state))
-  expect_true(is.character(pollster_data$voter_type))
-  expect_true(is.character(pollster_data$party))
-  expect_true(is.character(pollster_data$winner))
-  expect_true(is.numeric(pollster_data$percentage))
-})
-
-test_that("Pollster data has valid values", {
-  expect_true(all(!is.na(pollster_data)))
-  expect_true(all(pollster_data$date >= as.Date("2024-01-01") & 
-                    pollster_data$date <= as.Date("2024-10-18")))
-  expect_true(all(pollster_data$pollster_rating >= 1 & pollster_data$pollster_rating <= 5))
[…]

### 2024-11-01 scripts/04-test_analysis_data.R

@@ -1,5 +1,6 @@
 #### Preamble ####
-# Purpose: Testing cleaned pollster data for the 2024 US presidential election
+# Purpose: Testing cleaned pollster data for the 2024 US presidential election,
+#          ensuring data quality and consistency with expected formats
 # Authors: Tim Chen, Steven Li, Tommy Fu
 # Date: 18 October 2024
 # Contacts: 
@@ -10,78 +11,153 @@
 # - 03-clean_data.R must have been run
 # - The `tidyverse`, `lubridate`, and `testthat` package must be installed and loaded
 
+
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
 library(lubridate)
 
 # Read in the data
-pollster_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+pollster_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
-########################### Tests for Pollster Data########################### 
-
-test_that("Pollster data has correct structure", {
+############################ Test Data Structure ############################
+test_that("Data has correct structure and identifiers", {
+  # Test for required columns
   expected_cols <- c("poll_id", "pollster_id", "pollster", "numeric_grade",
                      "pollscore", "methodology", "transparency_score", 
                      "sample_size", "population", "population_full", "party", 
                      "answer", "pct", "state", "candidate_name", "end_date", 
                      "candidate_binary", "num_supporters")
   expect_equal(colnames(pollster_data), expected_cols)
+  
+  # Test relationship between poll_id and pollster_id
+  expect_true(all(!is.na(pollster_data$poll_id)))
+  expect_true(all(!is.na(pollster_data$pollster_id)))
 })
 
-test_that("Data types are correct", {
-  expect_true(is.numeric(pollster_data$poll_id))
-  expect_true(is.numeric(pollster_data$pollster_id))
-  expect_true(is.character(pollster_data$pollster))
-  expect_true(is.numeric(pollster_data$numeric_grade))
-  expect_true(is.numeric(pollster_data$pollscore))
-  expect_true(is.character(pollster_data$methodology))
-  expect_true(is.numeric(pollster_data$transparency_score))
-  expect_true(is.numeric(pollster_data$sample_size))
-  expect_true(is.character(pollster_data$population))
-  expect_true(is.character(pollster_data$population_full))
-  expect_true(is.character(pollster_data$party))
-  expect_true(is.character(pollster_data$answer))
-  expect_true(is.numeric(pollster_data$pct))
[…]

### 2024-11-01 scripts/07-replications.R

@@ -1,105 +1,0 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-# Final models for Harris and Trump
-mlr_harris_model_final <- readRDS(here::here(("models/mlr_harris_model_final.rds")))
-mlr_trump_model_final <- readRDS(here::here(("models/mlr_trump_model_final.rds")))
-
-# Predict poll percentages for Kamala Harris
-harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
-
-# Predict poll percentages for Donald Trump
-trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
-# Define electoral votes for each state (and Washington, D.C.)
-electoral_votes <- data.frame(
-  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", 
-            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
-            "Maine CD-1", "Maine CD-2", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
-            "Missouri", "Montana", "Nebraska", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", "Nevada", 
-            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
-            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
-            "District of Columbia"),
[…]

### 2024-11-01 paper/paper.qmd
testing round
Improved data section, added figures and tables to appendix

@@
 # Data {#sec-data}
-
 ## Overview
 Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news.
-This database compiles and standardizes polling results from various organizations, applying quality metrics and assessments to each poll.
-Several key variables from this dataset are crucial to our analysis:
+This database aggregates and standardizes polling results from various organizations, applying quality metrics and methodological assessments to each poll.
+In this section, we detail our selected variables, discuss key measurements, outline important limitations of our data, and our data cleaning process.
 
-The dataset employed in this analysis encompasses polling data from various reputable sources, capturing critical variables related to polling organizations, sample sizes, candidate support levels, and polling coverage (state-level or national).
-The primary variables of interest are outlined as follows:
+Our analysis focuses on several key variables that directly influence polling accuracy and electoral predictions. @tbl-overview_data presents a sample of our dataset.
 
-- **Poll ID**: A unique identifier for each poll entry, facilitating data tracking and management (e.g., 88590).
-- **Pollster**: The organization conducting the poll, which serves as a primary indicator of the poll’s methodological quality (e.g., YouGov, Ipsos).
-- **Pollscore**: A quantitative measure of the pollster’s reliability, where lower (negative) values suggest higher predictive accuracy (e.g., -1.1 for YouGov).
-- **Sample Size**: The number of respondents included in each poll, influencing the statistical precision and margin of error (e.g., 1414).
-- **Support Percentage (pct)**: The proportion of respondents expressing support for each candidate, serving as the dependent variable in subsequent analyses (e.g., 47% for Kamala Harris).
-- **State**: Indicates the geographical focus of the poll, either at the state level or nationwide (e.g., National).
-- **Candidate Name**: The candidate assessed in the poll, providing context for support levels and enabling candidate-specific analyses (e.g., Kamala Harris, Donald Trump).
-- **End Date**: The date on which the poll concluded, offering temporal alignment for longitudinal analyses (e.g., 2024-10-07).
+## Poll Quality Variables
 
-To ensure analytical rigor, the dataset underwent extensive preprocessing.
-This involved standardizing variable formats for compatibility with regression models, converting support percentages into numerical values, and systematically addressing missing data to minimize bias.
-These preprocessing steps ensured that the dataset adhered to the standards of a "tidy" dataset, enabling robust model development and enhancing interpretability of subsequent statistical analyses.
+@tbl-poll-metrics summarizes three primary metrics that assess poll quality in our dataset:
 
-@tbl-overview_data shows a sample of the dataset.
+- **Pollscore**: Represents "the error and bias attributable to a pollster, where negative numbers indicate better performance." Our data shows a mean pollscore of -1.06 (SD = 0.28), with values ranging from -1.50 to -0.50, indicating generally high-quality polling organizations in our sample.
+- **Numeric Grade**: A rating given to pollsters to indicate their quality and reliability.
+With a mean of 2.89 (SD = 0.10) and a range of 2.70 to 3.00, our dataset maintains high standards.
+We established a minimum threshold of 2.7 to ensure methodological rigor.
+- **Transparency Score**: A measure reflecting pollsters' transparency about their methodology, with 10 being the highest possible score.
+The data shows strong overall transparency with a mean of 8.59 (SD = 1.04), indicating that most polls in our sample maintain high standards of methodological disclosure.
+
+## Methodological Variables
+
+- **Methodology**: @fig-methodology-dist shows the distribution of polling approaches used to conduct the polls.
+Live Phone methods and Online Panels are the predominant methods, with hybrid approaches combining multiple methods (such as IVR/Online Panel/Text-to-Web) representing evolving polling techniques adapting to changing communication patterns.
+- **Sample Size**: @fig-sample-size illustrates the distribution of poll sample sizes, representing the total number of respondents participating in each poll.
+Our dataset shows considerable variation, with a mean of 1,114.76 respondents (SD = 583.72).
+The right-skewed distribution ranges from 450 to 4,253 respondents, with larger sample sizes typically associated with national polls.
+
+## Geographic and Temporal Variables
+
+- **State Coverage**: @tbl-state-polls reveals the distribution of polls across different U.S.
+states.
+National polls lead with 261 surveys, followed by concentrated polling in key battleground states: Pennsylvania (86), Wisconsin (82), and North Carolina (64).
+This distribution reflects strategic focus on states where the electoral outcome is less certain.
+- **End Date**: @fig-temporal-polling tracks polling frequency from July through October 2024.
+We observe increased activity during September and early October, with weekly poll counts peaking in late September.
+This pattern reflects intensified polling efforts as the election approaches.
+
+## Outcome Variable
+
+- **Support Percentage**: @fig-candidate-support displays our primary dependent variable: the percentage of respondents expressing support for each candidate.
+The distribution reveals a competitive race between Donald Trump and Kamala Harris, with support levels typically ranging between 40-50% for both candidates.
 
@@
 
-- **Poll Quality:** While we filter for high-quality polls using numeric grades, differences in polling methodologies may introduce systematic biases.
-The inclusion of pollster ratings helps account for historical accuracy but cannot completely eliminate these potential biases.
-- **Temporal Dynamics:** Our dataset provides a snapshot of voter preferences during a specific timeframe.
-This static nature means we cannot capture the full dynamics of voter preference evolution over the campaign period.
+- **Poll Quality:** While our pollscore and numeric grade filters help ensure data quality, these metrics are based on historical performance and may not fully capture current methodological improvements or deterioration.
+- **Temporal Dynamics:** Our dataset provides discrete snapshots of voter preferences rather than continuous measurement.
+This limitation is particularly relevant given the rapid evolution of political narratives and voter sentiment during presidential campaigns.
 - **Geographic Coverage:** Although we have national and state-level polling data, coverage varies by state.
@@
 - **Response Bias:** Despite careful methodology by pollsters, self-selection bias in survey participation and social desirability bias in responses remain potential concerns.
-
-## Outcome variables
-Our primary outcome variable is the percentage of support (*pct*) for Kamala Harris and Donald Trump in each poll.
-This measurement represents the proportion of respondents who indicate they would vote for each candidate if the election were held on the day of the poll.
-The variable directly captures voter preferences and serves as the foundation for our electoral predictions.
-
-@fig-support below illustrates the distribution of support for both candidates, revealing several notable patterns.
-First, the support percentages cluster between 40% and 60%, reflecting the competitive nature of the race.
-Second, the distributions show slight differences between candidates, with Harris's support displaying more variation than Trump's.
-This pattern might reflect differences in voter certainty or polling methodology across different states and time periods.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-support
-#| fig-cap: Distribution of Support for Kamala Harris and Donald Trump
-
-ggplot(data, aes(x = pct, fill = candidate_name)) +
-  geom_histogram(bins = 30, alpha = 0.8, position = "identity") +
-  facet_wrap(~ candidate_name) +
-  labs(x = "Percentage of Support",
-       y = "Count") +
-  scale_fill_discrete(name = "Candidates") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-```
-
-## Predictor variables
-
-Our model incorporates several key predictor variables, each chosen for its theoretical importance in explaining polling variations and electoral outcomes:
-
-- **Numeric Grade:** A composite measure (scale: 0-4) incorporating factors such as methodology rigor and historical accuracy
-- **Pollscore:** A measure of historical polling accuracy (range: -4 to +4, where negative scores indicate better performance)
-- **Transparency Score:** Quantifies the openness of polling methodology (scale: 0-10)
-- **Sample Size:** Number of respondents, typically ranging from 500 to 3000
-- **State:** State-level indicators capturing regional political variations
-- **Methodology:** Survey approach (e.g., online panel, phone interviews)
-
-The relationship between these predictors and polling outcomes is complex and often interconnected.
-For example, while larger sample sizes generally reduce sampling error, this effect may be moderated by the poll's methodology and quality metrics.
-Similarly, geographic variations in polling accuracy suggest that the relationship between predictors and outcomes may vary systematically across states.
-
-These variables were selected based on both theoretical foundations in electoral polling literature and practical considerations of data availability and quality.
-We focused on these core variables due to their consistent availability across polls and demonstrated importance in previous electoral forecasting efforts.
-
-**Edit note: Add a table in appendix to show sample with all variables used in model**
 
@@
 
-  > i. Converted end dates to standardized date format
+  > i. Converted *End Date* values to standardized date format
   > ii. Calculated absolute supporter numbers from percentages and sample sizes
   > iii. Created binary candidate indicators (Harris = 1, Trump = 0)
+
+Other variable such as ... were not selected for our analysis because ...
 
@@
 #| label: tbl-overview_data
+#| tbl-cap: Sample Overview of Selected Variables in the Polling Dataset
+
+set.seed(778) # For reproducability
 
 data_overview <- data %>%
-  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
+  select(pollster, pollscore, numeric_grade, transparency_score, sample_size, 
+         methodology, pct, state, candidate_name, end_date)
 
 # Transpose the first 5 rows to display them vertically
-data_overview_transposed <- data_overview %>%
-  head(5) %>%
+data_overview <- data_overview %>%
+  sample_n(5) %>%
   t()
 
-# Display the transposed table with a caption
-data_overview_transposed %>%
-  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
+# Display the transposed table
+kable(data_overview)
+```
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-poll-metrics
+#| tbl-cap: Summary Statistics of Key Poll Metrics
+
+data %>%
+  summarise(
+    across(c(sample_size, pollscore, numeric_grade, transparency_score),
+           list(
+             Mean = mean,
+             Median = median,
+             SD = sd,
+             Min = min,
+             Max = max
+           ), 
+           .names = "{.fn}_{.col}",
+           na.rm = TRUE)
+  ) %>%
+  pivot_longer(everything(), 
+               names_to = c("Statistic", "Variable"),
+               names_sep = "_",
+               values_to = "Value") %>%
+  pivot_wider(names_from = Variable,
+              values_from = Value) %>%
+  mutate(across(-Statistic, ~round(.x, 2))) %>%
+  kable()
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-methodology-dist
+#| fig-cap: Distribution of Polling Methodologies
+
+data %>%
+  ggplot(aes(x = fct_infreq(methodology))) +
+  geom_bar() +
+  coord_flip() +
+  theme_minimal() +
+  labs(x = "Methodology",
+       y = "Count") +
+  theme(axis.text.y = element_text(size = 8))
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-candidate-support
+#| fig-cap: Distribution of Support Percentages by Candidate
+
+ggplot(data, aes(x = pct, fill = candidate_name)) +
+  geom_histogram(bins = 30, position = "identity") +
+  facet_wrap(~ candidate_name) +
+  labs(x = "Percentage of Support",
+       y = "Count") +
+  scale_fill_discrete(name = "Candidates") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-state-polls
+#| tbl-cap: Polling Frequency by State
+
+data %>%
+  count(state) %>%
+  arrange(desc(n)) %>%
+  rename("State" = state,
+         "Number of Polls" = n) %>%
+  kable(format = "latex", booktabs = TRUE)
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-temporal-polling
+#| fig-cap: Number of Polls Over Time
+
+data %>%
+  mutate(end_date = as.Date(end_date)) %>%
+  mutate(week = floor_date(end_date, "week")) %>%
+  count(week) %>%
+  ggplot(aes(x = week, y = n)) +
+  geom_line() +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Week",
+       y = "Number of Polls") +
+  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d")
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-sample-size
+#| fig-cap: Distribution of Poll Sample Sizes
+
+data %>%
+  ggplot(aes(x = sample_size)) +
+  geom_histogram(binwidth = 100) +
+  theme_minimal() +
+  labs(x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(breaks = seq(0, max(data$sample_size), by = 500))
 ```
@@
 #| fig-cap: Predicted Vote Percentages by State
-knitr::include_graphics(here::here(("improved_vote_percentages_by_state.png")))
+# knitr::include_graphics(here::here(("improved_vote_percentages_by_state.png")))
 

### 2024-11-02 paper/paper.qmd
model refinement with discussion refinement

@@
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+# Load necessary libraries
+library(reshape2)
+library(ggplot2)
+
+# Load final models for Harris and Trump
+mlr_harris_model_final <- readRDS(here::here("models/mlr_harris_model_final.rds"))
+mlr_trump_model_final <- readRDS(here::here("models/mlr_trump_model_final.rds"))
+
+# Predict poll percentages for Kamala Harris and Donald Trump
+harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
+trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+# Define electoral votes for each state (and Washington, D.C.)
+# Correct the electoral_votes data frame
+electoral_votes <- data.frame(
+  state = c(
+    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", 
+    "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
+    "Kansas", "Kentucky", "Louisiana", 
+    # Maine entries
+    "Maine At-Large", "Maine CD-1", "Maine CD-2", 
+    "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", 
+    # Nebraska entries
+    "Nebraska At-Large", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", 
+    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", 
+    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", 
+    "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", 
+    "Wyoming"
+  ),
+  electoral_votes = c(
+    9, 3, 11, 6, 55, 9, 7, 3, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8,
+    # Maine electoral votes
+    2, 1, 1,
+    10, 11, 16, 10, 6, 10, 3,
+    # Nebraska electoral votes
+    2, 1, 1, 1,
+    6, 4, 14, 5, 29, 15, 3,
+    18, 7, 7, 20, 4, 9, 3,
+    11, 38, 6, 3, 13, 12, 5, 10,
+    3
+  )
+)
+
+# Aggregate predictions by state for Harris and Trump
+harris_state_avg <- aggregate(predicted_pct_harris ~ state, data = harris_data, FUN = mean)
+trump_state_avg <- aggregate(predicted_pct_trump ~ state, data = trump_data, FUN = mean)
+
+# Merge predictions for both candidates
+prediction_comparison <- merge(harris_state_avg, trump_state_avg, by = "state", all.x = TRUE)
+
+```
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+# Identify any states missing from the prediction data
+missing_states <- setdiff(electoral_votes$state, prediction_comparison$state)
+
+# Assign 50% support to both candidates for missing states
+missing_states_data <- data.frame(
+  state = missing_states,
+  predicted_pct_harris = 50,
+  predicted_pct_trump = 50
+)
+
+# Append missing states data to prediction_comparison
+prediction_comparison <- rbind(prediction_comparison, missing_states_data)
+
+# Merge electoral votes into the prediction_comparison to ensure each state has electoral votes
+prediction_comparison <- merge(prediction_comparison, electoral_votes, by = "state", all.x = TRUE)
+
+```
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+# Determine the winner for each state based on predicted percentages
+set.seed(304)
+prediction_comparison$winner <- ifelse(
+  prediction_comparison$predicted_pct_harris > prediction_comparison$predicted_pct_trump, 
+  "Harris", 
+  ifelse(prediction_comparison$predicted_pct_trump > prediction_comparison$predicted_pct_harris, 
+         "Trump", 
+         sample(c("Harris", "Trump"), nrow(prediction_comparison), replace = TRUE))
+)
+
+# Calculate total electoral votes for each candidate
+harris_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Harris"], na.rm = TRUE)
+trump_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Trump"], na.rm = TRUE)
+
+# Create a table with electoral votes
+electoral_votes_table <- data.frame(
+  Candidate = c("Harris", "Trump"),
+  Electoral_Votes = c(harris_electoral_votes, trump_electoral_votes)
+)
+
+# Display the table with kable
+kable(electoral_votes_table, caption = "Electoral Votes for Each Candidate", col.names = c("Candidate", "Electoral Votes"))
+```
+
 # Electoral College Prediction {#sec-predict}
 
-To forecast the winner of the 2024 U.S.
-Presidential election between Kamala Harris and Donald Trump, we performed the following series of steps, focusing on the distribution of electoral votes across states based on predicted polling percentages.
-This approach allows for an estimation of each candidate's electoral college support under the U.S.
-voting system.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-predicted-vote-pct
+#| fig.height: 5
+#| fig.cap: "Predicted Vote Percentages by State"
+
+# Load necessary libraries
+library(ggplot2)
+library(reshape2)
+library(dplyr)
+library(patchwork)
+
+# Melt data for easier plotting
+prediction_melted <- melt(
+  prediction_comparison, 
+  id.vars = c("state", "electoral_votes", "winner"),
+  measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
+)
+
+# Rename melted variables for readability
+colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
+prediction_melted$candidate <- ifelse(
+  prediction_melted$candidate == "predicted_pct_harris", 
+  "Harris", 
+  "Trump"
+)
+
+# Calculate the midpoint to split the states
+n_states <- length(unique(prediction_melted$state))
+midpoint <- ceiling(n_states / 2)
+
+# Order states by electoral votes in descending order for better visualization
+states_order <- prediction_melted %>%
+  distinct(state, electoral_votes) %>%
+  arrange(desc(electoral_votes)) %>%
+  pull(state)
+
+# Split the states into two groups
+left_states <- states_order[1:midpoint]
+right_states <- states_order[(midpoint + 1):length(states_order)]
+
+# Create two separate datasets
+prediction_left <- prediction_melted %>%
+  filter(state %in% left_states) %>%
+  mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
+
+prediction_right <- prediction_melted %>%
+  filter(state %in% right_states) %>%
+  mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
+
+# Function to create individual plots to avoid repetition
+create_plot <- function(data, show_legend = FALSE) {
+  ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
+    geom_bar(
+      stat = "identity", 
+      position = position_dodge(width = 0.9), 
+      width = 0.5  # Reduced bar width for more spacing
+    ) +
+    coord_flip() +
+    labs(
+      x = NULL,
+      y = "Predicted Vote Percentage (%)"
+    ) +
+    scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
+    theme_minimal() +
+    theme(
+      text = element_text(size = 9),  # Reduced overall text size
+      axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
+      axis.text.x = element_text(size = 9),
+      legend.title = element_blank(),
+      legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
+      legend.text = element_text(size = 9),
+      panel.grid.major.y = element_blank(),
+      plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
+    ) +
+    geom_text(
+      aes(label = round(value, 1)),
+      position = position_dodge(width = 0.9),
+      hjust = -0.2,    # Adjusted horizontal position for labels
+      size = 1.5,      # Smaller label size
+      show.legend = FALSE
+    ) +
+    scale_y_continuous(
+      expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
+      limits = c(0, 75)  # Adjust based on your data
+    )
+}
+
+# Create the left plot with legend
+plot_left <- create_plot(prediction_left, show_legend = TRUE)
+
+# Create the right plot without legend
+plot_right <- create_plot(prediction_right, show_legend = FALSE)
+
+# Combine the plots side by side with legend on top
+combined_plot <- plot_left + plot_right +
+  plot_layout(guides = "collect") &  # Collects the legend at the top
+  theme(legend.position = "top")  # Places the combined legend at the top
+
+# Print the combined plot
+combined_plot
+
+```
+
+To project the outcome of the 2024 U.S.
+Presidential election between Kamala Harris and Donald Trump, we conducted a statistical analysis that integrates predicted state-level polling percentages with the allocation of electoral votes.
+This methodology employs multiple linear regression models to estimate each candidate's support and translates these predictions into potential Electoral College results, adhering to the structure of the U.S.
+electoral system.
 
 1.
-Predicting State-Level Polling Percentages: We used final multiple linear regression models for both candidates, Harris and Trump.
-These models generated predicted polling percentages for each candidate across different states, capturing their expected levels of support based on the poll data used.
+**Prediction of State-Level Polling Percentages**: We utilized finalized multiple linear regression models for both Kamala Harris and Donald Trump.
+These models were applied to their respective datasets to generate predicted polling percentages for each candidate across various states.
+The predictions reflect the expected levels of support based on historical polling data and relevant covariates included in the models.
 
 2.
-Averaging Predicted Support by State: For both Harris and Trump, we calculated the average predicted polling percentage within each state.
-This step provides a state-level summary of each candidate's support based on our predictions, simplifying comparisons between them in each state.
+**Aggregation of Predicted Support by State**: For each candidate, we computed the average predicted polling percentage within each state.
+This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see Figure @fig-predicted-vote-pct).
 
 3.
-Assigning Electoral Votes: We then merged our state-level predictions with the distribution of electoral votes, as assigned by state in the Electoral College.
-This dataset provided the electoral vote count for each state, including special allocations for districts in Maine and Nebraska, which can split their electoral votes by congressional district.
+**Integration of Electoral Vote Allocations**: We established a dataset detailing the electoral vote distribution for each state, including the distinct allocations for Maine and Nebraska, which can split their electoral votes by congressional district.
+This dataset was merged with the state-level predicted support data to align electoral votes with the corresponding predicted support.
 
 4.
-Determining State Winners: Using the averaged predictions, we identified the winner in each state by comparing the predicted support percentages.
-Harris or Trump was deemed the winner in a state if their predicted polling percentage exceeded that of their opponent.
+**Addressing Missing Data**: We identified states lacking predicted polling data and assigned a neutral predicted support of 50% to both candidates in these states.
+This assumption accounts for the absence of data while ensuring that all states are included in the analysis.
 
 5.
-Aggregating Electoral Votes: Finally, we summed the electoral votes for each candidate across all states where they were predicted to win.
-This yielded total electoral vote counts for both Harris and Trump, which are essential for determining the predicted winner under the U.S.
-system, where 270 electoral votes are required to secure the presidency.
+**Determination of State Winners**: Based on the aggregated predicted percentages, we determined the projected winner in each state.
+A candidate was designated as the winner if their predicted polling percentage exceeded that of their opponent.
+In cases where the predicted percentages were equal, we randomly assigned the winner using a stochastic approach to reflect the uncertainty inherent in tied predictions.
 
 6.
-Interpreting the Results: With Harris predicted to receive 216 electoral votes compared to Trump’s 147, Harris demonstrates a stronger position in the Electoral College, though neither candidate reaches the 270-vote threshold required for a decisive victory.
-This outcome suggests that, while a clear electoral win is not forecasted, Harris holds a better chance of winning the election based on her lead in projected electoral votes.
-This lead places her closer to the threshold and may indicate a competitive advantage should undecided states or close races lean in her favor.
+**Calculation of Total Electoral Votes**: We summed the electoral votes from all states won by each candidate to compute their total electoral vote counts.
+This calculation is critical, as securing at least 270 electoral votes is necessary to win the presidency under the U.S.
+electoral system.
 
-This electoral vote aggregation based on state-level predictions provides an interpretation of how polling data (@fig-predicted-vote-pct), analyzed through multiple linear regression models, translates into likely electoral support for each candidate.
-It effectively models potential election outcomes while respecting the structure of the Electoral College​​.
+7.
+**Interpretation of Results**: The analysis indicated that Kamala Harris is projected to receive a total of **306 electoral votes**, while Donald Trump is projected to receive **232 electoral votes**.
+Harris surpasses the 270-vote threshold required for victory, suggesting a strong position in the Electoral College based on our predictions.
+This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes.
+
+This statistical approach effectively bridges the gap between state-level polling data and the Electoral College outcome.
+By employing multiple linear regression models to predict polling percentages and integrating these predictions with the electoral vote framework, we provide a nuanced analysis of potential election results.
+This methodology respects the intricacies of the U.S.
+electoral system and offers valuable insights into how predicted voter support may translate into electoral success.
 
@@
 
-## Key Findings
+## Summary of Findings
 
-This study set out to forecast the outcome of the 2024 U.S.
-Presidential Election between Kamala Harris and Donald Trump by developing a multiple linear regression (MLR) model based on comprehensive polling data.
-Importantly, the same MLR model was applied to both candidates, ensuring consistency in how predictors influenced the predicted support percentages for each.
+In this paper, we developed statistical models to forecast the outcome of the 2024 United States presidential election between Kamala Harris and Donald Trump.
+Utilizing polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database, we analyzed factors influencing candidate support through linear and multiple linear regression models.
+We incorporated variables such as pollster rating, transparency score, sample size, and state-level data to predict the percentage of support for each candidate across different states.
+By aggregating these state-level predictions and integrating electoral vote allocations, we projected that Kamala Harris would receive 306 electoral votes, surpassing the 270 required to win the presidency.
 
-The MLR model incorporated several significant predictors of voter support, including state-level factors, pollster ratings (pollscore), transparency scores, and sample size.
-By using the same model for both Harris and Trump, we were able to directly compare how these variables impacted each candidate's predicted support across different states.
+## Insights on Polling Methodologies and Voter Support
 
-For both Kamala Harris and Donald Trump, the model highlighted the substantial impact of state-level variations on their predicted support.
-Certain states showed higher predicted support for Harris, such as California and Maryland, which historically lean Democratic.
-Conversely, states like Indiana and Missouri exhibited higher predicted support for Trump, aligning with their Republican-leaning tendencies.
-The consistency of the model across both candidates allowed for a clear comparison of how state demographics and historical voting patterns influenced voter preferences.
+Our analysis indicates that pollster rating and state-level factors significantly influence reported support for both candidates.
+For Kamala Harris, the multiple linear regression model shows that higher pollster ratings and certain states are associated with increased support percentages.
+This suggests that polls conducted by more reliable pollsters may report higher support for Harris, possibly due to methodological rigor or sampling techniques that capture her voter base more accurately.
+For Donald Trump, pollster transparency and state differences play a notable role, indicating that how openly pollsters disclose their methodologies can affect the reported support for Trump.
 
-Pollster reliability indicators, such as numeric grades and transparency scores, were significant predictors for both candidates.
-This suggests that polls conducted by more credible organizations tended to report more accurate and potentially higher support levels for each candidate.
-The sample size also played a role, although its impact varied slightly between the two candidates.
-By using the same model, we ensured that these predictors were weighted equally in the analysis of both Harris's and Trump's support levels.
+## Implications for Electoral Forecasting
 
-When aggregating the state-level predictions to simulate the Electoral College outcomes, the findings suggest a competitive race.
-Harris is projected to receive 216 electoral votes, while Trump is predicted to secure 147 electoral votes.
-Neither candidate reaches the 270-vote threshold required to win the presidency, emphasizing the potential for a closely contested election.
-The remaining electoral votes are likely concentrated in battleground states, where voter preferences are more volatile and could ultimately determine the election outcome.
-
-## Model Strengths
-
-A primary strength of this analysis lies in the consistent application of the same MLR model to both candidates.
-This approach ensures that the comparison between Harris and Trump is based on the same criteria and that any differences in predicted support are attributable to variations in the data rather than differences in the modeling approach.
-
-By incorporating multiple predictors—including state-level data, pollster ratings, transparency scores, and sample size—the model provides a nuanced understanding of the factors influencing voter preferences.
-This multifaceted approach allows for more accurate predictions compared to models relying solely on national-level polling data or a single set of predictors.
-
-The use of stepwise model selection and checks for multicollinearity enhanced the robustness of the MLR model.
-Stepwise selection optimized the predictor variables, ensuring that only those contributing significantly to the model were included.
-Checking for multicollinearity through the Variance Inflation Factor (VIF) helped prevent unreliable estimates due to correlated predictors, thereby improving the model's validity for both candidates.
-
-Moreover, translating the predicted support percentages into Electoral College projections adds practical value to the analysis.
-Since the U.S.
-Presidential Election is determined by electoral votes rather than the popular vote, this approach offers a more realistic forecast of the election outcome.
-Using the same model for both candidates ensures that the Electoral College projections are directly comparable.
+The projection of Kamala Harris receiving 306 electoral votes underscores the importance of incorporating state-level analyses in electoral forecasting.
+Our findings highlight that national polling averages may not sufficiently capture the complexities of the Electoral College system.
+By focusing on state-specific predictions and accounting for variables that influence voter support in different regions, forecasters can achieve more accurate predictions.
+This approach acknowledges the heterogeneity of the electorate and the pivotal role of battleground states in determining election outcomes.
 
@@
 
-Despite the strengths, several limitations need to be acknowledged:
+Despite the strengths of our modeling approach, several limitations must be acknowledged.
+First, the reliance on historical polling data and pollster ratings assumes that past performance is indicative of future accuracy, which may not hold true if polling methodologies or voter behaviors change significantly.
+Second, the underrepresentation of certain states due to sparse polling data may affect the reliability of our state-level predictions.
+Assigning a default 50% support in states with missing data introduces uncertainty and may not reflect actual voter preferences.
+Additionally, the models may not fully account for dynamic factors such as late-breaking events, shifts in voter sentiment, or turnout variations that can influence election results.
 
-- Reliance on Polling Data Quality: The accuracy of the model is inherently dependent on the quality of the polling data used.
-While the analysis accounted for pollster ratings and transparency scores, inherent biases and methodological differences between polls could still affect the results.
-Sampling errors, non-response biases, and underrepresentation of certain demographic groups are persistent issues in polling data that can skew predictions for both candidates.
+## Future Research and Recommendations
 
-- Static Snapshot of Voter Preferences: The model provides a static view based on the most recent polling data available.
-It does not account for the dynamic nature of election campaigns, where voter preferences can shift rapidly due to various factors such as political events, debates, or emerging issues.
-This limitation means the model may not accurately predict future changes in voter sentiment leading up to the election for either candidate.
-
-- Exclusion of Other Influential Factors: The analysis did not incorporate other variables that can significantly impact election outcomes, such as economic indicators (e.g., unemployment rates, GDP growth), social movements, campaign strategies, or voter turnout efforts.
-The absence of these factors may limit the model's ability to fully capture the complexities of voter behavior for both Harris and Trump.
-
-- Electoral College Simplifications: The Electoral College simulation assumes that the candidate with the higher predicted support in a state wins all its electoral votes (except for Maine and Nebraska).
-This winner-takes-all approach does not account for the proportional allocation of electoral votes in those two states or potential variations in voter turnout that could affect the actual results.
-
-## Future Research 
-To enhance the predictive power and accuracy of election forecasting models, future research could consider the following approaches:
-
-- Incorporate Time-Series Analysis: Introducing a time-series component would allow the model to account for trends and shifts in voter preferences over time.
-This dynamic approach could capture the impact of events such as debates, policy announcements, or external factors on voter sentiment for both candidates.
-
-- Expand Predictive Variables: Including additional variables such as economic indicators, demographic data, and social media sentiment analysis could provide a more comprehensive view of the factors influencing voter behavior.
-Integrating data on unemployment rates or consumer confidence might reveal economic influences on candidate support.
-
-- Voter Turnout Models: Developing models to predict voter turnout could significantly improve election forecasts.
-Turnout can vary widely between elections and is often influenced by voter enthusiasm, mobilization efforts, and barriers to voting.
-Incorporating turnout probabilities could adjust the predicted support percentages to reflect more realistic electoral scenarios.
-
-- Advanced Modeling Techniques: Exploring advanced statistical methods or machine learning algorithms might capture complex nonlinear relationships between predictors and voter support.
-Techniques such as random forests, gradient boosting, or neural networks could uncover patterns not detectable through linear regression.
-
-- Account for Electoral Nuances: With variations in how electoral votes are allocated and the potential impact of third-party candidates or ranked-choice voting in some states, future models should consider these electoral nuances.
-Simulating different voting systems could provide insights into how alternative methods might affect election outcomes.
-
-- Cross-Validation with Alternative Data Sources: Validating the model against alternative data sources, such as exit polls or historical voting patterns, could test its robustness.
-Cross-validation helps ensure that the model is not overfitted to the polling data used and can generalize to different datasets.
+To enhance the accuracy of electoral forecasts, future research should consider incorporating real-time data and alternative data sources such as social media trends, economic indicators, and demographic shifts.
+Employing more sophisticated modeling techniques, such as hierarchical Bayesian models or machine learning algorithms, may capture complex interactions and non-linear relationships among variables.
+Furthermore, improving polling methodologies to address issues of response bias and sample representativeness is essential.
+Collaborations between pollsters, statisticians, and political scientists can foster the development of more robust predictive models that adapt to the evolving electoral landscape.
 
@@
 
-```{r, out.width="90%", out.height="70%", fig.align="center"}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-predicted-vote-pct
-#| fig-cap: Predicted Vote Percentages by State
-# knitr::include_graphics(here::here(("improved_vote_percentages_by_state.png")))
-
-```
-
 \newpage

### 2024-11-02 paper/paper.qmd
words changed to follow prose guideline

@@
 
-This statistical approach effectively bridges the gap between state-level polling data and the Electoral College outcome.
-By employing multiple linear regression models to predict polling percentages and integrating these predictions with the electoral vote framework, we provide a nuanced analysis of potential election results.
-This methodology respects the intricacies of the U.S.
+This statistical approach effectively connects state-level polling data to Electoral College projections.
+By employing multiple linear regression models to predict polling percentages and integrating these predictions with the electoral vote framework, we provide a detailed analysis of potential election results.
+This methodology reflects the structure of the U.S.
 electoral system and offers valuable insights into how predicted voter support may translate into electoral success.

### 2024-11-02 paper/paper.qmd

@@
 library(car)
+library(maps)
+library(stringr)
+library(reshape2)
+library(patchwork)
 
@@
 # Load necessary library
-library(ggplot2)
 
@@
 # Load necessary libraries
-library(reshape2)
-library(ggplot2)
 
@@
 
-# Load necessary libraries
-library(ggplot2)
-library(reshape2)
-library(dplyr)
-library(patchwork)
-
 # Melt data for easier plotting
@@
 **Aggregation of Predicted Support by State**: For each candidate, we computed the average predicted polling percentage within each state.
-This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see Figure @fig-predicted-vote-pct).
+This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see @fig-predicted-vote-pct).
 
@@
 Harris surpasses the 270-vote threshold required for victory, suggesting a strong position in the Electoral College based on our predictions.
-This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes.
+This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes (see @fig-map-state-winner).
 
@@
 electoral system and offers valuable insights into how predicted voter support may translate into electoral success.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-map-state-winner
+#| fig.cap: "Electoral College Predition Map"
+
+# Get US state map data
+us_states <- map_data("state")
+
+# Create a mapping of state names
+state_mapping <- data.frame(
+  state = c(
+    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware",
+    "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa",
+    "Kansas", "Kentucky", "Louisiana", "Maine At-Large", "Maine CD-1", "Maine CD-2", "Maryland",
+    "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana",
+    "Nebraska At-Large", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3",
+    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina",
+    "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island",
+    "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont",
+    "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming"
+  ),
+  region = tolower(c(
+    "alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut", "delaware",
+    "district of columbia", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa",
+    "kansas", "kentucky", "louisiana", "maine", "maine", "maine", "maryland",
+    "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "montana",
+    "nebraska", "nebraska", "nebraska", "nebraska",
+    "nevada", "new hampshire", "new jersey", "new mexico", "new york", "north carolina",
+    "north dakota", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode island",
+    "south carolina", "south dakota", "tennessee", "texas", "utah", "vermont",
+    "virginia", "washington", "west virginia", "wisconsin", "wyoming"
+  ))
+)
+
+# Simplify prediction data for mapping
+map_data <- prediction_comparison %>%
+  select(state, winner) %>%
+  left_join(state_mapping, by = "state")
+
+# Merge with map data
+map_data_final <- us_states %>%
+  left_join(map_data, by = "region")
+
+# Create the basic map without projection
+ggplot(data = map_data_final) +
+  geom_polygon(aes(x = long, y = lat, group = group, fill = winner), 
+               color = "white", size = 0.2) +
+  coord_fixed(1.3) +  # This replaces coord_map() with a simpler aspect ratio adjustment
+  scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080"),
+                    name = "Predicted Winner") +
+  theme_minimal() +
+  theme(
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank(),
+    plot.title = element_text(hjust = 0.5, size = 16),
+    legend.position = "bottom"
+  )
+```
 

### 2024-11-03 paper/paper.qmd
Some editting of the appendix writing

--- 2024-11-02 paper/paper.qmd
@@
 Through aggregating these state-level predictions, we simulate Electoral College outcomes to provide insights into each candidate's probability of securing the required 270 electoral votes.
-Our analysis also includes a detailed examination of YouGov's polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
+Our analysis also includes a detailed examination of [@YouGov] polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
 
@@
 # Data {#sec-data}
+
 ## Overview
 Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news.
-This database aggregates and standardizes polling results from various organizations, applying quality metrics and methodological assessments to each poll.
-In this section, we detail our selected variables, discuss key measurements, outline important limitations of our data, and our data cleaning process.
+This database compiles and standardizes polling results from various organizations, applying quality metrics and assessments to each poll.
+Several key variables from this dataset are crucial to our analysis:
 
-Our analysis focuses on several key variables that directly influence polling accuracy and electoral predictions. @tbl-overview_data presents a sample of our dataset.
+The dataset employed in this analysis encompasses polling data from various reputable sources, capturing critical variables related to polling organizations, sample sizes, candidate support levels, and polling coverage (state-level or national).
+The primary variables of interest are outlined as follows:
 
-## Poll Quality Variables
+- **Poll ID**: A unique identifier for each poll entry, facilitating data tracking and management (e.g., 88590).
+- **Pollster**: The organization conducting the poll, which serves as a primary indicator of the poll’s methodological quality (e.g., YouGov, Ipsos).
+- **Pollscore**: A quantitative measure of the pollster’s reliability, where lower (negative) values suggest higher predictive accuracy (e.g., -1.1 for YouGov).
+- **Sample Size**: The number of respondents included in each poll, influencing the statistical precision and margin of error (e.g., 1414).
+- **Support Percentage (pct)**: The proportion of respondents expressing support for each candidate, serving as the dependent variable in subsequent analyses (e.g., 47% for Kamala Harris).
+- **State**: Indicates the geographical focus of the poll, either at the state level or nationwide (e.g., National).
+- **Candidate Name**: The candidate assessed in the poll, providing context for support levels and enabling candidate-specific analyses (e.g., Kamala Harris, Donald Trump).
+- **End Date**: The date on which the poll concluded, offering temporal alignment for longitudinal analyses (e.g., 2024-10-07).
 
-@tbl-poll-metrics summarizes three primary metrics that assess poll quality in our dataset:
+To ensure analytical rigor, the dataset underwent extensive preprocessing.
+This involved standardizing variable formats for compatibility with regression models, converting support percentages into numerical values, and systematically addressing missing data to minimize bias.
+These preprocessing steps ensured that the dataset adhered to the standards of a "tidy" dataset, enabling robust model development and enhancing interpretability of subsequent statistical analyses.
 
-- **Pollscore**: Represents "the error and bias attributable to a pollster, where negative numbers indicate better performance." Our data shows a mean pollscore of -1.06 (SD = 0.28), with values ranging from -1.50 to -0.50, indicating generally high-quality polling organizations in our sample.
-- **Numeric Grade**: A rating given to pollsters to indicate their quality and reliability.
-With a mean of 2.89 (SD = 0.10) and a range of 2.70 to 3.00, our dataset maintains high standards.
-We established a minimum threshold of 2.7 to ensure methodological rigor.
-- **Transparency Score**: A measure reflecting pollsters' transparency about their methodology, with 10 being the highest possible score.
-The data shows strong overall transparency with a mean of 8.59 (SD = 1.04), indicating that most polls in our sample maintain high standards of methodological disclosure.
-
-## Methodological Variables
-
-- **Methodology**: @fig-methodology-dist shows the distribution of polling approaches used to conduct the polls.
-Live Phone methods and Online Panels are the predominant methods, with hybrid approaches combining multiple methods (such as IVR/Online Panel/Text-to-Web) representing evolving polling techniques adapting to changing communication patterns.
-- **Sample Size**: @fig-sample-size illustrates the distribution of poll sample sizes, representing the total number of respondents participating in each poll.
-Our dataset shows considerable variation, with a mean of 1,114.76 respondents (SD = 583.72).
-The right-skewed distribution ranges from 450 to 4,253 respondents, with larger sample sizes typically associated with national polls.
-
-## Geographic and Temporal Variables
-
-- **State Coverage**: @tbl-state-polls reveals the distribution of polls across different U.S.
-states.
-National polls lead with 261 surveys, followed by concentrated polling in key battleground states: Pennsylvania (86), Wisconsin (82), and North Carolina (64).
-This distribution reflects strategic focus on states where the electoral outcome is less certain.
-- **End Date**: @fig-temporal-polling tracks polling frequency from July through October 2024.
-We observe increased activity during September and early October, with weekly poll counts peaking in late September.
-This pattern reflects intensified polling efforts as the election approaches.
-
-## Outcome Variable
-
-- **Support Percentage**: @fig-candidate-support displays our primary dependent variable: the percentage of respondents expressing support for each candidate.
-The distribution reveals a competitive race between Donald Trump and Kamala Harris, with support levels typically ranging between 40-50% for both candidates.
+@tbl-overview_data shows a sample of the dataset.
 
@@
 
-- **Poll Quality:** While our pollscore and numeric grade filters help ensure data quality, these metrics are based on historical performance and may not fully capture current methodological improvements or deterioration.
-- **Temporal Dynamics:** Our dataset provides discrete snapshots of voter preferences rather than continuous measurement.
-This limitation is particularly relevant given the rapid evolution of political narratives and voter sentiment during presidential campaigns.
+- **Poll Quality:** While we filter for high-quality polls using numeric grades, differences in polling methodologies may introduce systematic biases.
+The inclusion of pollster ratings helps account for historical accuracy but cannot completely eliminate these potential biases.
+- **Temporal Dynamics:** Our dataset provides a snapshot of voter preferences during a specific timeframe.
+This static nature means we cannot capture the full dynamics of voter preference evolution over the campaign period.
 - **Geographic Coverage:** Although we have national and state-level polling data, coverage varies by state.
@@
 - **Response Bias:** Despite careful methodology by pollsters, self-selection bias in survey participation and social desirability bias in responses remain potential concerns.
+
+## Outcome variables
+Our primary outcome variable is the percentage of support (*pct*) for Kamala Harris and Donald Trump in each poll.
+This measurement represents the proportion of respondents who indicate they would vote for each candidate if the election were held on the day of the poll.
+The variable directly captures voter preferences and serves as the foundation for our electoral predictions.
+
+@fig-support below illustrates the distribution of support for both candidates, revealing several notable patterns.
+First, the support percentages cluster between 40% and 60%, reflecting the competitive nature of the race.
+Second, the distributions show slight differences between candidates, with Harris's support displaying more variation than Trump's.
+This pattern might reflect differences in voter certainty or polling methodology across different states and time periods.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-support
+#| fig-cap: Distribution of Support for Kamala Harris and Donald Trump
+
+ggplot(data, aes(x = pct, fill = candidate_name)) +
+  geom_histogram(bins = 30, alpha = 0.8, position = "identity") +
+  facet_wrap(~ candidate_name) +
+  labs(x = "Percentage of Support",
+       y = "Count") +
+  scale_fill_discrete(name = "Candidates") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+## Predictor variables
+
+Our model incorporates several key predictor variables, each chosen for its theoretical importance in explaining polling variations and electoral outcomes:
+
+- **Numeric Grade:** A composite measure (scale: 0-4) incorporating factors such as methodology rigor and historical accuracy
+- **Pollscore:** A measure of historical polling accuracy (range: -4 to +4, where negative scores indicate better performance)
+- **Transparency Score:** Quantifies the openness of polling methodology (scale: 0-10)
+- **Sample Size:** Number of respondents, typically ranging from 500 to 3000
+- **State:** State-level indicators capturing regional political variations
+- **Methodology:** Survey approach (e.g., online panel, phone interviews)
+
+The relationship between these predictors and polling outcomes is complex and often interconnected.
+For example, while larger sample sizes generally reduce sampling error, this effect may be moderated by the poll's methodology and quality metrics.
+Similarly, geographic variations in polling accuracy suggest that the relationship between predictors and outcomes may vary systematically across states.
+
+These variables were selected based on both theoretical foundations in electoral polling literature and practical considerations of data availability and quality.
+We focused on these core variables due to their consistent availability across polls and demonstrated importance in previous electoral forecasting efforts.
+
+**Edit note: Add a table in appendix to show sample with all variables used in model**
 
@@
 
-  > i. Converted *End Date* values to standardized date format
+  > i. Converted end dates to standardized date format
   > ii. Calculated absolute supporter numbers from percentages and sample sizes
   > iii. Created binary candidate indicators (Harris = 1, Trump = 0)
-
-Other variable such as ... were not selected for our analysis because ...
 
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Load necessary libraries
-library(reshape2)
-library(ggplot2)
-
-# Load final models for Harris and Trump
-mlr_harris_model_final <- readRDS(here::here("models/mlr_harris_model_final.rds"))
-mlr_trump_model_final <- readRDS(here::here("models/mlr_trump_model_final.rds"))
-
-# Predict poll percentages for Kamala Harris and Donald Trump
-harris_data$predicted_pct_harris <- predict(mlr_harris_model_final, newdata = harris_data)
-trump_data$predicted_pct_trump <- predict(mlr_trump_model_final, newdata = trump_data)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Define electoral votes for each state (and Washington, D.C.)
-# Correct the electoral_votes data frame
-electoral_votes <- data.frame(
-  state = c(
-    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", 
-    "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
-    "Kansas", "Kentucky", "Louisiana", 
-    # Maine entries
-    "Maine At-Large", "Maine CD-1", "Maine CD-2", 
-    "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", 
-    # Nebraska entries
-    "Nebraska At-Large", "Nebraska CD-1", "Nebraska CD-2", "Nebraska CD-3", 
-    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", 
-    "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", 
-    "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", 
-    "Wyoming"
-  ),
-  electoral_votes = c(
-    9, 3, 11, 6, 55, 9, 7, 3, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8,
-    # Maine electoral votes
-    2, 1, 1,
-    10, 11, 16, 10, 6, 10, 3,
-    # Nebraska electoral votes
-    2, 1, 1, 1,
-    6, 4, 14, 5, 29, 15, 3,
-    18, 7, 7, 20, 4, 9, 3,
-    11, 38, 6, 3, 13, 12, 5, 10,
-    3
-  )
-)
-
-# Aggregate predictions by state for Harris and Trump
-harris_state_avg <- aggregate(predicted_pct_harris ~ state, data = harris_data, FUN = mean)
-trump_state_avg <- aggregate(predicted_pct_trump ~ state, data = trump_data, FUN = mean)
-
-# Merge predictions for both candidates
-prediction_comparison <- merge(harris_state_avg, trump_state_avg, by = "state", all.x = TRUE)
-
-```
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Identify any states missing from the prediction data
-missing_states <- setdiff(electoral_votes$state, prediction_comparison$state)
-
-# Assign 50% support to both candidates for missing states
-missing_states_data <- data.frame(
-  state = missing_states,
-  predicted_pct_harris = 50,
-  predicted_pct_trump = 50
-)
-
-# Append missing states data to prediction_comparison
-prediction_comparison <- rbind(prediction_comparison, missing_states_data)
-
-# Merge electoral votes into the prediction_comparison to ensure each state has electoral votes
-prediction_comparison <- merge(prediction_comparison, electoral_votes, by = "state", all.x = TRUE)
-
-```
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-# Determine the winner for each state based on predicted percentages
-set.seed(304)
-prediction_comparison$winner <- ifelse(
-  prediction_comparison$predicted_pct_harris > prediction_comparison$predicted_pct_trump, 
-  "Harris", 
-  ifelse(prediction_comparison$predicted_pct_trump > prediction_comparison$predicted_pct_harris, 
-         "Trump", 
-         sample(c("Harris", "Trump"), nrow(prediction_comparison), replace = TRUE))
-)
-
-# Calculate total electoral votes for each candidate
-harris_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Harris"], na.rm = TRUE)
-trump_electoral_votes <- sum(prediction_comparison$electoral_votes[prediction_comparison$winner == "Trump"], na.rm = TRUE)
-
-# Create a table with electoral votes
-electoral_votes_table <- data.frame(
-  Candidate = c("Harris", "Trump"),
-  Electoral_Votes = c(harris_electoral_votes, trump_electoral_votes)
-)
-
-# Display the table with kable
-kable(electoral_votes_table, caption = "Electoral Votes for Each Candidate", col.names = c("Candidate", "Electoral Votes"))
-```
-
 # Electoral College Prediction {#sec-predict}
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-predicted-vote-pct
-#| fig.height: 5
-#| fig.cap: "Predicted Vote Percentages by State"
-
-# Load necessary libraries
-library(ggplot2)
-library(reshape2)
-library(dplyr)
-library(patchwork)
-
-# Melt data for easier plotting
-prediction_melted <- melt(
-  prediction_comparison, 
-  id.vars = c("state", "electoral_votes", "winner"),
-  measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
-)
-
-# Rename melted variables for readability
-colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
-prediction_melted$candidate <- ifelse(
-  prediction_melted$candidate == "predicted_pct_harris", 
-  "Harris", 
-  "Trump"
-)
-
-# Calculate the midpoint to split the states
-n_states <- length(unique(prediction_melted$state))
-midpoint <- ceiling(n_states / 2)
-
-# Order states by electoral votes in descending order for better visualization
-states_order <- prediction_melted %>%
-  distinct(state, electoral_votes) %>%
-  arrange(desc(electoral_votes)) %>%
-  pull(state)
-
-# Split the states into two groups
-left_states <- states_order[1:midpoint]
-right_states <- states_order[(midpoint + 1):length(states_order)]
-
-# Create two separate datasets
-prediction_left <- prediction_melted %>%
-  filter(state %in% left_states) %>%
-  mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
-
-prediction_right <- prediction_melted %>%
-  filter(state %in% right_states) %>%
-  mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
-
-# Function to create individual plots to avoid repetition
-create_plot <- function(data, show_legend = FALSE) {
-  ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
-    geom_bar(
-      stat = "identity", 
-      position = position_dodge(width = 0.9), 
-      width = 0.5  # Reduced bar width for more spacing
-    ) +
-    coord_flip() +
-    labs(
-      x = NULL,
-      y = "Predicted Vote Percentage (%)"
-    ) +
-    scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
-    theme_minimal() +
-    theme(
-      text = element_text(size = 9),  # Reduced overall text size
-      axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
-      axis.text.x = element_text(size = 9),
-      legend.title = element_blank(),
-      legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
-      legend.text = element_text(size = 9),
-      panel.grid.major.y = element_blank(),
-      plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
-    ) +
-    geom_text(
-      aes(label = round(value, 1)),
-      position = position_dodge(width = 0.9),
-      hjust = -0.2,    # Adjusted horizontal position for labels
-      size = 1.5,      # Smaller label size
-      show.legend = FALSE
-    ) +
-    scale_y_continuous(
-      expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
-      limits = c(0, 75)  # Adjust based on your data
-    )
-}
-
-# Create the left plot with legend
-plot_left <- create_plot(prediction_left, show_legend = TRUE)
-
-# Create the right plot without legend
-plot_right <- create_plot(prediction_right, show_legend = FALSE)
-
-# Combine the plots side by side with legend on top
-combined_plot <- plot_left + plot_right +
-  plot_layout(guides = "collect") &  # Collects the legend at the top
-  theme(legend.position = "top")  # Places the combined legend at the top
-
-# Print the combined plot
-combined_plot
-
-```
-
-To project the outcome of the 2024 U.S.
-Presidential election between Kamala Harris and Donald Trump, we conducted a statistical analysis that integrates predicted state-level polling percentages with the allocation of electoral votes.
-This methodology employs multiple linear regression models to estimate each candidate's support and translates these predictions into potential Electoral College results, adhering to the structure of the U.S.
-electoral system.
+To forecast the winner of the 2024 U.S.
+Presidential election between Kamala Harris and Donald Trump, we performed the following series of steps, focusing on the distribution of electoral votes across states based on predicted polling percentages.
+This approach allows for an estimation of each candidate's electoral college support under the U.S.
+voting system.
 
 1.
-**Prediction of State-Level Polling Percentages**: We utilized finalized multiple linear regression models for both Kamala Harris and Donald Trump.
-These models were applied to their respective datasets to generate predicted polling percentages for each candidate across various states.
-The predictions reflect the expected levels of support based on historical polling data and relevant covariates included in the models.
+Predicting State-Level Polling Percentages: We used final multiple linear regression models for both candidates, Harris and Trump.
+These models generated predicted polling percentages for each candidate across different states, capturing their expected levels of support based on the poll data used.
 
 2.
-**Aggregation of Predicted Support by State**: For each candidate, we computed the average predicted polling percentage within each state.
-This aggregation provides a concise state-level summary of anticipated support, facilitating direct comparisons between the candidates in each state (see Figure @fig-predicted-vote-pct).
+Averaging Predicted Support by State: For both Harris and Trump, we calculated the average predicted polling percentage within each state.
+This step provides a state-level summary of each candidate's support based on our predictions, simplifying comparisons between them in each state.
 
 3.
-**Integration of Electoral Vote Allocations**: We established a dataset detailing the electoral vote distribution for each state, including the distinct allocations for Maine and Nebraska, which can split their electoral votes by congressional district.
-This dataset was merged with the state-level predicted support data to align electoral votes with the corresponding predicted support.
+Assigning Electoral Votes: We then merged our state-level predictions with the distribution of electoral votes, as assigned by state in the Electoral College.
+This dataset provided the electoral vote count for each state, including special allocations for districts in Maine and Nebraska, which can split their electoral votes by congressional district.
 
 4.
-**Addressing Missing Data**: We identified states lacking predicted polling data and assigned a neutral predicted support of 50% to both candidates in these states.
-This assumption accounts for the absence of data while ensuring that all states are included in the analysis.
+Determining State Winners: Using the averaged predictions, we identified the winner in each state by comparing the predicted support percentages.
+Harris or Trump was deemed the winner in a state if their predicted polling percentage exceeded that of their opponent.
 
 5.
-**Determination of State Winners**: Based on the aggregated predicted percentages, we determined the projected winner in each state.
-A candidate was designated as the winner if their predicted polling percentage exceeded that of their opponent.
-In cases where the predicted percentages were equal, we randomly assigned the winner using a stochastic approach to reflect the uncertainty inherent in tied predictions.
+Aggregating Electoral Votes: Finally, we summed the electoral votes for each candidate across all states where they were predicted to win.
+This yielded total electoral vote counts for both Harris and Trump, which are essential for determining the predicted winner under the U.S.
+system, where 270 electoral votes are required to secure the presidency.
 
 6.
-**Calculation of Total Electoral Votes**: We summed the electoral votes from all states won by each candidate to compute their total electoral vote counts.
-This calculation is critical, as securing at least 270 electoral votes is necessary to win the presidency under the U.S.
-electoral system.
+Interpreting the Results: With Harris predicted to receive 216 electoral votes compared to Trump’s 147, Harris demonstrates a stronger position in the Electoral College, though neither candidate reaches the 270-vote threshold required for a decisive victory.
+This outcome suggests that, while a clear electoral win is not forecasted, Harris holds a better chance of winning the election based on her lead in projected electoral votes.
+This lead places her closer to the threshold and may indicate a competitive advantage should undecided states or close races lean in her favor.
 
-7.
-**Interpretation of Results**: The analysis indicated that Kamala Harris is projected to receive a total of **306 electoral votes**, while Donald Trump is projected to receive **232 electoral votes**.
-Harris surpasses the 270-vote threshold required for victory, suggesting a strong position in the Electoral College based on our predictions.
-This outcome implies that Harris is likely to win the election, given her substantial lead in projected electoral votes.
-
-This statistical approach effectively connects state-level polling data to Electoral College projections.
-By employing multiple linear regression models to predict polling percentages and integrating these predictions with the electoral vote framework, we provide a detailed analysis of potential election results.
-This methodology reflects the structure of the U.S.
-electoral system and offers valuable insights into how predicted voter support may translate into electoral success.
+This electoral vote aggregation based on state-level predictions provides an interpretation of how polling data (@fig-predicted-vote-pct), analyzed through multiple linear regression models, translates into likely electoral support for each candidate.
+It effectively models potential election outcomes while respecting the structure of the Electoral College​​.
 
@@
 
-## Summary of Findings
+## Key Findings
 
-In this paper, we developed statistical models to forecast the outcome of the 2024 United States presidential election between Kamala Harris and Donald Trump.
-Utilizing polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database, we analyzed factors influencing candidate support through linear and multiple linear regression models.
-We incorporated variables such as pollster rating, transparency score, sample size, and state-level data to predict the percentage of support for each candidate across different states.
-By aggregating these state-level predictions and integrating electoral vote allocations, we projected that Kamala Harris would receive 306 electoral votes, surpassing the 270 required to win the presidency.
+This study set out to forecast the outcome of the 2024 U.S.
+Presidential Election between Kamala Harris and Donald Trump by developing a multiple linear regression (MLR) model based on comprehensive polling data.
+Importantly, the same MLR model was applied to both candidates, ensuring consistency in how predictors influenced the predicted support percentages for each.
 
-## Insights on Polling Methodologies and Voter Support
+The MLR model incorporated several significant predictors of voter support, including state-level factors, pollster ratings (pollscore), transparency scores, and sample size.
+By using the same model for both Harris and Trump, we were able to directly compare how these variables impacted each candidate's predicted support across different states.
 
-Our analysis indicates that pollster rating and state-level factors significantly influence reported support for both candidates.
-For Kamala Harris, the multiple linear regression model shows that higher pollster ratings and certain states are associated with increased support percentages.
-This suggests that polls conducted by more reliable pollsters may report higher support for Harris, possibly due to methodological rigor or sampling techniques that capture her voter base more accurately.
-For Donald Trump, pollster transparency and state differences play a notable role, indicating that how openly pollsters disclose their methodologies can affect the reported support for Trump.
+For both Kamala Harris and Donald Trump, the model highlighted the substantial impact of state-level variations on their predicted support.
+Certain states showed higher predicted support for Harris, such as California and Maryland, which historically lean Democratic.
+Conversely, states like Indiana and Missouri exhibited higher predicted support for Trump, aligning with their Republican-leaning tendencies.
+The consistency of the model across both candidates allowed for a clear comparison of how state demographics and historical voting patterns influenced voter preferences.
 
-## Implications for Electoral Forecasting
+Pollster reliability indicators, such as numeric grades and transparency scores, were significant predictors for both candidates.
+This suggests that polls conducted by more credible organizations tended to report more accurate and potentially higher support levels for each candidate.
+The sample size also played a role, although its impact varied slightly between the two candidates.
+By using the same model, we ensured that these predictors were weighted equally in the analysis of both Harris's and Trump's support levels.
 
-The projection of Kamala Harris receiving 306 electoral votes underscores the importance of incorporating state-level analyses in electoral forecasting.
-Our findings highlight that national polling averages may not sufficiently capture the complexities of the Electoral College system.
-By focusing on state-specific predictions and accounting for variables that influence voter support in different regions, forecasters can achieve more accurate predictions.
-This approach acknowledges the heterogeneity of the electorate and the pivotal role of battleground states in determining election outcomes.
+When aggregating the state-level predictions to simulate the Electoral College outcomes, the findings suggest a competitive race.
+Harris is projected to receive 216 electoral votes, while Trump is predicted to secure 147 electoral votes.
+Neither candidate reaches the 270-vote threshold required to win the presidency, emphasizing the potential for a closely contested election.
+The remaining electoral votes are likely concentrated in battleground states, where voter preferences are more volatile and could ultimately determine the election outcome.
+
+## Model Strengths
+
+A primary strength of this analysis lies in the consistent application of the same MLR model to both candidates.
+This approach ensures that the comparison between Harris and Trump is based on the same criteria and that any differences in predicted support are attributable to variations in the data rather than differences in the modeling approach.
+
+By incorporating multiple predictors—including state-level data, pollster ratings, transparency scores, and sample size—the model provides a nuanced understanding of the factors influencing voter preferences.
+This multifaceted approach allows for more accurate predictions compared to models relying solely on national-level polling data or a single set of predictors.
+
+The use of stepwise model selection and checks for multicollinearity enhanced the robustness of the MLR model.
+Stepwise selection optimized the predictor variables, ensuring that only those contributing significantly to the model were included.
+Checking for multicollinearity through the Variance Inflation Factor (VIF) helped prevent unreliable estimates due to correlated predictors, thereby improving the model's validity for both candidates.
+
+Moreover, translating the predicted support percentages into Electoral College projections adds practical value to the analysis.
+Since the U.S.
+Presidential Election is determined by electoral votes rather than the popular vote, this approach offers a more realistic forecast of the election outcome.
+Using the same model for both candidates ensures that the Electoral College projections are directly comparable.
 
@@
 
-Despite the strengths of our modeling approach, several limitations must be acknowledged.
-First, the reliance on historical polling data and pollster ratings assumes that past performance is indicative of future accuracy, which may not hold true if polling methodologies or voter behaviors change significantly.
-Second, the underrepresentation of certain states due to sparse polling data may affect the reliability of our state-level predictions.
-Assigning a default 50% support in states with missing data introduces uncertainty and may not reflect actual voter preferences.
-Additionally, the models may not fully account for dynamic factors such as late-breaking events, shifts in voter sentiment, or turnout variations that can influence election results.
+Despite the strengths, several limitations need to be acknowledged:
 
-## Future Research and Recommendations
+- Reliance on Polling Data Quality: The accuracy of the model is inherently dependent on the quality of the polling data used.
+While the analysis accounted for pollster ratings and transparency scores, inherent biases and methodological differences between polls could still affect the results.
+Sampling errors, non-response biases, and underrepresentation of certain demographic groups are persistent issues in polling data that can skew predictions for both candidates.
 
-To enhance the accuracy of electoral forecasts, future research should consider incorporating real-time data and alternative data sources such as social media trends, economic indicators, and demographic shifts.
-Employing more sophisticated modeling techniques, such as hierarchical Bayesian models or machine learning algorithms, may capture complex interactions and non-linear relationships among variables.
-Furthermore, improving polling methodologies to address issues of response bias and sample representativeness is essential.
-Collaborations between pollsters, statisticians, and political scientists can foster the development of more robust predictive models that adapt to the evolving electoral landscape.
+- Static Snapshot of Voter Preferences: The model provides a static view based on the most recent polling data available.
+It does not account for the dynamic nature of election campaigns, where voter preferences can shift rapidly due to various factors such as political events, debates, or emerging issues.
+This limitation means the model may not accurately predict future changes in voter sentiment leading up to the election for either candidate.
+
+- Exclusion of Other Influential Factors: The analysis did not incorporate other variables that can significantly impact election outcomes, such as economic indicators (e.g., unemployment rates, GDP growth), social movements, campaign strategies, or voter turnout efforts.
+The absence of these factors may limit the model's ability to fully capture the complexities of voter behavior for both Harris and Trump.
+
+- Electoral College Simplifications: The Electoral College simulation assumes that the candidate with the higher predicted support in a state wins all its electoral votes (except for Maine and Nebraska).
+This winner-takes-all approach does not account for the proportional allocation of electoral votes in those two states or potential variations in voter turnout that could affect the actual results.
+
+## Future Research 
+To enhance the predictive power and accuracy of election forecasting models, future research could consider the following approaches:
+
+- Incorporate Time-Series Analysis: Introducing a time-series component would allow the model to account for trends and shifts in voter preferences over time.
+This dynamic approach could capture the impact of events such as debates, policy announcements, or external factors on voter sentiment for both candidates.
+
+- Expand Predictive Variables: Including additional variables such as economic indicators, demographic data, and social media sentiment analysis could provide a more comprehensive view of the factors influencing voter behavior.
+Integrating data on unemployment rates or consumer confidence might reveal economic influences on candidate support.
+
+- Voter Turnout Models: Developing models to predict voter turnout could significantly improve election forecasts.
+Turnout can vary widely between elections and is often influenced by voter enthusiasm, mobilization efforts, and barriers to voting.
+Incorporating turnout probabilities could adjust the predicted support percentages to reflect more realistic electoral scenarios.
+
+- Advanced Modeling Techniques: Exploring advanced statistical methods or machine learning algorithms might capture complex nonlinear relationships between predictors and voter support.
+Techniques such as random forests, gradient boosting, or neural networks could uncover patterns not detectable through linear regression.
+
+- Account for Electoral Nuances: With variations in how electoral votes are allocated and the potential impact of third-party candidates or ranked-choice voting in some states, future models should consider these electoral nuances.
+Simulating different voting systems could provide insights into how alternative methods might affect election outcomes.
+
+- Cross-Validation with Alternative Data Sources: Validating the model against alternative data sources, such as exit polls or historical voting patterns, could test its robustness.
+Cross-validation helps ensure that the model is not overfitted to the polling data used and can generalize to different datasets.
 
@@
 ## YouGov Pollster Methodology Overview and Evaluation
-YouGov conducts online surveys through their proprietary panel of U.S.
-adults, using nonprobability sampling methods combined with sophisticated weighting procedures to achieve representative results.
-Their approach balances speed and cost-effectiveness with statistical rigor through careful sample selection and data quality controls.
+YouGov utilizes an online panel with non-probability sampling, adjusting for demographic imbalances through weighting techniques.
+This approach ensures that survey results are representative of the broader population, despite the limitations of non-random sampling.
+
+The primary strength of YouGov's methodology lies in its efficiency and cost-effectiveness, enabling rapid data collection.
+However, the reliance on non-probability sampling can introduce biases, which are partially mitigated by their rigorous weighting process.
+While effective in providing timely insights, this method requires careful consideration of its limitations in representativeness.
 
 ### Survey Population and Sampling
-YouGov's target population typically comprises all U.S.
-adults or adult citizens, with their sampling frame consisting of their opt-in online panel covering approximately 95% of Americans.
-For general population surveys, they aim for 1,000-2,000 respondents, selected based on demographic and political characteristics to match the target population.
+YouGov’s surveys target U.S.
+adults, utilizing an opt-in online panel that covers approximately 95% of the population.
+For general population surveys, they typically sample between 1,000 and 2,000 respondents.
+These participants are selected to reflect key demographic and political characteristics, ensuring the sample is representative of the broader population.
 
 ### Panel Recruitment and Participation
-Panel members are recruited through advertising and website partnerships, with surveys offered in multiple languages including Spanish to ensure broad representation.
-Participants receive points exchangeable for small monetary rewards, though many report being motivated by the desire to contribute to research.
+[@YouGov] recruits panel members through targeted advertising and partnerships with various websites.
+Surveys are available in multiple languages, including Spanish, to ensure inclusivity and broad demographic representation.
+Participants are incentivized with points redeemable for small monetary rewards, though many join to contribute to research efforts.
 
 ### Quality Control
-YouGov employs several measures to maintain data quality:
+[@YouGov] employs several measures to maintain data quality:
 
@@
 ### Non-response and Weighting
-To address potential biases, YouGov applies statistical weighting based on demographics (age, gender, race, education) and political factors (voting behavior, party identification).
+To address potential biases, [@YouGov] applies statistical weighting based on demographics (age, gender, race, education) and political factors (voting behavior, party identification).
 Their weighting process considers multiple characteristics simultaneously to better reflect real-world demographic intersections.
@@
 ### Strengths and Limitations
-The methodology's primary strengths include rapid data collection, cost-effectiveness, and the ability to track opinions over time.
-However, the nonprobability sampling approach may introduce biases, and the online-only format could underrepresent certain populations.
-While weighting helps address these limitations, it cannot fully account for all potential sources of bias.
+The key strengths of YouGov’s methodology are its rapid data collection, cost-effectiveness, and capacity to track public opinion trends over time.
+However, the reliance on non-probability sampling can introduce biases, and the online-only format may lead to underrepresentation of certain populations.
+While sophisticated weighting techniques mitigate some of these issues, they cannot fully eliminate all potential biases.
 
@@
 ## Idealized Survey Methodology
-This idealized survey methodology outlines a comprehensive plan for forecasting the US presidential election within a budget of $100,000.
-The approach is designed to be statistically sound, practical, and capable of accurately predicting election outcomes by considering both the popular vote and electoral college implications.
+This idealized survey methodology presents a structured plan for forecasting the U.S.
+presidential election within a $100,000 budget.
+It balances statistical rigor with practicality to ensure accurate predictions of both the popular vote and Electoral College outcomes.
+The methodology includes advanced sampling techniques, robust survey design, and comprehensive data analysis to capture key demographic and political variables.
+By addressing potential biases and focusing on cost-effective data collection, this plan ensures reliable and representative insights into voter behavior.
 
@@
 #| label: tbl-overview_data
-#| tbl-cap: Sample Overview of Selected Variables in the Polling Dataset
-
-set.seed(778) # For reproducability
 
 data_overview <- data %>%
-  select(pollster, pollscore, numeric_grade, transparency_score, sample_size, 
-         methodology, pct, state, candidate_name, end_date)
+  select(poll_id, pollster, pollscore, sample_size, pct, state, candidate_name, end_date)
 
 # Transpose the first 5 rows to display them vertically
-data_overview <- data_overview %>%
-  sample_n(5) %>%
+data_overview_transposed <- data_overview %>%
+  head(5) %>%
   t()
 
-# Display the transposed table
-kable(data_overview)
-```
+# Display the transposed table with a caption
+data_overview_transposed %>%
+  kable(caption = "Sample Overview of Selected Variables in the Polling Dataset")
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-poll-metrics
-#| tbl-cap: Summary Statistics of Key Poll Metrics
-
-data %>%
-  summarise(
-    across(c(sample_size, pollscore, numeric_grade, transparency_score),
-           list(
-             Mean = mean,
-             Median = median,
-             SD = sd,
-             Min = min,
-             Max = max
-           ), 
-           .names = "{.fn}_{.col}",
-           na.rm = TRUE)
-  ) %>%
-  pivot_longer(everything(), 
-               names_to = c("Statistic", "Variable"),
-               names_sep = "_",
-               values_to = "Value") %>%
-  pivot_wider(names_from = Variable,
-              values_from = Value) %>%
-  mutate(across(-Statistic, ~round(.x, 2))) %>%
-  kable()
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-methodology-dist
-#| fig-cap: Distribution of Polling Methodologies
-
-data %>%
-  ggplot(aes(x = fct_infreq(methodology))) +
-  geom_bar() +
-  coord_flip() +
-  theme_minimal() +
-  labs(x = "Methodology",
-       y = "Count") +
-  theme(axis.text.y = element_text(size = 8))
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-candidate-support
-#| fig-cap: Distribution of Support Percentages by Candidate
-
-ggplot(data, aes(x = pct, fill = candidate_name)) +
-  geom_histogram(bins = 30, position = "identity") +
-  facet_wrap(~ candidate_name) +
-  labs(x = "Percentage of Support",
-       y = "Count") +
-  scale_fill_discrete(name = "Candidates") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-state-polls
-#| tbl-cap: Polling Frequency by State
-
-data %>%
-  count(state) %>%
-  arrange(desc(n)) %>%
-  rename("State" = state,
-         "Number of Polls" = n) %>%
-  kable(format = "latex", booktabs = TRUE)
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-temporal-polling
-#| fig-cap: Number of Polls Over Time
-
-data %>%
-  mutate(end_date = as.Date(end_date)) %>%
-  mutate(week = floor_date(end_date, "week")) %>%
-  count(week) %>%
-  ggplot(aes(x = week, y = n)) +
-  geom_line() +
-  geom_point() +
-  theme_minimal() +
-  labs(x = "Week",
-       y = "Number of Polls") +
-  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d")
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-sample-size
-#| fig-cap: Distribution of Poll Sample Sizes
-
-data %>%
-  ggplot(aes(x = sample_size)) +
-  geom_histogram(binwidth = 100) +
-  theme_minimal() +
-  labs(x = "Sample Size",
-       y = "Count") +
-  scale_x_continuous(breaks = seq(0, max(data$sample_size), by = 500))
 ```

### 2024-11-03 paper/paper.qmd

--- 2024-11-02 paper/paper.qmd
@@
 ## YouGov Pollster Methodology Overview and Evaluation
-YouGov conducts online surveys through their proprietary panel of U.S.
-adults, using nonprobability sampling methods combined with sophisticated weighting procedures to achieve representative results.
-Their approach balances speed and cost-effectiveness with statistical rigor through careful sample selection and data quality controls.
+YouGov[@YouGov] utilizes an online panel with non-probability sampling, adjusting for demographic imbalances through weighting techniques.
+This approach ensures that survey results are representative of the broader population, despite the limitations of non-random sampling.
+
+The primary strength of YouGov's methodology lies in its efficiency and cost-effectiveness, enabling rapid data collection.
+However, the reliance on non-probability sampling can introduce biases, which are partially mitigated by their rigorous weighting process.
+While effective in providing timely insights, this method requires careful consideration of its limitations in representativeness.
 
 ### Survey Population and Sampling
-YouGov's target population typically comprises all U.S.
-adults or adult citizens, with their sampling frame consisting of their opt-in online panel covering approximately 95% of Americans.
-For general population surveys, they aim for 1,000-2,000 respondents, selected based on demographic and political characteristics to match the target population.
+YouGov’s surveys target U.S.
+adults, utilizing an opt-in online panel that covers approximately 95% of the population.
+For general population surveys, they typically sample between 1,000 and 2,000 respondents.
+These participants are selected to reflect key demographic and political characteristics, ensuring the sample is representative of the broader population.
 
 ### Panel Recruitment and Participation
-Panel members are recruited through advertising and website partnerships, with surveys offered in multiple languages including Spanish to ensure broad representation.
-Participants receive points exchangeable for small monetary rewards, though many report being motivated by the desire to contribute to research.
+YouGov[@YouGov] recruits panel members through targeted advertising and partnerships with various websites.
+Surveys are available in multiple languages, including Spanish, to ensure inclusivity and broad demographic representation.
+Participants are incentivized with points redeemable for small monetary rewards, though many join to contribute to research efforts.
 
 ### Quality Control
-YouGov employs several measures to maintain data quality:
+YouGov[@YouGov] employs several measures to maintain data quality:
 
@@
 ```
+@tbl-overview_data provides an overview of key variables from different polls conducted by various organizations, including Washington Post/George Mason University, CNN/SSRS, Siena/NYT, and YouGov.
+It highlights differences in methodologies, such as live phone interviews, online panels, and mixed methods, as well as sample sizes ranging from 708 to 1,142 respondents.
+Pollscores, which reflect pollster reliability, vary from -1.5 (indicating higher reliability for Siena/NYT) to -0.6 (for CNN/SSRS).
+Transparency scores show that YouGov and CNN/SSRS exhibit full transparency with perfect scores of 10.
+
+The table also includes data on candidate support percentages, with Donald Trump polling between 42% and 48% across various states and Kamala Harris receiving 48% in Michigan.
+This snapshot reveals the diversity in poll designs, geographical focus, and methodological rigor, which are crucial for understanding the variability in polling data.
 
@@
 ```
+@tbl-poll-metrics presents the summary statistics for key poll metrics, including sample size, pollscore (pollster reliability), numeric grade, and transparency score.
+The average sample size across polls is 1,114.76 respondents, with a standard deviation of 583.72, indicating variability in the number of participants.
+The median sample size is 1,000, with a minimum of 450 and a maximum of 4,253 respondents.
+
+The average pollscore is -1.06, with a median of -1.10, reflecting generally high reliability across pollsters (lower scores indicate greater reliability).
+The numeric grade, which measures methodological rigor, averages 2.89 with minimal variability (SD of 0.10).
+The transparency score averages 8.59, suggesting most polls are transparent about their methods, though scores range from 6.00 to 10.00.
+These statistics highlight the overall quality and variation in polling practices used in the analysis.
 
@@
 ```
+@fig-methodology-dist illustrates the distribution of various polling methodologies used in the dataset.
+The most frequently employed methods include "Live Phone," "Online Panel," and "Live Phone/Text-to-Web," highlighting a preference for direct communication methods combined with online components.
+Less common methodologies, such as "Live Phone/Online Panel/Text" and "Live Phone/Email," appear at the lower end of the distribution.
+This diversity in polling methods reflects the varied approaches taken by pollsters to gather representative data, balancing traditional techniques with modern, web-based options.
 
@@
 ```
+@fig-candidate-support shows the distribution of support percentages for Donald Trump and Kamala Harris.
+Both distributions are centered around the mid-40% to mid-50% range, reflecting a competitive race.
+Trump's support is more concentrated around the 45-50% range, indicating relatively consistent polling results.
+Harris’s distribution is slightly more varied, suggesting greater fluctuation in her polling support.
+This visualization highlights the close contest between the two candidates, with neither consistently achieving overwhelming support across polls.
 
@@
 ```
+@tbl-state-polls illustrates the number of polls conducted over time, showing weekly fluctuations in polling activity.
+There is a noticeable increase in the number of polls around mid-September, peaking at over 160 polls in a week, followed by a decline as the election date approaches.
+This pattern suggests heightened polling activity during key points in the campaign, likely driven by major events or debates, before tapering off closer to the election.
+The variability highlights how polling frequency can change in response to the evolving political landscape.
 
@@
 ```
+@fig-temporal-polling provides the polling frequency by state, highlighting where polling efforts were concentrated.
+National polls dominate with 261 instances, while key battleground states like Pennsylvania (86 polls), Wisconsin (82 polls), and North Carolina (64 polls) also see significant attention.
+These states are critical for determining election outcomes, explaining the higher frequency.
+Conversely, states like Missouri, Montana, and Nebraska show fewer polls (4 each), reflecting less competitive races or a perceived lower impact on the overall election result.
+This distribution underscores the strategic focus on swing states in polling efforts.
 
@@
 
+@fig-sample-size shows the distribution of sample sizes across the polls in the dataset.
+The majority of polls have sample sizes clustered around 1,000 respondents, which is a typical size for achieving a balance between precision and cost.
+A few polls have significantly larger sample sizes, extending beyond 2,000, with some even exceeding 4,000 respondents.
+This suggests that while most polls aim for a moderate sample size, there are instances where more extensive sampling is conducted, possibly for greater accuracy or in critical regions.
+The distribution reflects the common practice of balancing statistical reliability with logistical constraints.
+
 ```{r}

### 2024-11-03 paper/paper.qmd
changed data to parquet as well as adding model formula

--- 2024-11-02 paper/paper.qmd
@@
 options(repos = c(CRAN = "https://cran.rstudio.com/"))
-
 library(tidyverse)
@@
 library(patchwork)
+library(kableExtra)
+library(arrow)
 
 # Load the dataset
-data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
@@
 
-To analyze factors influencing candidate support, we use a linear regression model with sample size as the primary predictor.
-This model examines the relationship between sample size and the percentage of support for Kamala Harris and Donald Trump across different polls.
-By including sample size as a predictor, we can assess whether larger or smaller sample sizes impact reported support levels for each candidate.
+To investigate the factors influencing voter support for Kamala Harris and Donald Trump, we employed linear regression models using polling data.
+Our analysis aimed to assess how predictors such as sample size, pollster ratings, transparency scores, and state-level variables affect the reported percentage of support for each candidate.
 
@@
 
-For Kamala Harris, the linear regression results show that sample size has a statistically insignificant effect on her support percentage.
-This implies that other factors, such as pollster methodology or regional biases, may play a more substantial role in shaping her reported support.
-
-In contrast, the results for Donald Trump indicate a weak but statistically significant negative effect of sample size on his support percentage.
-This suggests that, on average, larger sample sizes tend to show slightly lower support for Trump, though the effect size remains small.
+We began with simple linear regression models to examine the relationship between sample size (`sample_size`) and the percentage of support (`pct`) for each candidate.
 
@@
 
-To better capture the complexity of voter support, we constructed multiple linear regression (MLR) models for both Kamala Harris and Donald Trump.
-These models incorporate key predictors, including pollster rating (represented by pollscore), transparency score, sample size, and state-level data.
-By including these variables, the MLR models aim to control for multiple factors influencing voter support and to enhance the accuracy of our predictions.
-
-The MLR model for Kamala Harris indicates that factors such as state and pollster rating significantly impact her predicted level of support across polls.
-The results, shown in @fig-mlr-harris, suggest that Harris's support varies substantially by state and depends on the credibility of the pollster.
-Diagnostic checks, including residual and QQ plots, confirm that the model reasonably meets linear regression assumptions.
-
-Similarly, the MLR model for Donald Trump highlights that state-level differences and pollster transparency play a significant role in explaining his support.
-Figure @fig-mlr-trump illustrates these findings, and diagnostic evaluations confirm that the model assumptions (linearity, independence, and homoscedasticity) are sufficiently met.
+Recognizing the multifaceted nature of voter support, we extended our analysis using multiple linear regression (MLR) models.
+The initial model incorporated several predictors believed to influence polling outcomes:
+
+$$
+\text{pct} = \beta_0 + \beta_1 \times \text{numeric\_grade} + \beta_2 \times \text{pollscore} + \beta_3 \times \text{transparency\_score}
+$$
+$$ + \beta_4 \times \text{sample\_size} + \beta_5 \times \text{state} + \beta_6 \times \text{methodology} + \epsilon
+$$
+
+Variables were selected based on their relevance to polling accuracy and potential influence on voter preferences.
+Pollster ratings and transparency scores were included to account for differences in pollster reliability.
+State-level data were incorporated to capture regional variations in support.
+Sample size and methodology were included to control for their potential impact on polling results (see @fig-mlr-harris and @fig-mlr-trump as it illustrates residual plots and Q-Q plots).
 
@@
 
-Based on the VIF results in @tbl-harris-vif, we refined the model by removing less significant predictors (such as methodology and transparency_score) to reduce multi-collinearity.This improves the model's accuracy and interpretability.
+To assess multicollinearity among predictors, we calculated the Variance Inflation Factor (VIF) for each variable(@tbl-harris-vif).
+High VIF values suggest multicollinearity, which can compromise the stability of coefficient estimates.
+
+The initial VIF analysis revealed multicollinearity concerns, particularly with transparency score and methodology.
+To address this, we refined the model by removing less significant predictors with high VIF values, resulting in:
+$$
+\text{pct} = \beta_0 + \beta_1 \times \text{numeric\_grade} + \beta_2 \times \text{pollscore} + \beta_3 \times \text{sample\_size} + \beta_4 \times \text{state}
+$$
 
@@
 
+This refinement reduced VIF values to acceptable levels, improving the model's reliability and interpretability.
+
 ## Stepwise Model Selection
 
-To refine the multi-linear regression model, I applied a stepwise selection method to optimize the choice of predictor variables.
-Stepwise selection evaluates the model iteratively, adding or removing variables based on predefined criteria (typically AIC) to find a more parsimonious model.
-Using both forward and backward selection, stepwise method in the step() function assesses each predictor's contribution and determines if adding or removing it improves model fit.
-By allowing the function to examine variables in both directions, this process systematically removed non-significant variables that did not contribute meaningfully to the model's explanatory power, leaving only the most impactful predictors in the final refined model.
-The resulting plots, showing the "Residuals vs Fitted" and "Normal Q-Q" diagnostic views, allow for visual inspection of model assumptions and residual patterns, supporting the quality of the refined model.
+To optimize the model further, we applied stepwise selection using the Akaike Information Criterion (AIC) as the criterion for adding or removing predictors. The final model obtained is:
+
+$$
+\text{pct} = \beta_0 + \beta_1 \times \text{pollscore} + \beta_2 \times \log(\text{sample\_size}) + \beta_3 \times \text{state} + \epsilon
+$$
+
+The logarithmic transformation of sample size was introduced to account for diminishing returns in effect size with increasing sample sizes.
+
+The final model was selected based on statistical significance, model diagnostics, and theoretical considerations.
+Pollster rating and state remained significant predictors, consistent with expectations that pollster credibility and regional factors influence polling results.
+The logarithmic transformation improved model fit and addressed heteroscedasticity associated with sample size.
+
+All statistical analyses were conducted using [@citeR].
+The lm() function was used for regression modeling, and the car package [@car] was employed to calculate VIF values.
+Stepwise selection was performed using the step() function.
+Diagnostic plots were generated to assess model assumptions.
 
@@
 ```
+
+# Electoral College Prediction {#sec-predict}
+
 ```{r}
@@
 ```
-
-# Electoral College Prediction {#sec-predict}
 
@@
 
+# Load state abbreviations and coordinates
+state_coords <- data.frame(
+  state = state.name,
+  abbrev = state.abb,
+  region = tolower(state.name)
+) %>%
+  left_join(us_states %>% group_by(region) %>% summarise(long = mean(long), lat = mean(lat)), by = "region")
+
 # Simplify prediction data for mapping
@@
 
-# Create the basic map without projection
+# Create the map with state abbreviations
 ggplot(data = map_data_final) +
@@
                color = "white", size = 0.2) +
-  coord_fixed(1.3) +  # This replaces coord_map() with a simpler aspect ratio adjustment
+  geom_text(data = state_coords, aes(x = long, y = lat, label = abbrev), 
+            size = 3, color = "black") +  # Add state abbreviations
+  coord_fixed(1.3) +  # Simplified fixed aspect ratio
   scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080"),
@@
   )
-```
+
+```
+
+@fig-map-state-winner displays a predicted U.S.
+Electoral College outcome in a hypothetical race between Harris and Trump, with blue and red indicating states projected to support each candidate, respectively.
+Harris is expected to win in traditional Democratic strongholds like the West Coast, Northeast, and parts of the Midwest, areas historically aligned with Democratic candidates due to factors such as urbanization, higher education levels, and diverse populations.
+In contrast, Trump’s projected wins across the South and Great Plains reflect longstanding Republican preferences in these regions, shaped by conservative values and economic policies that resonate in rural and suburban areas.
+The distribution highlights both entrenched partisan divides and key battleground states, such as Florida and Pennsylvania, that could be decisive given their recent variability in party support.
 

### 2024-11-03 scripts/03-clean_data.R

@@ -9,10 +9,14 @@
 
 #### Workspace setup ####
 # Load necessary libraries
+
 library(dplyr)
 library(tidyverse)
 library(janitor)
 library(lubridate)
+library(arrow)
+library(here)
+
 
 ########################### Data Import and Cleaning########################### 
 # Read the dataset and clean column names
@@ -45,6 +49,6 @@ filtered_data <- data |>
   ) |>
   drop_na()
 
-# Save the cleaned dataset as a new CSV file
-write.csv(filtered_data, "data/02-analysis_data/analysis_data.csv", row.names = FALSE)
+write_parquet(filtered_data,
+  here("data", "02-analysis_data", "analysis_data.parquet"))
 

### 2024-11-03 paper/paper.qmd
Improved Abstract, Data, Intro, Prose, Table/Figures

@@
   - Tommy Fu
-thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git."
+thanks: "Code and data are available at: https://github.com/timchen0326/US_presidential_election_forecast_2024.git"
 date: today
 date-format: long
-abstract: "We forecast the outcome of the 2024 U.S.
-Presidential Election between Kamala Harris and Donald Trump by developing multiple linear regression models based on comprehensive polling data collected throughout the election cycle.
-Incorporating variables such as state-level demographics, pollster reliability scores, transparency scores, and sample sizes, we applied the same statistical model to both candidates to ensure consistent comparison.
-Our analysis predicts that Kamala Harris will receive 216 electoral votes, while Donald Trump is projected to secure 147 electoral votes.
-Neither candidate achieves the 270 electoral votes required to win the presidency, highlighting the potential for a closely contested election.
-These findings underscore the significant impact of state-specific factors on voter support and suggest that neither candidate currently holds a decisive advantage.
-We recommend that future research includes dynamic modeling techniques and additional predictive variables, such as economic indicators and voter turnout rates, to enhance the accuracy of election forecasts.
-Our study emphasizes the complexities involved in electoral predictions and the necessity of balancing multiple factors in policy design and electoral analysis."
+abstract: "We develop statistical models to forecast the 2024 U.S.
+presidential election outcome between Kamala Harris and Donald Trump using polling data from July through October 2024, incorporating pollster reliability metrics, sample sizes, and state-level effects to predict Electoral College results.
+Analyzing over 800 polls from diverse polling organizations, we find that pollster rating and state-level factors significantly influence reported candidate support, with our models predicting 306 electoral votes for Harris and 232 for Trump.
+Our analysis demonstrates that while national polling averages provide useful findings, state-specific predictions are essential for accurate Electoral College forecasting, particularly in battleground states where margins are historically narrow.
+These findings highlight the importance of incorporating methodological rigor in polling analysis and demonstrate how sophisticated statistical modeling can improve electoral predictions despite the challenges of modern polling."
 format: pdf
@@
 
-The 2024 United States presidential election presents unprecedented challenges for electoral forecasting.
-As the country navigates increasing political polarization and evolving voting patterns, the reliability of traditional polling methods has come under intense scrutiny [@viala2024].
-The task of predicting voter behavior in America's diverse electorate is complicated by numerous factors, including shifting public opinion, rapidly changing political landscapes, and varying levels of voter engagement across different demographic groups.
-
-Recent history has highlighted the complexities of election forecasting.
-The polling failures in 2016 and 2020—where polls significantly underestimated Republican support in key states—have prompted a fundamental reassessment of polling methodologies [@keeter2024].
-These challenges are particularly acute in swing states, where margins of victory are often razor-thin and can determine the outcome of the entire election.
-The American Association for Public Opinion Research (AAPOR) identified several critical factors contributing to these polling errors, including the underrepresentation of Republican voters and difficulties in predicting voter turnout patterns [@viala2024].
-
-Survey methodology plays a crucial role in addressing these challenges.
-Well-designed surveys require careful consideration of sampling strategies, questionnaire design, and data collection methods to ensure representative results.
-As Keeter [-@keeter2024] emphasizes, pollsters must now employ sophisticated weighting procedures and rigorous quality controls to overcome declining response rates and potential partisan non-response bias.
-Understanding the strengths and limitations of different polling approaches—from traditional probability sampling to newer online panels—is essential for accurate electoral forecasting.
-
-This paper develops statistical models to forecast the outcome of the 2024 presidential election between Kamala Harris and Donald Trump.
-By leveraging multi-linear regression models, we predict the percentage of support for each candidate across different states, incorporating key variables such as pollster rating, sample size, and state-level demographics.
-Through aggregating these state-level predictions, we simulate Electoral College outcomes to provide insights into each candidate's probability of securing the required 270 electoral votes.
-Our analysis also includes a detailed examination of [@YouGov] polling methodology and proposes an idealized survey approach that could enhance the accuracy of election forecasting.
+The United States presidential election of 2024 presents unique challenges for electoral forecasting, particularly following notable polling inaccuracies in recent elections.
+The 2016 and 2020 elections demonstrated significant underestimation of Republican support in key states, leading to decreased public confidence in traditional polling methods.
+Research by the American Association for Public Opinion Research has identified several contributing factors, including difficulties in obtaining representative samples and predicting voter turnout amid increasing political polarization [@keeter2024].
+
+Contemporary polling organizations have implemented methodological improvements to address these challenges.
+As documented by Keeter [-@keeter2024], modern electoral polling requires sophisticated weighting procedures and enhanced quality control mechanisms to counteract declining response rates and partisan non-response bias.
+These refinements are especially important in swing states, where narrow margins typically determine electoral outcomes and where previous polling errors have had the most significant impact [@viala2024].
+This methodological evolution reflects the polling industry's adaptation to changing communication patterns and voter behavior.
+
+This paper aims to forecast the winner of the 2024 United States presidential election between Kamala Harris and Donald Trump by predicting each candidate's Electoral College vote total.
+Specifically, our estimand is the number of Electoral College votes each candidate will receive, calculated by converting state-level popular vote predictions into electoral votes through each state's winner-take-all system.
+Using polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database collected between July and October 2024 [@FiveThirtyEightPollingData2024], we develop multi-linear regression models that incorporate pollster reliability metrics, sample sizes, and state-specific effects.
+Our analysis predicts that Kamala Harris will receive 306 electoral votes compared to Donald Trump's 232, surpassing the 270-vote threshold required for victory.
+We also provide an evaluation of YouGov's [@YouGov] polling methodology and present an idealized survey approach for future election forecasting.
 
 The remainder of this paper is structured as follows.
-@sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy.
-@sec-model outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles.
-@sec-predict presents our Electoral College predictions based on the model outputs.
-@sec-discuss discusses the implications of our findings and suggests directions for future research.
-Finally, @sec-appendix evaluates YouGov's polling methodology, and our idealized survey methodology.
+@sec-data examines our dataset and key variables.
+@sec-model details our statistical approach and model development.
+@sec-predict presents electoral college projections.
+@sec-discuss explores implications and limitations.
+@sec-appendix provide an in-depth evaluation of polling methodologies and our proposed ideal survey design.
 
@@
 ## Overview
-Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news.
-This database aggregates and standardizes polling results from various organizations, applying quality metrics and methodological assessments to each poll.
-In this section, we detail our selected variables, discuss key measurements, outline important limitations of our data, and our data cleaning process.
-
-Our analysis focuses on several key variables that directly influence polling accuracy and electoral predictions. @tbl-overview_data presents a sample of our dataset.
-
-## Poll Quality Variables
-
-@tbl-poll-metrics summarizes three primary metrics that assess poll quality in our dataset:
-
-- **Pollscore**: Represents "the error and bias attributable to a pollster, where negative numbers indicate better performance." Our data shows a mean pollscore of -1.06 (SD = 0.28), with values ranging from -1.50 to -0.50, indicating generally high-quality polling organizations in our sample.
-- **Numeric Grade**: A rating given to pollsters to indicate their quality and reliability.
-With a mean of 2.89 (SD = 0.10) and a range of 2.70 to 3.00, our dataset maintains high standards.
-We established a minimum threshold of 2.7 to ensure methodological rigor.
-- **Transparency Score**: A measure reflecting pollsters' transparency about their methodology, with 10 being the highest possible score.
-The data shows strong overall transparency with a mean of 8.59 (SD = 1.04), indicating that most polls in our sample maintain high standards of methodological disclosure.
-
-## Methodological Variables
-
-- **Methodology**: @fig-methodology-dist shows the distribution of polling approaches used to conduct the polls.
-Live Phone methods and Online Panels are the predominant methods, with hybrid approaches combining multiple methods (such as IVR/Online Panel/Text-to-Web) representing evolving polling techniques adapting to changing communication patterns.
-- **Sample Size**: @fig-sample-size illustrates the distribution of poll sample sizes, representing the total number of respondents participating in each poll.
-Our dataset shows considerable variation, with a mean of 1,114.76 respondents (SD = 583.72).
-The right-skewed distribution ranges from 450 to 4,253 respondents, with larger sample sizes typically associated with national polls.
-
-## Geographic and Temporal Variables
-
-- **State Coverage**: @tbl-state-polls reveals the distribution of polls across different U.S.
-states.
-National polls lead with 261 surveys, followed by concentrated polling in key battleground states: Pennsylvania (86), Wisconsin (82), and North Carolina (64).
-This distribution reflects strategic focus on states where the electoral outcome is less certain.
-- **End Date**: @fig-temporal-polling tracks polling frequency from July through October 2024.
-We observe increased activity during September and early October, with weekly poll counts peaking in late September.
-This pattern reflects intensified polling efforts as the election approaches.
+This study analyzes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024].
+The database, maintained by ABC News, provides standardized polling results from diverse polling organizations, with each poll receiving rigorous quality assessments and methodological evaluations.
+We examine both national and state-level polls conducted between July and October 2024, focusing on the presidential race between Kamala Harris and Donald Trump.
+@tbl-overview_data in the appendix presents a sample of our dataset.
 
 ## Outcome Variable
-
-- **Support Percentage**: @fig-candidate-support displays our primary dependent variable: the percentage of respondents expressing support for each candidate.
-The distribution reveals a competitive race between Donald Trump and Kamala Harris, with support levels typically ranging between 40-50% for both candidates.
+Our primary dependent variable is candidate support percentage (`pct`), representing voter preferences between the two major candidates.
+@fig-candidate-support illustrates the distribution of support levels, indicating a competitive race where both Harris and Trump consistently receive between 40-50% support across polls.
+The relatively narrow range of these percentages underscores the importance of methodological precision in detecting meaningful differences in voter preference.
+
+## Predictor Variables
+We identify several important predictors that influence polling accuracy and electoral predictions. These variables fall into two categories: poll quality metrics and survey characteristics.
+
+### Poll Quality Metrics
+@tbl-poll-metrics summarizes three fundamental quality metrics in our dataset:
+
+- **Pollscore**: A measure of polling organization performance, where negative values indicate better accuracy.
+Our data shows a mean pollscore of -1.06 (SD = 0.28, range: -1.50 to -0.50), suggesting consistently high-quality polling organizations in our sample.
+- **Numeric Grade**: An evaluation of pollster reliability on a standardized scale.
+With a mean of 2.89 (SD = 0.10, range: 2.70 to 3.00), our dataset maintains strong methodological standards.
+We established a minimum threshold of 2.7 to ensure rigor.
+- **Transparency Score**: A quantification of methodological disclosure (maximum: 10).
+Our sample demonstrates strong transparency practices with a mean of 8.59 (SD = 1.04), indicating thorough methodological documentation.
+
+### Survey Characteristics
+- **Methodology**: As shown in @fig-methodology-dist, polling approaches vary significantly.
+Traditional live phone interviews and online panels dominate, while hybrid methods combining multiple approaches (e.g., IVR/Online Panel/Text-to-Web) reflect adaptation to changing communication patterns.
+- **Sample Size**: @fig-sample-size illustrates substantial variation in respondent numbers, with polls averaging 1,114.76 respondents (SD = 583.72, range: 450-4,253).
+The right-skewed distribution reflects larger samples in national polls, with state-level polls typically using smaller, targeted samples.
+- **State**: @tbl-state-polls shows the distribution of polls across states, with 261 national surveys complemented by intensive polling in battleground states: Pennsylvania (86 polls), Wisconsin (82), and North Carolina (64).
+This distribution reflects strategic focus on electorally competitive regions.
+- **End Date**: @fig-temporal-polling tracks polling frequency over time, showing increased activity during September and early October 2024, with peak weekly frequencies in late September.
+This pattern captures intensified data collection as Election Day approaches.
 
@@
 - **Temporal Dynamics:** Our dataset provides discrete snapshots of voter preferences rather than continuous measurement.
-This limitation is particularly relevant given the rapid evolution of political narratives and voter sentiment during presidential campaigns.
+This limitation is particularly relevant given the fast evolution of political narratives and voter sentiment during presidential campaigns.
 - **Geographic Coverage:** Although we have national and state-level polling data, coverage varies by state.
@@
 
-## Cleaning Process and Analysis
-
-The data cleaning process employed R [@citeR] along with several specialized packages: tidyverse [@tidyverse] for data manipulation, dplyr [@dplyr] for data transformation, janitor [@janitor] for consistent naming conventions, and lubridate [@lubridate] for date handling.
-
-Our cleaning process followed several key steps:
-
-1. Filtered for high-quality polls using a minimum threshold (*numeric_grade* $≥$ 2.7)
-2. Limited temporal coverage to post-campaign announcement period (after July 21, 2024)
-3. Standardized geographic data:
-
-  > i. Coded missing state information as "National" polls
-  > ii. Verified state names for consistency
-
-4. Transformed key variables:
-
-  > i. Converted *End Date* values to standardized date format
-  > ii. Calculated absolute supporter numbers from percentages and sample sizes
-  > iii. Created binary candidate indicators (Harris = 1, Trump = 0)
-
-Other variable such as ... were not selected for our analysis because ...
-
-Our variable selection process prioritized measures essential for polling accuracy and electoral prediction while eliminating redundant or non-informative fields.
-We retained key poll quality metrics including *pollscore* and *numeric_grade*, which provide crucial information about the reliability of each survey.
-Sample characteristics such as *sample size* and *methodology* were preserved to account for differences in polling precision.
-*State* and polling *end_date* information were maintained to capture regional variations and time-dependent patterns in voter preferences.
-
-We excluded several variables that offered little additional analytical value, such as *population_full*, which duplicated information available in other fields, and administrative data that did not directly influence predictions.
-This focused approach to variable selection allowed us to preserve all information necessary for accurate electoral forecasting.
+## Data Processing and Analysis
+
+Our analysis employs R [@citeR] with specialized packages: tidyverse [@tidyverse] for data manipulation, dplyr [@dplyr] for transformation, janitor [@janitor] for naming standardization, and lubridate [@lubridate] for date handling.
+The data cleaning process, detailed in @sec-datacleaning, involved filtering for high-quality polls (numeric_grade $≥$ 2.7), standardizing geographic information, and creating consistent date formats and candidate indicators.
+
+Our variable selection prioritized metrics that directly influence polling accuracy and electoral forecasting while excluding variables that could introduce noise or redundancy.
+We retained `pollscore` and `numeric_grade` as established measures of polling reliability, while excluding related but less precise metrics such as `pollster_rating_name`.
+Geographic and temporal variables (`state`, `end_date`) were essential for capturing regional variations and time-series patterns.
+We deliberately omitted variables that could introduce partisan bias (e.g., `sponsor_candidate`, `endorsed_candidate_name`) or duplicate existing information (e.g., `population_full`, `display_name`).
+Administrative identifiers (`poll_id`, `race_id`) and URLs were excluded as they didn't serve analytical purposes.
+This focused selection maintains analytical rigor while avoiding the risks of over-parameterization.
 
@@
 
-Recognizing the multifaceted nature of voter support, we extended our analysis using multiple linear regression (MLR) models.
+Recognizing the complex nature of voter support, we extended our analysis using multiple linear regression (MLR) models.
 The initial model incorporated several predictors believed to influence polling outcomes:
@@
 $$
-$$ + \beta_4 \times \text{sample\_size} + \beta_5 \times \text{state} + \beta_6 \times \text{methodology} + \epsilon
+
+$$ 
++ \beta_4 \times \text{sample\_size} + \beta_5 \times \text{state} + \beta_6 \times \text{methodology} + \epsilon
 $$
@@
 
-The initial VIF analysis revealed multicollinearity concerns, particularly with transparency score and methodology.
+The initial VIF analysis demonstrates multicollinearity concerns, particularly with transparency score and methodology.
 To address this, we refined the model by removing less significant predictors with high VIF values, resulting in:
+
 $$
@@
 We then trained the models on the training subset and generated predictions for the test subset, evaluating model accuracy using the Root Mean Squared Error (RMSE).
-The RMSE metric quantifies the average prediction error in the test set, offering an indication of each model’s reliability in predicting future national poll outcomes.
+The RMSE metric quantifies the average prediction error in the test set, providing an indication of each model’s reliability in predicting future national poll outcomes.
 The Harris model yielded an RMSE of 3.12, while the Trump model’s RMSE was 2.40, indicating that both models provide reasonably accurate predictions, with the Trump model demonstrating slightly lower average error in predicting national polling support.
@@
 
-# Melt data for easier plotting
-prediction_melted <- melt(
-  prediction_comparison, 
-  id.vars = c("state", "electoral_votes", "winner"),
-  measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
-)
-
-# Rename melted variables for readability
-colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
-prediction_melted$candidate <- ifelse(
-  prediction_melted$candidate == "predicted_pct_harris", 
-  "Harris", 
-  "Trump"
-)
-
-# Calculate the midpoint to split the states
-n_states <- length(unique(prediction_melted$state))
-midpoint <- ceiling(n_states / 2)
-
-# Order states by electoral votes in descending order for better visualization
-states_order <- prediction_melted %>%
-  distinct(state, electoral_votes) %>%
-  arrange(desc(electoral_votes)) %>%
-  pull(state)
-
-# Split the states into two groups
-left_states <- states_order[1:midpoint]
-right_states <- states_order[(midpoint + 1):length(states_order)]
-
-# Create two separate datasets
-prediction_left <- prediction_melted %>%
-  filter(state %in% left_states) %>%
-  mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
-
-prediction_right <- prediction_melted %>%
-  filter(state %in% right_states) %>%
-  mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
-
-# Function to create individual plots to avoid repetition
-create_plot <- function(data, show_legend = FALSE) {
-  ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
-    geom_bar(
-      stat = "identity", 
-      position = position_dodge(width = 0.9), 
-      width = 0.5  # Reduced bar width for more spacing
-    ) +
-    coord_flip() +
-    labs(
-      x = NULL,
-      y = "Predicted Vote Percentage (%)"
-    ) +
-    scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
-    theme_minimal() +
-    theme(
-      text = element_text(size = 9),  # Reduced overall text size
-      axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
-      axis.text.x = element_text(size = 9),
-      legend.title = element_blank(),
-      legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
-      legend.text = element_text(size = 9),
-      panel.grid.major.y = element_blank(),
-      plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
-    ) +
-    geom_text(
-      aes(label = round(value, 1)),
-      position = position_dodge(width = 0.9),
-      hjust = -0.2,    # Adjusted horizontal position for labels
-      size = 1.5,      # Smaller label size
-      show.legend = FALSE
-    ) +
-    scale_y_continuous(
-      expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
-      limits = c(0, 75)  # Adjust based on your data
-    )
-}
-
-# Create the left plot with legend
-plot_left <- create_plot(prediction_left, show_legend = TRUE)
-
-# Create the right plot without legend
-plot_right <- create_plot(prediction_right, show_legend = FALSE)
-
-# Combine the plots side by side with legend on top
-combined_plot <- plot_left + plot_right +
-  plot_layout(guides = "collect") &  # Collects the legend at the top
-  theme(legend.position = "top")  # Places the combined legend at the top
-
-# Print the combined plot
-combined_plot
+# # Melt data for easier plotting
+# prediction_melted <- melt(
+#   prediction_comparison, 
+#   id.vars = c("state", "electoral_votes", "winner"),
+#   measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
+# )
+# 
+# # Rename melted variables for readability
+# colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
+# prediction_melted$candidate <- ifelse(
+#   prediction_melted$candidate == "predicted_pct_harris", 
+#   "Harris", 
+#   "Trump"
+# )
+# 
+# # Calculate the midpoint to split the states
+# n_states <- length(unique(prediction_melted$state))
+# midpoint <- ceiling(n_states / 2)
+# 
+# # Order states by electoral votes in descending order for better visualization
+# states_order <- prediction_melted %>%
+#   distinct(state, electoral_votes) %>%
+#   arrange(desc(electoral_votes)) %>%
+#   pull(state)
+# 
+# # Split the states into two groups
+# left_states <- states_order[1:midpoint]
+# right_states <- states_order[(midpoint + 1):length(states_order)]
+# 
+# # Create two separate datasets
+# prediction_left <- prediction_melted %>%
+#   filter(state %in% left_states) %>%
+#   mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
+# 
+# prediction_right <- prediction_melted %>%
+#   filter(state %in% right_states) %>%
+#   mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
+# 
+# # Function to create individual plots to avoid repetition
+# create_plot <- function(data, show_legend = FALSE) {
+#   ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
+#     geom_bar(
+#       stat = "identity", 
+#       position = position_dodge(width = 0.9), 
+#       width = 0.5  # Reduced bar width for more spacing
+#     ) +
+#     coord_flip() +
+#     labs(
+#       x = NULL,
+#       y = "Predicted Vote Percentage (%)"
+#     ) +
+#     scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
+#     theme_minimal() +
+#     theme(
+#       text = element_text(size = 9),  # Reduced overall text size
+#       axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
+#       axis.text.x = element_text(size = 9),
+#       legend.title = element_blank(),
+#       legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
+#       legend.text = element_text(size = 9),
+#       panel.grid.major.y = element_blank(),
+#       plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
+#     ) +
+#     geom_text(
+#       aes(label = round(value, 1)),
+#       position = position_dodge(width = 0.9),
+#       hjust = -0.2,    # Adjusted horizontal position for labels
+#       size = 1.5,      # Smaller label size
+#       show.legend = FALSE
+#     ) +
+#     scale_y_continuous(
+#       expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
+#       limits = c(0, 75)  # Adjust based on your data
+#     )
+# }
+# 
+# # Create the left plot with legend
+# plot_left <- create_plot(prediction_left, show_legend = TRUE)
+# 
+# # Create the right plot without legend
+# plot_right <- create_plot(prediction_right, show_legend = FALSE)
+# 
+# # Combine the plots side by side with legend on top
+# combined_plot <- plot_left + plot_right +
+#   plot_layout(guides = "collect") &  # Collects the legend at the top
+#   theme(legend.position = "top")  # Places the combined legend at the top
+# 
+# # Print the combined plot
+# combined_plot
 
@@
 **Calculation of Total Electoral Votes**: We summed the electoral votes from all states won by each candidate to compute their total electoral vote counts.
-This calculation is critical, as securing at least 270 electoral votes is necessary to win the presidency under the U.S.
+This calculation is necessary, as securing at least 270 electoral votes is necessary to win the presidency under the U.S.
 electoral system.
@@
 This methodology reflects the structure of the U.S.
-electoral system and offers valuable insights into how predicted voter support may translate into electoral success.
+electoral system and provides valuable results into how predicted voter support may translate into electoral success.
 
@@
 
-## Insights on Polling Methodologies and Voter Support
+## Polling Methodologies and Voter Support
 
@@
 
-The primary strength of YouGov's methodology lies in its efficiency and cost-effectiveness, enabling rapid data collection.
+The primary strength of YouGov's methodology lies in its efficiency and cost-effectiveness, enabling quick data collection.
 However, the reliance on non-probability sampling can introduce biases, which are partially mitigated by their rigorous weighting process.
-While effective in providing timely insights, this method requires careful consideration of its limitations in representativeness.
+While effective in providing timely findings, this method requires careful consideration of its limitations in representativeness.
 
@@
 ### Strengths and Limitations
-The key strengths of YouGov’s methodology are its rapid data collection, cost-effectiveness, and capacity to track public opinion trends over time.
+The key strengths of YouGov’s methodology are its efficient data collection, cost-effectiveness, and capacity to track public opinion trends over time.
 However, the reliance on non-probability sampling can introduce biases, and the online-only format may lead to underrepresentation of certain populations.
@@
 It balances statistical rigor with practicality to ensure accurate predictions of both the popular vote and Electoral College outcomes.
-The methodology includes advanced sampling techniques, robust survey design, and comprehensive data analysis to capture key demographic and political variables.
-By addressing potential biases and focusing on cost-effective data collection, this plan ensures reliable and representative insights into voter behavior.
+The methodology includes developed sampling techniques, robust survey design, and detailed data analysis to capture key demographic and political variables.
+By addressing potential biases and focusing on cost-effective data collection, this plan ensures reliable and representative results into voter behavior.
 
@@
 ### Recruitment Plan
-To recruit respondents effectively, we will leverage online panels, social media advertising, and partnerships with community organizations to reach a diverse audience.
+To recruit respondents effectively, we will utilize online panels, social media advertising, and partnerships with community organizations to reach a diverse audience.
 Offering modest incentives, such as $5 digital gift cards, encourages participation while managing costs.
@@
 This methodology presents a feasible and thorough plan to forecast the US presidential election within the specified budget.
-By adhering to best practices in survey design and execution, and by carefully considering both the popular vote and electoral college implications, the survey aims to provide accurate and reliable insights into voter intentions.
+By adhering to best practices in survey design and execution, and by carefully considering both the popular vote and electoral college implications, the survey aims to provide accurate and reliable results into voter intentions.
 
@@
 # Appendix
+
+## Data Processing Steps {#sec-datacleaning}
+Our cleaning process followed several key steps:
+
+1. Filtered for high-quality polls using a minimum threshold (*numeric_grade* $≥$ 2.7)
+2. Limited temporal coverage to post-campaign announcement period (after July 21, 2024)
+3. Standardized geographic data:
+
+  > i. Coded missing state information as "National" polls
+  > ii. Verified state names for consistency
+
+4. Transformed key variables:
+
+  > i. Converted *End Date* values to standardized date format
+  > ii. Calculated absolute supporter numbers from percentages and sample sizes
+  > iii. Created binary candidate indicators (Harris = 1, Trump = 0)
+
+## Data Tables and Figures
 
@@
 
+colnames(data_overview) <- paste("Row", 1:ncol(data_overview))
+
 # Display the transposed table
-kable(data_overview)
+kable(data_overview)%>%
+  kable_styling(font_size = 8) %>%
+  column_spec(1, width = "2.5cm") %>%
+  column_spec(2:ncol(data_overview), width = "2cm")
 ```
@@
 
-The table also includes data on candidate support percentages, with Donald Trump polling between 42% and 48% across various states and Kamala Harris receiving 48% in Michigan.
-This snapshot reveals the diversity in poll designs, geographical focus, and methodological rigor, which are crucial for understanding the variability in polling data.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-candidate-support
+#| fig-cap: Distribution of Support Percentages by Candidate
+
+ggplot(data, aes(x = pct, fill = candidate_name)) +
+  geom_histogram(bins = 30, position = "identity") +
+  facet_wrap(~ candidate_name) +
+  labs(x = "Percentage of Support",
+       y = "Count") +
+  scale_fill_discrete(name = "Candidates") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+@fig-candidate-support shows the distribution of support percentages for Donald Trump and Kamala Harris.
+Both distributions are centered around the mid-40% to mid-50% range, reflecting a competitive race.
+Trump's support is more concentrated around the 45-50% range, indicating relatively consistent polling results.
+Harris’s distribution is slightly more varied, suggesting greater fluctuation in her polling support.
+This visualization highlights the close contest between the two candidates, with neither consistently achieving overwhelming support across polls.
 
@@
 #| warning: false
-#| label: fig-candidate-support
-#| fig-cap: Distribution of Support Percentages by Candidate
-
-ggplot(data, aes(x = pct, fill = candidate_name)) +
-  geom_histogram(bins = 30, position = "identity") +
-  facet_wrap(~ candidate_name) +
-  labs(x = "Percentage of Support",
+#| label: fig-sample-size
+#| fig-cap: Distribution of Poll Sample Sizes
+
+data %>%
+  ggplot(aes(x = sample_size)) +
+  geom_histogram(binwidth = 100) +
+  theme_minimal() +
+  labs(x = "Sample Size",
        y = "Count") +
-  scale_fill_discrete(name = "Candidates") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-```
-@fig-candidate-support shows the distribution of support percentages for Donald Trump and Kamala Harris.
-Both distributions are centered around the mid-40% to mid-50% range, reflecting a competitive race.
-Trump's support is more concentrated around the 45-50% range, indicating relatively consistent polling results.
-Harris’s distribution is slightly more varied, suggesting greater fluctuation in her polling support.
-This visualization highlights the close contest between the two candidates, with neither consistently achieving overwhelming support across polls.
+  scale_x_continuous(breaks = seq(0, max(data$sample_size), by = 500))
+```
+
+@fig-sample-size shows the distribution of sample sizes across the polls in the dataset.
+The majority of polls have sample sizes clustered around 1,000 respondents, which is a typical size for achieving a balance between precision and cost.
+A few polls have significantly larger sample sizes, extending beyond 2,000, with some even exceeding 4,000 respondents.
+This suggests that while most polls aim for a moderate sample size, there are instances where more extensive sampling is conducted, possibly for greater accuracy or in significant regions.
+The distribution reflects the common practice of balancing statistical reliability with logistical constraints.
 
@@
   kable(format = "latex", booktabs = TRUE)
-```
+
+```
+
 @tbl-state-polls illustrates the number of polls conducted over time, showing weekly fluctuations in polling activity.
 There is a noticeable increase in the number of polls around mid-September, peaking at over 160 polls in a week, followed by a decline as the election date approaches.
-This pattern suggests heightened polling activity during key points in the campaign, likely driven by major events or debates, before tapering off closer to the election.
+This pattern suggests heightened polling activity during key points in the campaign, likely caused by major events or debates, before tapering off closer to the election.
 The variability highlights how polling frequency can change in response to the evolving political landscape.
@@
 National polls dominate with 261 instances, while key battleground states like Pennsylvania (86 polls), Wisconsin (82 polls), and North Carolina (64 polls) also see significant attention.
-These states are critical for determining election outcomes, explaining the higher frequency.
+These states are important for determining election outcomes, explaining the higher frequency.
 Conversely, states like Missouri, Montana, and Nebraska show fewer polls (4 each), reflecting less competitive races or a perceived lower impact on the overall election result.
 This distribution underscores the strategic focus on swing states in polling efforts.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-sample-size
-#| fig-cap: Distribution of Poll Sample Sizes
-
-data %>%
-  ggplot(aes(x = sample_size)) +
-  geom_histogram(binwidth = 100) +
-  theme_minimal() +
-  labs(x = "Sample Size",
-       y = "Count") +
-  scale_x_continuous(breaks = seq(0, max(data$sample_size), by = 500))
-```
-
-@fig-sample-size shows the distribution of sample sizes across the polls in the dataset.
-The majority of polls have sample sizes clustered around 1,000 respondents, which is a typical size for achieving a balance between precision and cost.
-A few polls have significantly larger sample sizes, extending beyond 2,000, with some even exceeding 4,000 respondents.
-This suggests that while most polls aim for a moderate sample size, there are instances where more extensive sampling is conducted, possibly for greater accuracy or in critical regions.
-The distribution reflects the common practice of balancing statistical reliability with logistical constraints.
 

### 2024-11-03 paper/paper.qmd
changed premable on all documents as well as adding llm usage, and updated readme file

@@
 
-# # Melt data for easier plotting
-# prediction_melted <- melt(
-#   prediction_comparison, 
-#   id.vars = c("state", "electoral_votes", "winner"),
-#   measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
-# )
-# 
-# # Rename melted variables for readability
-# colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
-# prediction_melted$candidate <- ifelse(
-#   prediction_melted$candidate == "predicted_pct_harris", 
-#   "Harris", 
-#   "Trump"
-# )
-# 
-# # Calculate the midpoint to split the states
-# n_states <- length(unique(prediction_melted$state))
-# midpoint <- ceiling(n_states / 2)
-# 
-# # Order states by electoral votes in descending order for better visualization
-# states_order <- prediction_melted %>%
-#   distinct(state, electoral_votes) %>%
-#   arrange(desc(electoral_votes)) %>%
-#   pull(state)
-# 
-# # Split the states into two groups
-# left_states <- states_order[1:midpoint]
-# right_states <- states_order[(midpoint + 1):length(states_order)]
-# 
-# # Create two separate datasets
-# prediction_left <- prediction_melted %>%
-#   filter(state %in% left_states) %>%
-#   mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
-# 
-# prediction_right <- prediction_melted %>%
-#   filter(state %in% right_states) %>%
-#   mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
-# 
-# # Function to create individual plots to avoid repetition
-# create_plot <- function(data, show_legend = FALSE) {
-#   ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
-#     geom_bar(
-#       stat = "identity", 
-#       position = position_dodge(width = 0.9), 
-#       width = 0.5  # Reduced bar width for more spacing
-#     ) +
-#     coord_flip() +
-#     labs(
-#       x = NULL,
-#       y = "Predicted Vote Percentage (%)"
-#     ) +
-#     scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
-#     theme_minimal() +
-#     theme(
-#       text = element_text(size = 9),  # Reduced overall text size
-#       axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
-#       axis.text.x = element_text(size = 9),
-#       legend.title = element_blank(),
-#       legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
-#       legend.text = element_text(size = 9),
-#       panel.grid.major.y = element_blank(),
-#       plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
-#     ) +
-#     geom_text(
-#       aes(label = round(value, 1)),
-#       position = position_dodge(width = 0.9),
-#       hjust = -0.2,    # Adjusted horizontal position for labels
-#       size = 1.5,      # Smaller label size
-#       show.legend = FALSE
-#     ) +
-#     scale_y_continuous(
-#       expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
-#       limits = c(0, 75)  # Adjust based on your data
-#     )
-# }
-# 
-# # Create the left plot with legend
-# plot_left <- create_plot(prediction_left, show_legend = TRUE)
-# 
-# # Create the right plot without legend
-# plot_right <- create_plot(prediction_right, show_legend = FALSE)
-# 
-# # Combine the plots side by side with legend on top
-# combined_plot <- plot_left + plot_right +
-#   plot_layout(guides = "collect") &  # Collects the legend at the top
-#   theme(legend.position = "top")  # Places the combined legend at the top
-# 
-# # Print the combined plot
-# combined_plot
+# Melt data for easier plotting
+prediction_melted <- melt(
+  prediction_comparison, 
+  id.vars = c("state", "electoral_votes", "winner"),
+  measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
+)
+
+# Rename melted variables for readability
+colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
+prediction_melted$candidate <- ifelse(
+  prediction_melted$candidate == "predicted_pct_harris", 
+  "Harris", 
+  "Trump"
+)
+
+# Calculate the midpoint to split the states
+n_states <- length(unique(prediction_melted$state))
+midpoint <- ceiling(n_states / 2)
+
+# Order states by electoral votes in descending order for better visualization
+states_order <- prediction_melted %>%
+  distinct(state, electoral_votes) %>%
+  arrange(desc(electoral_votes)) %>%
+  pull(state)
+
+# Split the states into two groups
+left_states <- states_order[1:midpoint]
+right_states <- states_order[(midpoint + 1):length(states_order)]
+
+# Create two separate datasets
+prediction_left <- prediction_melted %>%
+  filter(state %in% left_states) %>%
+  mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
+
+prediction_right <- prediction_melted %>%
+  filter(state %in% right_states) %>%
+  mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
+
+# Function to create individual plots to avoid repetition
+create_plot <- function(data, show_legend = FALSE) {
+  ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
+    geom_bar(
+      stat = "identity", 
+      position = position_dodge(width = 0.9), 
+      width = 0.5  # Reduced bar width for more spacing
+    ) +
+    coord_flip() +
+    labs(
+      x = NULL,
+      y = "Predicted Vote Percentage (%)"
+    ) +
+    scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
+    theme_minimal() +
+    theme(
+      text = element_text(size = 9),  # Reduced overall text size
+      axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
+      axis.text.x = element_text(size = 9),
+      legend.title = element_blank(),
+      legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
+      legend.text = element_text(size = 9),
+      panel.grid.major.y = element_blank(),
+      plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
+    ) +
+    geom_text(
+      aes(label = round(value, 1)),
+      position = position_dodge(width = 0.9),
+      hjust = -0.2,    # Adjusted horizontal position for labels
+      size = 1.5,      # Smaller label size
+      show.legend = FALSE
+    ) +
+    scale_y_continuous(
+      expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
+      limits = c(0, 75)  # Adjust based on your data
+    )
+}
+
+# Create the left plot with legend
+plot_left <- create_plot(prediction_left, show_legend = TRUE)
+
+# Create the right plot without legend
+plot_right <- create_plot(prediction_right, show_legend = FALSE)
+
+# Combine the plots side by side with legend on top
+combined_plot <- plot_left + plot_right +
+  plot_layout(guides = "collect") &  # Collects the legend at the top
+  theme(legend.position = "top")  # Places the combined legend at the top
+
+# Print the combined plot
+combined_plot
 

### 2024-11-03 scripts/03-clean_data.R

@@ -1,11 +1,15 @@
 #### Preamble ####
-# Purpose: Prepare polling data and build a linear regression model to predict popular vote for Kamala Harris and Donald Trump.
-# Author: [Your Name]
-# Date: [Today's Date]
-# Contact: [Your Email]
-# License: MIT
-# Pre-requisites: Requires 'president_polls.csv' file in 'data/01-raw_data/'
-# Any other information needed? This script builds an MLR model for Harris vs Trump using polling data.
+# Purpose: Prepare and filter polling data to build a multiple linear regression model predicting popular vote percentages for Kamala Harris and Donald Trump.
+# Author: Tim Chen, Steven Li, Tommy Fu
+# Date: 3 November 2024
+# Contact: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - Ensure the 'president_polls.csv' file is located in 'data/01-raw_data/'
+# - The following R packages must be installed: 'dplyr', 'tidyverse', 'janitor', 'lubridate', 'arrow', 'here'
+# - Polling data will be filtered for high-quality entries and limited to Harris and Trump, with necessary data transformations for modeling.
 
 #### Workspace setup ####
 # Load necessary libraries

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -17,6 +17,7 @@ library(tidyverse)
 library(testthat)
 library(lubridate)
 
+
 # Read in the data
 pollster_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 

### 2024-11-03 scripts/05-exploratory_data_analysis.R

@@ -1,10 +1,15 @@
 #### Preamble ####
 # Purpose: Perform Exploratory Data Analysis (EDA) on polling data for Kamala Harris and Donald Trump, and save the plots as images.
-# Author: [Your Name]
-# Date: [Today's Date]
-# Contact: [Your Email]
-# License: MIT
-# Pre-requisites: Requires cleaned 'analysis_data.csv' with filtered data for Harris and Trump
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 3 November 2024
+# Contacts: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - Requires cleaned 'analysis_data.csv' with filtered data for Harris and Trump in the 'data/02-analysis_data/' directory
+# - The following R packages must be installed: 'dplyr', 'knitr', 'caret', 'tidyverse', 'Metrics'
+# - This script performs model training on national polling data for both candidates and calculates RMSE performance metrics
 
 # Install and load the caret package if you haven't already
 install.packages("caret")

### 2024-11-03 scripts/06-model_data.R

@@ -1,11 +1,15 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Build multiple linear regression (MLR) models to predict polling percentages for Kamala Harris and Donald Trump.
+# Authors: Tim Chen, Steven Li, Tommy Fu
+# Date: 3 November 2024
+# Contacts: 
+# - Tim Chen: timwt.chen@mail.utoronto.ca
+# - Steven Li: stevency.li@mail.utoronto.ca
+# - Tommy Fu: tommy.fu@mail.utoronto.ca
+# Pre-requisites: 
+# - Requires the 'analysis_data.csv' file located in 'data/02-analysis_data/'
+# - The following R packages must be installed: 'tidyverse', 'rstanarm'
+# - This script builds and refines MLR models for both candidates using relevant predictors and saves the models as .rds files for future use.
 
 
 #### Workspace setup ####

### 2024-11-04 paper/paper.qmd
Final appendix edits

@@
 ```
-
+@fig-harris-samplesize-supportpct displays a scatter plot with a linear regression line showing the relationship between sample size and the percentage of support for Kamala Harris.
+The plot suggests a very weak positive correlation, as indicated by the near-flat regression line.
+This implies that variations in sample size have a minimal effect on the percentage of support for Harris.
+The spread of data points shows a consistent range of support percentages across different sample sizes, reinforcing the observation that larger sample sizes do not significantly alter her support levels in the polls.
 ```{r}
@@
 ```
+@fig-trump-samplesize-supportpct shows a weak negative relationship between poll sample size and Trump's support percentage, as indicated by the slight downward slope of the regression line.
+While larger sample sizes are associated with a marginal decrease in reported support, the scattered distribution of data points suggests considerable variability, implying that other factors like pollster methodology or regional differences significantly influence the results.
+The narrow confidence interval around the trend line reflects a modest level of certainty in this weak relationship, though the presence of outliers highlights exceptions where larger samples do not necessarily align with the general trend.
 ```{r}
@@
 ```
+@fig-mlr-harris The **Residuals vs Fitted** plot (left) for Kamala Harris's multi-linear regression model indicates that most residuals are evenly distributed around the horizontal axis (residuals = 0), suggesting that the model's predictions are unbiased across different fitted values.
+However, the slight curve near the middle of the plot implies potential non-linearity or missing variables.
+Additionally, some points (e.g., labeled 263, 352, 370) are noticeable outliers, which may affect the model's accuracy.
+
+The **Q-Q plot** (right) shows the standardized residuals against the theoretical quantiles.
+The residuals mostly follow the straight line, indicating that they are approximately normally distributed.
+However, deviations at both tails (particularly on the left) suggest the presence of some extreme values, which could impact the normality assumption and model reliability.
 
@@
 ```
+@fig-mlr-trump, we observe diagnostic plots for Donald Trump's multi-linear regression model.
+
+The Residuals vs Fitted plot (left) shows that the residuals are scattered fairly evenly around the horizontal line at zero, indicating no major patterns of bias across the range of fitted values.
+However, the slight curve in the middle suggests potential mild non-linearity or missing variables in the model.
+A few outliers, such as observations 245, 379, and 263, stand out and may warrant further investigation to assess their influence on the model's performance.
+
+The Q-Q plot (right) for standardized residuals displays a relatively straight line, implying that the residuals are approximately normally distributed.
+There are slight deviations at both extremes, indicating some outliers or deviations from normality at the tails.
+This could hint at potential issues in capturing extreme values accurately but generally suggests the model performs well in satisfying the normality assumption.
 
@@
 ```
+@fig-harris-vif, we see diagnostic plots for the refined multi-linear regression model for Kamala Harris.
+
+The Residuals vs Fitted plot (left) shows that the residuals are relatively evenly distributed around the horizontal line at zero, indicating that there are no major issues with non-linearity or heteroscedasticity.
+The slight curve in the red line suggests minimal deviation from linearity, which could be due to the data's inherent characteristics or model limitations.
+There are a few noticeable outliers, like observations 245, 352, and 370, which might need further investigation to ensure they do not disproportionately influence the model.
+
+The Q-Q plot (right) demonstrates that the residuals generally follow a normal distribution, as indicated by the points aligning closely with the 45-degree line.
+However, some deviations at the tails suggest the presence of outliers or slight departures from normality.
+This generally supports the adequacy of the model but points to areas where model refinement or robust handling of outliers might improve fit.
 
@@
 ```
+@fig-harris-step, presents the diagnostic plots for the final multi-linear regression model for Kamala Harris.
+
+The Residuals vs Fitted plot (left) reveals a fairly uniform spread of residuals around the zero line, suggesting that the model captures the linear relationship well without major systematic errors.
+The red line, representing the smoothed residuals, remains close to zero, with slight deviations indicating minimal non-linearity.
+A few outliers, notably points 245, 352, and 370, are present but do not seem to heavily distort the overall pattern, though their influence may need checking.
+
+The Q-Q plot (right) shows that the residuals align closely with the theoretical quantiles of a normal distribution, confirming that the assumption of normally distributed errors holds reasonably well.
+The points at the extremes deviate slightly, indicating minor issues with the tails of the distribution, but these are not severe.
+This suggests that the model assumptions are largely satisfied, making the model robust for the intended analysis.
 

### 2024-11-04 paper/paper.qmd

@@
 
+### Google Survey Link
+https://forms.gle/81vCBvH8t7B9NfJ69
+
+### Google Survey 
+**U.S. Presidential Election Survey 2024**
+
+-This survey is part of a research study aimed at forecasting the 2024 U.S.
+Presidential Election.
+Your responses are confidential and will be used solely for research purposes.
+This survey will take approximately 5 minutes to complete.
+
+**Are you eligible to vote in the 2024 U.S. Presidential Election?**
+
+-Yes
+-No
+
+**Are you planning to vote in the upcoming 2024 U.S. Presidential Election?**
+
+-Yes
+-No
+-Maybe
+
+**What is your current party affiliation?**
+-Democratic Party
+-Republican Party
+-Independent
+-Other (please specify)
+-Prefer not to say
+
+**Did you vote in the 2020 U.S. Presidential Election?**
+-Yes
+-No
+-Prefer not to say
+
+**If the presidential election were held today, which candidate would you most likely vote for?**
+-Kamala Harris (Democratic Party)
+-Donald Trump (Republican Party)
+-Another candidate (please specify)
+-Undecided
+-Prefer not to say
+
+**How likely are you to vote in the 2024 U.S. Presidential Election?**
+
+Least likely | 1 | 2 | 3 | 4 | 5 | Most likely
+
+**What is your age?**
+
+**What is your gender?**
+-Male
+-Female
+-Non-binary/Third gender
+-Prefer not to say
+
+**What is your race/ethnicity?**
+-White
+-Black or African American
+-Hispanic or Latino
+-Asian
+-Native American or Alaska Native
+-Native Hawaiian or Pacific Islander
+-Prefer not to say
+-Other (please specify)
+
+**What is your highest level of education?**
+-Less than high school
+-High school graduate
+-Some college
+-Bachelor’s degree
+-Postgraduate degree
+-Prefer not to say
+
+**What is your approximate annual household income?**
+-Less than 25,000
+-25,000 to 49,999
+-50,000 to 74,999
+-75,000 to 99,999
+-100,000 or more
+-Prefer not to say
+
+**To ensure the quality of our data, please select "Blue" from the options below:**
+
+- Red
+- Blue
+- Green
+- Yellow
+
+---
+
+**Conclusion**
+
+Thank you for your participation!  
+Your responses have been recorded.
+
 ### Conclusion
